\chapter{Concept and Architecture}\label{ch:concept}

This chapter presents the core contribution of the thesis: an axis library framework that decomposes
cache-aware search algorithms into interchangeable, orthogonal building-block axes, the layered
engine architecture that realizes it, the ABI-stable module interface, the nineteen-axis permutation
matrix, the probe PRT-ART, and the builder that orchestrates the three-stage evaluation.

\section{The Axis Library Framework}\label{sec:axis-framework}

The central idea is to treat an algorithm not as a monolith but as a
\emph{composition of orthogonal axes}, each axis representing one sub-decision of a cache-aware
index (how pages are laid out, how nodes branch, how memory is allocated, how concurrency is
handled, and so on). A concrete algorithm is then a point in the Cartesian product of all axes, and
a state-of-the-art design is reconstructed as one explicit configuration. The decisive property is
\emph{modular interchangeability}: a researcher can study one new axis variant in isolation --- and
measure its effect, per axis and for the whole algorithm --- without re-implementing all the others.
This turns the comparison of two \texttt{std::map} implementations
(Section~\ref{sec:stdmap-interface}) into a controlled, reproducible experiment.

This decomposition rests on two observations. First, search algorithms share, in the large, the same
\emph{tree anatomy} in different expressions --- from balanced search trees through tries and radix
trees to custom constructs; even a sorted key--value list is, strictly speaking, a single leaf of a
tree whose values point onward to data leaves. The tree structure is therefore universal across
search-algorithm architectures and forms the canvas on which the axes are placed. Non-tree-like
search methods --- e.g.\ a hash table --- fit into the same axis system by assigning the
tree-specific axes (path compression, node type, range index) pass-through variants and carrying
only the shared axis subset; they remain the same genus with a restricted profile. A separate genus
would arise only if its base axis set differed fundamentally. Second, an instructive metaphor treats
a concrete algorithm as an \emph{organism} whose \emph{organs} are the axes: every algorithm has a
node-layout organ, a traversal organ, an allocation organ, and so on, each organ in a different
expression. Permuting the axes then resembles a genetic experiment --- exchanging organs between
organisms yields new composite algorithms with previously unknown cache-line behaviour. The research
goal is to find this shared anatomy and, within the resulting design
space~\cite{idreos2018periodic, idreos2018datacalculator}, the fastest composite for a given
data-type-specific load pattern. It is precisely this modular interchangeability that solves the
separability problem of the task: because each axis can be swapped against its alternatives in
isolation while all other axes stay fixed, the weighting contribution of every single constituent to
the overall performance becomes individually measurable.

\section{Four-Subsystem Separation (M Model)}\label{sec:m-model}

The diploma project, the cache-engine builder, the cache engine, and the probe are
\emph{four formally separated subsystems} with distinct responsibilities. The separation is
orthogonal to the three-repository split (Section~\ref{sec:repos}): builder and engine live in the
same repository but are an executable and a library, respectively.

\begin{enumerate}
  \item \textbf{\texttt{messung\_driver}} (Diplomarbeit/Code) --- the evaluation orchestrator (outer
  loop): reads the three measurement-series XML configurations (A/B/C), triggers the builder per
  series, and collects the results for the appendix.
  \item \textbf{CacheEngineBuilder} (\texttt{cache-engine/apps}) --- an autonomous platform-probing
  system running, via the \texttt{ExperimentDriver} library, the seven-phase pipeline (enumerate,
  codegen, compile, load, execute, measure, persist) --- plus two opt-in phases (hot-compile of
  missing modules, ABI-contract functional test) --- and enumerating the permutation space per
  registered engine.
  \item \textbf{CacheEngine} (\texttt{cache-engine/libs}) --- the tool library: it provides, for
  each of the nineteen building-block axes (Section~\ref{sec:axes}), the default implementations and
  the associated cache-strategy families as a C++23 concept API, from which the builder assembles
  the permutation configurations.
  \item \textbf{Probe (PRT-ART)} (\texttt{prt-art}) --- an algorithm implementation registered with
  the cache engine as an \texttt{IExecutionEngine}; it supplies its own axis organs and consumes
  cache-engine services during its own lookup/insert/scan operations.
\end{enumerate}

The defining property of the M model is the \emph{bidirectional}
cache-engine~$\leftrightarrow$~probe relationship: (a) the cache engine instantiates the probe per
permutation configuration (phase 5, \textsc{execute}: \texttt{create\_instance}), and (b) the probe
calls cache-engine services (telemetry, prefetch, heuristics) during its own lookups at run time
(also phase 5, \texttt{run\_workload}); aggregation happens in phase 6 (\textsc{measure}). This
symmetry enables the \emph{multi-probe} capability: measurement series~A compares PRT-ART against
many SOTA adapters in a single builder pipeline.

Here the \texttt{ExecutionEngine} is not a fifth subsystem but the common measurable root --- the
\texttt{IExecutionEngine} interface through which these four subsystems are connected: the cache
engine provides the axis building-block library, while the ExecutionEngine contributes the
orthogonal execution and measurement infrastructure --- as the abstract root over everything
measurable (Section~\ref{sec:three-layer}), under which the probe and every engine register as an
\texttt{IExecutionEngine}.

\section{The One Architecture}\label{sec:three-layer}

The model has \emph{one} architecture --- a single, self-contained hierarchy with no parallel trees.
At its root stands the \texttt{ExecutionEngine} as the common, measurable abstraction over
everything measurable; beneath it the only branching is between \emph{living organisms} (anatomies
with axis organs) and axis-less \emph{viruses} (e.g.\ pure graph algorithms), as siblings of the
same root.

\scriptsize
\begin{verbatim}
    IExecutionEngine                               root: everything measurable
      |
      +-- IAnatomyBase : IExecutionEngine          living organism
      |     |
      |     +-- SearchAlgorithm subclass           = "organism" (mammal)
      |           |
      |           +-- SearchAlgorithmAnatomy<C>     = its BODY (19 organs)
      |           +-- SearchAlgorithmAbiAdapter<A>  = ABI runtime VIEW ("SearchEngine")
      |
      +-- IVirusExecutionEngine : IExecutionEngine  virus (axis-less, graph)
\end{verbatim}
  \normalsize

  In this model a search algorithm is a \emph{living organism}, and ``organism'' and
  \texttt{SearchAlgorithm} denote the same thing --- metaphor and technical term are two names for
  one entity. The organism is addressed from outside through its genus interface (the test dock, a
  \texttt{std::map}-like interface); its interior is its \emph{anatomy}. The anatomy is not a second
  model beside the search algorithm but its \emph{body}: the fixed composition of its nineteen axes,
  each corresponding to an organ. Across the module boundary the same organism is exposed as an ABI
  view --- the ABI/runtime layering
  \texttt{CacheEngine}\,$\to$\,\texttt{ExecutionEngine}\,$\to$\,\texttt{SearchEngine} is exactly
  this view of the same organism across the module boundary, not a parallel construct:
  \texttt{CacheEngine} and \texttt{ExecutionEngine} have clearly separated roles here --- the
  \texttt{CacheEngine} is the \emph{axis building-block library} and provides the default axis
  building blocks, whereas the \texttt{ExecutionEngine} supplies the
  \emph{execution and measurement tools}: the OS primitives (cache-line-aligned allocate,
  NUMA-conforming read-pin, coherence-aware write) and, in experiment mode, the measurement
  aggregator over exactly one observer interface. ``SearchEngine'' is the search-API view of the
  same organism, measured through exactly this one observer interface over all nineteen organs.
  There is thus a single chain --- root (\texttt{ExecutionEngine}) through the genus test dock down
  to the organism, which is the same thing as \texttt{SearchAlgorithm} and whose body is the anatomy
  with its nineteen axes.

  \section{ABI-Stable C++23 Module Interface}\label{sec:abi}

  The ABI follows the variadic-template principle: one parameter $\Rightarrow$ \texttt{std::vector}
  API, two parameters $\Rightarrow$ \texttt{std::map} API, $N>2 \Rightarrow$
  \texttt{std::map<K,\,std::tuple<V$_1$\ldots V$_N$>>}. Complex keys are reduced to fixed-length
  16-byte binary strings by a fingerprint component. The module boundary is binary-stable so that
  the builder can load pre-compiled permutation modules (LoadLibrary/dlopen) and verify the ABI
  contract per module.

  \section{The Building-Block Axes}\label{sec:axes}

  After the N-phase extension and the consolidation into the search-algorithm anatomy, the matrix
  comprises \textbf{nineteen main axes plus roughly fifty-seven sub-axes}: (1)~search-algorithm,
  (2)~cache-traversal, (3)~mapping, (4)~path-compression, (5)~node-type, (6)~memory-layout,
  (7)~allocator (allocation, reclamation, NUMA, huge-page, free-list), (8)~prefetch, (9)~concurrency
  (pattern, locking-mode), (10)~serialization, (11)~telemetry, (12)~value-handle, (13)~ISA,
  (14)~index-organization, (15)~I/O-dispatch, (16)~migration, (17)~filter, (18)~queuing-q1, and
  (19)~queuing-q2 --- the seventeen search axes plus the two queuing axes, exactly the nineteen
  organs that the \texttt{IsComposition} concept enforces per algorithm. Per axis a
  \texttt{std::variant} enumerates all concretizations (the allocator axis alone offers about two
  dozen); the Cartesian product spans far more than $10^{11}$ configurations
  (Section~\ref{sec:explosion}).

  \section{PRT-ART as Probe}\label{sec:prt-art}

  PRT-ART is a \emph{probe}: an abstract organism that supplies some axis organs itself and, for the
  rest, falls back at compile time (\texttt{resolve\_baustein.hpp}) onto the cache-engine default
  building blocks where it does not override an axis; the hybrid
  \texttt{PrtArtSearchEngine<Ts\ldots>} API stays untouched. Its run-time binding is through the
  execution-engine adapter, not through a search-engine layer of its own. PRT-ART contributes its
  own implementations in the page, layout, free-list, prefetch, concurrency-pattern, and measurement
  axes (4+2 pool allocator, distance-estimator prefetch, OLC with reserved value blocks, and the
  H1/H2/H3 metrics); for the remaining axes it reuses existing SOTA building blocks via permutation
  configuration.

  \section{CacheEngineBuilder and Three-Stage Evaluation}\label{sec:builder}

  The builder orchestrates a three-stage evaluation: \textbf{stage~1} (cache-engine only) uses the
  default variant per axis; \textbf{stage~2} (probe only) replaces the overridden axes with the
  probe variants and keeps cache-engine defaults elsewhere; \textbf{stage~3} (full join,
  non-redundant) combines the default and all probe variants. Each stage emits thousands of
  non-redundant permutation binaries, each loaded at run time as an ABI-stable C++23 module. Stages
  map directly onto the three mandatory measurement series of Chapter~\ref{ch:methodology}. From the
  fine-grained measurement results the pipeline already provides the required ranking; the automatic
  selection of the best overall algorithm and its delivery as a standalone, ABI-stable binary behind
  the unified search-algorithm interface are envisioned as an extension (objective, outlook).

  \subsection{Heuristic Profile Evaluation and Binary Selection}\label{sec:heuristic-profile-selection}

  Beyond the bare ranking, an end-to-end evaluation step is envisioned as an objective that turns
  the measurement results into a delivered configuration. The following three stages are
  \emph{outlook and objective} and, in the current state, \emph{not yet implemented}.

  \emph{First}, from the ranking of the three measurement series~A/B/C the best configuration is
  determined per load profile and data type. Since each series yields a complete axis measurement
  vector per permutation binary, for a given load/data-type profile one can identify the axis
  composition that holds the foremost rank under that profile.

  \emph{Second}, the intended selection mechanism --- which infers the best configuration
  automatically from the measurement results --- is an \emph{ML classification}: from the axis
  features and the features of the load profile (access distribution, write/read ratio, key/value
  data type) the suitable axis configuration is predicted. This heuristic selection mechanism is
  \emph{future work} and not yet realized.

  \emph{Third}, the result of this evaluation is stored as a reusable \emph{XML load profile} that
  binds the determined best configuration to a named load/data-type profile and can thus be used
  directly as input on the next run --- symmetric to the measurement-series XML configurations from
  Section~\ref{sec:m-model}.
