\chapter{Implementation and Concrete Algorithm Components}\label{ch:impl}

% [SKELETON] Detail focus. After a brief known SOTA, each PRT-ART algorithm as a figure + description;
% then axis implementation (one overall-concept figure per axis) and all levels of the measurement system.

This chapter descends into detail: first the repository architecture, then each key point of PRT-ART individually
and visually, finally the axis implementation and the levels of the measurement system.

\section{Three-Repository Architecture}\label{sec:repos}
The implementation is distributed across three Git repositories with clearly separated responsibilities:
\begin{itemize}
  \item \texttt{comdare-cache-engine} --- \textbf{how} measurement is done: the building-block library and the builder
  (XML config parser, \texttt{generate\_\allowbreak module\_\allowbreak from\_\allowbreak profile} code
  generation, permutation loop, module loader, and the seven-phase \texttt{ExperimentDriver}), the ABI
  headers, 30~SOTA profiles under \texttt{algorithm\_profiles/sota}, and the test suites; the source tree
  follows a Pitchfork layout (\texttt{apps/}, \texttt{libs/}, \texttt{libs/common}).
  \item \texttt{comdare-prt-art} --- the \textbf{probe}: the hybrid \texttt{PrtArtSearchEngine}
  (vector/map/tuple API), its ABI adapters, and its building blocks (4+2 allocator pools, OLC with reserved
  blocks, multi-level layout, path-oriented prefetch, density tracker, hypothesis metrics H1/H2/H3).
  \item \texttt{probst-Diplomarbeit-cache-engine} --- \textbf{what} is tested: the \texttt{messung\_driver},
  the experiment configurations, the evaluation pipeline, and the manuscript.
\end{itemize}
Orthogonally, the terminology separates two engine terms that must not be conflated. The
\texttt{CacheEngine} (\texttt{cache-engine/libs}) is the \emph{library of axis building blocks}: it
provides per axis the interchangeable design constituents as concept/CRTP components and contains \emph{no}
measurement logic. The \texttt{ExecutionEngine}, by contrast, is the execution and measurement
infrastructure lying across all axes: as the abstract measurable root (\texttt{IExecutionEngine},
Section~\ref{sec:measurement-system}) the concrete \texttt{execution\_engine<Strategy>} provides the OS
primitives (cache-line-aligned allocation, NUMA-conformant read pinning, coherence-aware writing) and ---
exclusively in experiment mode (\texttt{-DCOMDARE\_\allowbreak EXPERIMENT\_\allowbreak MODE=ON}) --- the
measurement aggregator. In short: the \texttt{CacheEngine} provides the \emph{building blocks}, the
\texttt{ExecutionEngine} the \emph{tools}; this separation keeps the production path measurement-free.

\section{PRT-ART in Detail}\label{sec:prtart-impl}
Building on the known state of the art, we describe each key point of PRT-ART individually and visually.
PRT-ART contributes several key points of its own, each at a different axis. On the \emph{allocator axis}, a
\emph{4+2 pool allocator} (four size-class pools plus two reserved) replaces the algorithm-hardwired
allocation with a cache-line-aligned, low-fragmentation free list. On the \emph{concurrency axis} --- itself
split into pattern (8.1) and locking mode (8.2) --- PRT-ART contributes \emph{optimistic lock coupling (OLC)
with reserved value blocks}, coordinated by a concurrency manager that handles the disciplines OLC, HTM,
STM, lock-free, RCU, and hazard pointers; reclamation is here a sub-axis of the allocator (epoch, RCU,
hazard-pointer, and QSBR strategies selectable per permutation). On the \emph{prefetch axis} a
\emph{path-oriented bundle prefetch} fetches the hot search paths ahead of time, on the \emph{layout axis} a
\emph{multi-level layout}. Particularly cache-critical is the \emph{telemetry axis}: per-node counters cause
cache-line ping-pong on multi-core systems and falsify precisely the measurements they collect. Following
the Kuehn research line~\cite{kuehn2023bplustree}, PRT-ART therefore offers cache-coherence-aware strategies
--- \emph{leaf-only counters} (the main variant), \emph{sampling} (every $N$th leaf access), and
\emph{offline recompute} (bottom-up aggregation before reordering) ---, while the classical per-node counter
remains only as a labelled anti-pattern. The density tracker and the hypothesis metrics H1/H2/H3 the probe
adds as its own telemetry organs.
Condensed into a gallery, Figure~\ref{fig:prtart-gallery} shows the six own key points of PRT-ART, each
located at its axis.
\begin{figure}[htbp]\centering
\resizebox{0.97\linewidth}{!}{%
\begin{tikzpicture}[font=\footnotesize,
  pa/.style={draw,rounded corners,fill=orange!10,minimum width=4.7cm,minimum height=2.25cm,text width=4.4cm,align=center,inner sep=3pt}]
  \node[pa] (p1) at (0,0){\textbf{4+2 pool allocator}\\{\scriptsize axis T6}\\[2pt]{\tiny four size-class pools $+$ two reserved; cache-line-aligned, low-fragmentation free list instead of hard-wired allocation}};
  \node[pa] (p2) at (5.1,0){\textbf{OLC $+$ reserved value blocks}\\{\scriptsize axis T8}\\[2pt]{\tiny optimistic lock coupling; concurrency manager over OLC/HTM/STM/lock-free/RCU/hazard; reclamation as allocator sub-axis}};
  \node[pa] (p3) at (10.2,0){\textbf{Path-oriented bundle prefetch}\\{\scriptsize axis T7}\\[2pt]{\tiny loads the hot search paths as a bundle ahead of time; entry points from the subtree footprint (\texttt{p27\_bundle\_finder})}};
  \node[pa] (p4) at (0,-2.85){\textbf{Multi-level layout}\\{\scriptsize axis T5}\\[2pt]{\tiny multi-level memory layout across the node levels; layout-driven stride (CLA 64-byte vs.\ AoS 16-byte)}};
  \node[pa] (p5) at (5.1,-2.85){\textbf{Coherence-aware telemetry}\\{\scriptsize axis T10}\\[2pt]{\tiny leaf-only counters (main), sampling (every $N$th), offline recompute; per-node counter only as anti-pattern}};
  \node[pa] (p6) at (10.2,-2.85){\textbf{Density tracker $+$ H1/H2/H3}\\{\scriptsize axis T10 (own organs)}\\[2pt]{\tiny own measurement organs: occupancy density tracker and the probe's three hypothesis metrics}};
\end{tikzpicture}}
\caption{Gallery of the six own key points of PRT-ART, each at its axis: a 4+2 pool allocator (T6), optimistic lock coupling with reserved value blocks (T8), path-oriented bundle prefetch (T7), a multi-level layout (T5), and cache-coherence-aware telemetry (T10) together with the density tracker and the hypothesis metrics H1/H2/H3. The remaining fourteen composition axes the probe draws via the compile-time fallback from the cache-engine.}\label{fig:prtart-gallery}
\end{figure}

\section{Axis Implementation and Measurement-System Levels}\label{sec:axes-impl}
Each axis is implemented as a generalized interface; alongside lie the levels of the measurement system from code
generation to the seven-phase pipeline.
Each axis is a topic directory containing a two-layer \emph{concept} (a broad topic concept and a narrow
axis concept) and a CRTP base class secured by a \texttt{requires} clause. The probe-versus-default choice
per axis is resolved at compile time via \texttt{resolve\_baustein.hpp} (\texttt{std::conditional\_t} over
tag specializations), so that the hot path contains neither \texttt{virtual} calls nor runtime switches ---
the zero-cost abstraction from Section~\ref{sec:software-means}. Every SOTA design and every allocator is
integrated via a thin \emph{adapter} that provides the original implementation behind the engine ABI: a
\texttt{COMDARE\_HAVE\_<X>} compile flag guards an \texttt{\#if defined} block with a safe fallback (e.g.\
\texttt{std::malloc}), so that the build remains independently runnable; the integration of hierarchical
prefetching~\cite{zhang2025hierarchical} additionally provides a compile-time call-graph analyzer
(\texttt{p27\_bundle\_finder}, a C++23 port of the original tool \texttt{hp\_soft} that identifies
functions whose subtree footprint exceeds a threshold as bundle entry points).

The builder transforms each profile into a pre-compiled C++23 module; each permutation is identified by a
bit-packed \emph{permutation flag} structure that grows, with the N-phase refinement, from nine banks
(around 50~bits) to a wider, multi-bank \emph{permutation identifier} with sub-bank bit fields for the
nineteen axes and their sub-axes; a constraint filter excludes invalid combinations before each module
generation. Over these building blocks runs the \emph{seven-phase measurement pipeline} (Figure~\ref{fig:pipeline}) of the
\texttt{ExperimentDriver} per series: (1)~\textbf{Enumerate} (parse XML configs, build the permutation
list), (2)~\textbf{Codegen} (one \texttt{comdare\_perm\_<fp>.cpp} per permutation), (3)~\textbf{Compile},
(4)~\textbf{Load} (LoadLibrary/dlopen of all modules), (5)~\textbf{Execute} (\texttt{create\_instance} +
profile-aware \texttt{run\_workload}), (6)~\textbf{Measure} (aggregate the binary measurement records), and
(7)~\textbf{Persist} (\texttt{measurements.csv}/\texttt{.json}); two opt-in phases hot-compile missing
modules and verify the ABI contract per module. The downstream evaluation toolchain is a sequence of small
programs --- \texttt{sample\_data\_generator} (deterministic test data without real hardware) $\to$
\texttt{messung\_driver} (drives the series) $\to$ \texttt{binary\_to\_csv} (deserializes the binary
measurement records) $\to$ \texttt{csv\_to\_latex} (booktabs tables with per-profile algorithm fact sheets)
$\to$ \texttt{diagram\_generator} (A4-aware TikZ plots) $\to$ \texttt{latex\_to\_pdf} ---; the
measurement path is compiled in only under \texttt{-DCOMDARE\_\allowbreak EXPERIMENT\_\allowbreak MODE=ON},
so that the production path carries no measurement overhead, and continuous integration (GitLab CI plus a
mirrored GitHub workflow) runs across all three repositories.
% AXIS-T0..T18 gallery: fig:axes-gallery (4.3) DONE
\begin{figure}[htbp]\centering
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[font=\footnotesize,>=Stealth,align=center,
  ph/.style={draw,rounded corners,fill=blue!6,minimum width=1.6cm,minimum height=0.85cm},
  hi/.style={draw,rounded corners,fill=orange!18,very thick,minimum width=1.6cm,minimum height=0.85cm},
  opt/.style={draw,rounded corners,dashed,fill=orange!8,minimum width=2.0cm,minimum height=0.65cm,font=\scriptsize},
  pb/.style={draw,rounded corners,fill=orange!16,very thick,minimum width=3.7cm,minimum height=1.05cm,text width=3.5cm,font=\scriptsize}]
  \node[font=\bfseries\scriptsize,anchor=west] at (-0.95,0.95){Path A --- \texttt{ExperimentDriver} (in-process, 7 phases)};
  \node[ph] (p1) at (0,0){\textbf{1}\\Enumerate};
  \node[ph] (p2) at (1.95,0){\textbf{2}\\Codegen};
  \node[hi] (p3) at (3.9,0){\textbf{3}\\Compile};
  \node[ph] (p4) at (5.85,0){\textbf{4}\\Load};
  \node[hi] (p5) at (7.8,0){\textbf{5}\\Execute};
  \node[hi] (p6) at (9.75,0){\textbf{6}\\Measure};
  \node[ph] (p7) at (11.7,0){\textbf{7}\\Persist};
  \foreach \a/\b in {p1/p2,p2/p3,p3/p4,p4/p5,p5/p6,p6/p7}{\draw[->,thick] (\a)--(\b);}
  \node[font=\scriptsize,text=orange!60!black] at (3.9,0.66){\textsc{Build}};
  \node[font=\scriptsize,text=orange!60!black] at (8.78,0.66){\textsc{Measure}};
  \node[opt] (o1) at (2.9,-1.4){opt-in:\\hot-compile};
  \node[opt] (o2) at (4.9,-1.4){opt-in:\\ABI verify};
  \draw[->,dashed] (o1)--(p3);
  \draw[->,dashed] (o2)--(p4);
  \node[font=\bfseries\scriptsize,anchor=west] at (-0.95,-2.5){Path B --- \texttt{BuildOrchestrator} $+$ \texttt{perm\_runner} (out-of-process, recommended)};
  \node[pb] (b1) at (2.0,-3.95){\textbf{BUILD} \texttt{Build\-Orchestrator}: builds \emph{all} tier binaries first (resumable, RAM-aware)};
  \node[pb] (b2) at (8.0,-3.95){\textbf{MEASURE} \texttt{perm\_runner}: loads $+$ measures each tier binary individually (1 DLL/process, via \texttt{IObservableTier})};
  \draw[->,very thick] (b1)--node[above,font=\scriptsize]{tier binaries} (b2);
\end{tikzpicture}}
\caption{The measurement pipeline in two forms. \textbf{Path~A} (top) is the in-process seven-phase pipeline
of the \texttt{ExperimentDriver} per measurement series (Enumerate $\to$ Codegen $\to$ Compile $\to$ Load
$\to$ Execute $\to$ Measure $\to$ Persist; two opt-in phases hot-compile missing modules and verify the ABI
contract). Highlighted are the genuinely substantial phases: \textsc{Compile} (the \emph{build} of the tier
binaries) and \textsc{Execute}/\textsc{Measure} (the \emph{measurement}); Codegen/Load/Persist are pure
plumbing. \textbf{Path~B} (bottom, the recommended route) separates the two tasks out-of-process: the
\texttt{BuildOrchestrator} builds \emph{all tier binaries first} (resumable, RAM-aware), after which
\texttt{perm\_runner} loads and measures each binary individually (one DLL per process) through the observer
interface \texttt{IObservableTier}.}\label{fig:pipeline}
\end{figure}
The nineteen composition axes are not abstract but each a concrete \emph{strategy concept} with several
interchangeable building blocks, from which a concrete composition picks exactly one per slot.
Figure~\ref{fig:axes-gallery} gathers all nineteen as a gallery --- per axis the short name, the sub-axis
family and a representative selection of the building blocks; the complete building-block matrix with all
origin sources is given in Appendix~\ref{app:blocks}, the SOTA contributions per axis in Table~\ref{tab:axes-overview}.
\begin{figure}[htbp]\centering
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[font=\footnotesize,
  ax/.style={draw,rounded corners,fill=blue!7,minimum width=3.35cm,minimum height=2.0cm,text width=3.15cm,align=center,inner sep=2pt}]
  \node[ax] (t0) at (0,0){\textbf{T0 Search algorithm}\\{\tiny SA1--SA4}\\{\tiny Array256, OriginalART/HOT, KAry, Interpol., Hash \dots\,(17)}\\{\tiny\itshape Def.\ Array256}};
  \node[ax] (t1) at (3.6,0){\textbf{T1 Cache traversal}\\{\tiny CT1--CT2}\\{\tiny Linear-fanout, Hash-lookup, Binary-fanout}\\{\tiny\itshape Def.\ Linear}};
  \node[ax] (t2) at (7.2,0){\textbf{T2 Slot mapping}\\{\tiny MP1--MP2}\\{\tiny Direct-placement, Pool-relative}\\{\tiny\itshape Def.\ Direct}};
  \node[ax] (t3) at (10.8,0){\textbf{T3 Path compression}\\{\tiny PC1--PC3}\\{\tiny None, Byte-wise (ART), Patricia}\\{\tiny\itshape Def.\ None}};
  \node[ax] (t4) at (14.4,0){\textbf{T4 Node type}\\{\tiny NT1--NT3}\\{\tiny Node4 / 16 / 48 / 256}\\{\tiny\itshape Def.\ Node256}};
  \node[ax] (t5) at (0,-2.4){\textbf{T5 Memory layout}\\{\tiny HM1--HM4}\\{\tiny AoS, Cache-line-aligned, SoA, Bitmap (LOUDS), AoSoA}\\{\tiny\itshape Def.\ Cache-line}};
  \node[ax] (t6) at (3.6,-2.4){\textbf{T6 Allocator}\\{\tiny AA1--AA7}\\{\tiny Hoard, Slab, Mimalloc, Jemalloc, TCMalloc, Buddy \dots\,(25)}\\{\tiny\itshape Def.\ Std-malloc}};
  \node[ax] (t7) at (7.2,-2.4){\textbf{T7 Prefetch}\\{\tiny PF1--PF3}\\{\tiny None, Distance-estimator (ART), Hardware, Path-orient.}\\{\tiny\itshape Def.\ None}};
  \node[ax] (t8) at (10.8,-2.4){\textbf{T8 Concurrency}\\{\tiny CC1--CC2}\\{\tiny None, Blocking, R/W, OLC, Lock-free, RCU, Hazard}\\{\tiny\itshape Def.\ None}};
  \node[ax] (t9) at (14.4,-2.4){\textbf{T9 Serialization}\\{\tiny SR1--SR3}\\{\tiny Raw-binary, VarLen (LEB128), Succinct, LZ4/Delta}\\{\tiny\itshape Def.\ Raw-binary}};
  \node[ax] (t10) at (0,-4.8){\textbf{T10 Telemetry}\\{\tiny TM1--TM3}\\{\tiny Leaf-only counter, Density-tracker, HDR-histogram}\\{\tiny\itshape Def.\ Leaf-only}};
  \node[ax] (t11) at (3.6,-4.8){\textbf{T11 Value handle}\\{\tiny VH1--VH3}\\{\tiny Inline, External-pool, RCU-shared, MVCC, Chain-ref}\\{\tiny\itshape Def.\ Inline}};
  \node[ax] (t12) at (7.2,-4.8){\textbf{T12 ISA}\\{\tiny IS1--IS3}\\{\tiny Amd64 (SSE2), AArch64 (NEON), PowerPC, RISC-V}\\{\tiny\itshape Def.\ Amd64}};
  \node[ax] (t13) at (10.8,-4.8){\textbf{T13 Index org.}\\{\tiny io1--io3}\\{\tiny Clustered, Heap, IOT (embedded), Non-clustered}\\{\tiny\itshape Def.\ Heap}};
  \node[ax] (t14) at (14.4,-4.8){\textbf{T14 I/O dispatch}\\{\tiny IO1--IO3}\\{\tiny In-memory, Buffered, Direct (O\_DIRECT), Mmap}\\{\tiny\itshape Def.\ In-memory}};
  \node[ax] (t15) at (0,-7.2){\textbf{T15 Migration}\\{\tiny MG1--MG3}\\{\tiny None, Hot/Cold, Tier-based, Adaptive (ML)}\\{\tiny\itshape Def.\ None}};
  \node[ax] (t16) at (3.6,-7.2){\textbf{T16 Filter}\\{\tiny FT1--FT3}\\{\tiny Bloom, Cuckoo, Range-SuRF (LOUDS), Xor}\\{\tiny\itshape (no None)}};
  \node[ax] (t17) at (7.2,-7.2){\textbf{T17 Op. buffer}\\{\tiny QS1--QS6}\\{\tiny None, Append, FIFO/LIFO, Ring, Priority \dots\,(15)}\\{\tiny\itshape Def.\ No-buffer}};
  \node[ax] (t18) at (10.8,-7.2){\textbf{T18 Flush policy}\\{\tiny FS1--FS4}\\{\tiny Eager, Lazy, Watermark, Timed, Adaptive-LSM}\\{\tiny\itshape Def.\ Lazy}};
  \node[draw,rounded corners,fill=green!8,minimum width=3.35cm,minimum height=2.0cm,text width=3.15cm,align=center,inner sep=2pt] (leg) at (14.4,-7.2){\textbf{Reading}\\{\tiny Each axis $=$ a \texttt{Concept} with interchangeable building blocks (\emph{Strategy}); a composition picks exactly one per slot.}};
\end{tikzpicture}}
\caption{Gallery of the nineteen composition axes (T0--T18): per axis the implementation concept --- short name, sub-axis family and a representative selection of the interchangeable building blocks. Each axis is a \texttt{Concept} interface with several strategy building blocks (\emph{Strategy} pattern); a concrete composition occupies exactly one per slot. The \emph{default} (\enquote{Def.}) is a derived comparison baseline, not a code flag; for the large axes T0/T6/T17 the total building-block count is given in parentheses. The complete matrix with origin sources is given in Appendix~\ref{app:blocks}.}\label{fig:axes-gallery}
\end{figure}

With this, the apparatus is fully described --- from the repositories through the probe to the nineteen
axes and their measurement-system levels. Chapter~\ref{ch:eval} lays out the methodology by which this
apparatus is assessed.
