\chapter{Implementation}\label{ch:impl}

This chapter describes the implementation thematically (not as a chronological development log): the
three-repository split, the SOTA and allocator adapters, the permutation code generation and flag
system, the concept/CRTP realization of the axes, concurrency and reclamation, the
cache-coherence-aware telemetry, and the measurement pipeline.

\section{Three-Repository Architecture}\label{sec:repos}

The implementation is distributed over three Git repositories with clearly separated
responsibilities:

\begin{itemize}
  \item \texttt{comdare-cache-engine} --- \textbf{how} things are measured: the tool library and the
  builder. Key components: the builder (XML config parser, \texttt{generate\_module\_from\_profile}
  code generation, permutation loop, module loader, and the seven-phase \texttt{ExperimentDriver});
  the ABI headers (\texttt{search\_engine}, \texttt{execution\_engine}, \texttt{baustein\_variants},
  \texttt{resolve\_baustein}); 30 SOTA profiles under \texttt{algorithm\_profiles/sota}; and the
  test suites. The source tree follows a Pitchfork-style layout (\texttt{apps/} for executables,
  \texttt{libs/} for libraries, with \texttt{libs/common} for cross-cutting code).
  \item \texttt{comdare-prt-art} --- the \textbf{probe}: the hybrid \texttt{PrtArtSearchEngine}
  (vector/map/tuple API), its ABI adapters, its building blocks (4+2 allocator pools, OLC with
  reserved blocks, multi-level layout, path-oriented prefetch, density tracker, hypothesis metrics
  H1/H2/H3), and the probe-side reimplementations.
  \item \texttt{probst-Diplomarbeit-cache-engine} --- \textbf{what} is tested: the
  \texttt{messung\_driver}, the experiment configurations, the evaluation pipeline, and the
  manuscript.
\end{itemize}

\section{Adapters for State-of-the-Art Algorithms and Allocators}\label{sec:adapters}

Each SOTA design and each allocator is integrated through a thin adapter that exposes the original
implementation behind the engine ABI. The adapters follow a uniform pattern: a
\texttt{COMDARE\_HAVE\_<X>} compile flag guards an \texttt{\#if defined} block that wires the
original API, with a safe fallback (e.g.\ \texttt{std::malloc} or a typed exception) when the
original is unavailable. This keeps the build self-contained while allowing the original sources to
be activated per dependency. The hierarchical-prefetch integration (P27) additionally ships a
build-time call-graph analyzer (\texttt{p27\_bundle\_finder}, a C++23 port of the original
\texttt{hp\_soft} tool) that identifies functions whose subtree footprint exceeds a threshold as
bundle entry points.

\section{Permutation Code Generation and Flag System}\label{sec:codegen}

The builder turns each profile into a pre-compiled C++23 module
(\texttt{generate\_\allowbreak module\_\allowbreak from\_\allowbreak profile}). Every permutation is
identified by a bit-packed \emph{permutation flag} structure. With the N-phase refinement of the
matrix, the flag system grows from 9 banks ($\sim$50 bit) to a wider, multi-bank bit-packed
\emph{permutation identifier} with sub-bank bitfields, mirroring the nineteen axes and their
sub-axes. A constraint filter masks out invalid axis combinations before any module is generated.

\section{Concept/CRTP Realization of the Axes}\label{sec:concept-crtp}

Each axis is a topic directory containing a two-layer concept (a broad topic concept and a narrow
axis concept) and a CRTP base class secured by a \texttt{requires} clause. The probe-versus-default
selection per axis is resolved at compile time via \texttt{resolve\_baustein.hpp}
(\texttt{std::conditional\_t} over tag specializations), so the hot path contains neither
\texttt{virtual} calls nor run-time switches --- the zero-cost abstraction of
Section~\ref{sec:cpp23}.

\section{Concurrency and Reclamation}\label{sec:concurrency}

Concurrency is itself an axis (8.1 pattern, 8.2 locking mode), realized through a concurrency
manager that coordinates the disciplines OLC, HTM, STM, lock-free, RCU, and hazard pointers.
Reclamation is a sub-axis of the allocator (6.2): epoch-based, RCU, hazard-pointer, and QSBR
strategies are selectable per permutation. PRT-ART contributes OLC with reserved value blocks as its
concurrency variant.

\section{Telemetry against Cache-Coherence Effects}\label{sec:telemetry}

Per-node counters cause cache-line ping-pong on multi-core systems, which distorts the very
measurements they collect. Following the Kuehn line of work (P28), the telemetry axis therefore
offers cache-coherence-aware strategies: \emph{leaf-only counters} (the main variant),
\emph{sampling} (every $N$-th leaf access), and \emph{offline recompute} (bottom-up aggregation
before reordering); the classic per-node counter is retained only as a labeled anti-pattern for
comparison.

\section{Measurement Pipeline}\label{sec:pipeline}

The evaluation tool chain is a sequence of small programs: \texttt{sample\_data\_generator}
(deterministic test data without real hardware) $\rightarrow$ \texttt{messung\_driver} (drives the
series) $\rightarrow$ \texttt{binary\_to\_csv} (deserializes the binary measurement records)
$\rightarrow$ \texttt{csv\_to\_latex} (booktabs tables with per-profile algorithm fact sheets)
$\rightarrow$ \texttt{diagram\_generator} (A4-aware TikZ plots) $\rightarrow$
\texttt{latex\_to\_pdf}. The measurement path in the execution engine is compiled in only under
\texttt{-DCOMDARE\_EXPERIMENT\_MODE=ON}, so the production path carries no measurement overhead.
Continuous integration (GitLab CI plus a mirrored GitHub workflow) runs across all three
repositories.
