\chapter{Evaluation Methodology}\label{ch:methodology}

This chapter describes how measurements are taken; the concrete results follow in
Chapter~\ref{ch:results}.

\section{Three Measurement Series}\label{sec:series}

The architecture prescribes three measurement series, all executed by the
\texttt{messung\_driver} through the \texttt{ExperimentDriver} library, configured either
as three fixed series or via an external \texttt{messreihen.xml}.

\begin{description}
  \item[Series A --- PRT-ART vs.\ the state of the art.] Compares the probe against the
        documented SOTA algorithms. Two modes: \emph{A\_defined} (quick verification
        against the 8 Rank-1 profiles) and \emph{A\_full} (all 30 SOTA profiles; the
        default, per the directive to measure as completely as possible).
  \item[Series B --- cache-engine permutations.] The pure ``state of the art'' baseline:
        all SOTA profiles without the PRT-ART probe.
  \item[Series C --- merge of old and new.] Concrete merge points between PRT-ART building
        blocks and cache-engine permutations, e.g.\ PRT-ART OLC + tcmalloc + ART page;
        PRT-ART 4+2 pool + HOT page; PRT-ART path prefetch + B\textsuperscript{2}-tree
        page; PRT-ART multi-level layout + Wormhole page.
\end{description}

Each of these three series is collected at three granularities (per the thesis assignment):
\emph{micro-benchmarking} compares each individual design component (axis) in isolation,
through the common interface of its category, against load profiles;
\emph{macro-benchmarking} measures the individual \texttt{std::map} interface operations
(insert, point and range lookup, update, delete) against sample loads; and
\emph{whole-system benchmarking} runs the established YCSB load profiles over every
recombination of the building blocks. The ranking across these three granularities yields
which decision at which place is beneficial for which load pattern.

\section{ExperimentDriver Phases}\label{sec:experimentdriver}

The \texttt{ExperimentDriver} runs seven phases per series: (1)~\textbf{enumerate} (parse
the XML configs and build the permutation list); (2)~\textbf{codegen} (one
\texttt{comdare\_perm\_<fp>.cpp} per permutation, plus
\texttt{generate\_module\_from\_profile} for the SOTA profiles); (3)~\textbf{compile};
(4)~\textbf{load} (LoadLibrary/dlopen of all modules); (5)~\textbf{execute}
(\texttt{create\_instance} + profile-aware \texttt{run\_workload}); (6)~\textbf{measure}
(aggregate the binary measurement records); (7)~\textbf{persist}
(\texttt{measurements.csv} + \texttt{.json}). Two opt-in phases hot-compile missing modules
and verify the ABI contract per module.

\section{Hypotheses}\label{sec:hypotheses}

\begin{description}
  \item[H1 (page-type cost)] The mean access cost per page-type variant differs
        systematically by workload: dense page types are expected to win under Zipfian
        (YCSB-A) and range-optimized pages under range scans (YCSB-E).
  \item[H2 (code quality per source)] The code-quality score inherited from the original
        source repository correlates measurably with achieved throughput.
  \item[H3 (inline vs.\ external vs.\ chain distribution)] The observed distribution of the
        three value-handle variants correlates with page density: dense pages favour
        inline, sparse pages external, and PRT-ART heuristics chain-ref.
\end{description}

\section{Workloads and Datasets}\label{sec:workloads}

Phase~5 routes the YCSB workload per permutation from the profile's traversal tag
(Table~\ref{tab:workload-routing}); modules without a profile use the global default
workload.

\begin{table}[h]
\centering
\caption{Profile-aware workload routing}\label{tab:workload-routing}
\begin{tabular}{lll}
\toprule
\textbf{Traversal tag} & \textbf{YCSB workload} & \textbf{Characteristic} \\
\midrule
\texttt{RANGE\_SCAN} / \texttt{RANGE\_HOT\_PATH} & YCSB-E & range queries \\
\texttt{HOT\_PATH} / \texttt{PRTART\_HOT\_PATH} & YCSB-A & 50/50 read/update, Zipfian \\
\texttt{STANDARD} (default) & YCSB-C & read-only \\
\bottomrule
\end{tabular}
\end{table}

The data basis consists of real string datasets of differing characteristics
(Table~\ref{tab:datasets}); each carries a reproducibility record of source, checksum, line
count, preprocessing, and the seed rule for the synthetic miss, update, and delete keys.

\begin{table}[h]
\centering
\caption{Real datasets and their characteristics}\label{tab:datasets}
\begin{tabular}{ll}
\toprule
\textbf{Dataset} & \textbf{Characteristic} \\
\midrule
\texttt{url} & very long keys, long common prefixes, medium alphabet \\
\texttt{dna} & small alphabet, high repetition, deep prefix entropy \\
\texttt{protein} & biological sequences, small to medium alphabet \\
\texttt{xml} & structured strings with recurring prefixes \\
\texttt{tpcds-id} & database-like IDs, sorted and unsorted \\
\texttt{trec-terms} & dictionary/search-index terms with defined miss keys \\
\bottomrule
\end{tabular}
\end{table}

\section{Experimental Platforms}\label{sec:platforms}

Measurements target a local workstation for development verification and the TU~Dresden
ZIH cluster (Barnard CPU, Capella GPU) for the full sweeps, plus the Rank-3 platforms
(Ryzen~9~9950X3D, Intel~i9~14900KS, and an HBM platform). An \texttt{IPlatformProbe}
(phase~1) reports, per platform, the runnable subset of axes (e.g.\ AVX-512 only where
available, HBM-aware only on HBM hardware). Per platform, a scalar, an AVX2, and --- on
Barnard --- an AVX-512 build are prepared; on hybrid CPUs, P- and E-cores are measured
separately (distinct counter domains \texttt{cpu\_core}/\texttt{cpu\_atom}) rather than
aggregated together.

\section{Fairness and Reproducibility}\label{sec:fairness}

Fair comparability is ensured by fixed rules. Every comparison separates a common minimal
mode (common-denominator: external value handles, no PRT-ART special paths) from the PRT-ART
native mode (inline switching, cache engine, page-type scheduler). All methods receive
identical, normalized byte keys and identical miss, update, and delete sets; warmup,
randomization, and thread pinning are logged per platform and workload. Missing capabilities
of a baseline --- e.g.\ range scans of a hash table --- are reported as N/A rather than
covertly emulated through auxiliary structures; for each third-party library, compiler,
flags, ISA path, allocator, and commit hash are logged. For reproducibility, every
configuration is measured in several separate runs whose spread remains part of the
statement: percentiles are determined via HDR histograms rather than averaged.

\section{Permutation Explosion and Reduction}\label{sec:explosion}

With the N-phase extension from 11 to 19 axes plus $\sim$57 sub-axes, the space grows from
$\sim$5.5 billion to far more than $10^{11}$ permutations. Three mechanisms tame the
explosion: (1)~the \texttt{<expected\_workload>} profile filter routes only compatible
workloads; (2)~\texttt{MessreihenMode::Defined} runs only the declared tuples
($\sim$100--1000 per series) that feed the manuscript; (3)~\texttt{Full} (and the more
realistic \texttt{Full-Sampled}, e.g.\ every 1000th permutation) runs only on the ZIH
cluster, respecting axis compatibility reported by the platform probe.
