\chapter{Conclusion and Outlook}\label{ch:conclusion}

% [SKELETON] from old 08_conclusion.

This chapter takes stock: it answers the research questions RQ0--RQ4 from Section~\ref{sec:rqs}, names
the limits of the state reached, and sketches the next steps.

\section{Answering the Research Questions}\label{sec:answers}
\paragraph{RQ0 (Main question).} By how much active, measurement-driven cache-aware decisions outperform
their passive, statically dimensioned counterparts can be quantified conclusively only after the
end-to-end runs per platform (Section~\ref{sec:outlook}). What is prepared is the apparatus that isolates
exactly this difference: the same axis-constant structure is permuted once actively (measurement-driven
page-type, layout, prefetch, and value-placement decision) and once passively (statically dimensioned) and compared
under an identical compiler and flag basis on hybrid CPUs and Sapphire Rapids.

\paragraph{RQ1 (Composition).} The nineteen permutation axes (search-algorithm, cache-traversal, mapping,
path-compression, node-type, memory-layout, allocator, prefetch, concurrency, serialization, telemetry,
value-handle, ISA, index-organization, I/O-dispatch, migration, filter, and the two queuing axes) span a
configuration space of around $10^{14}$ permutations. State-of-the-art designs are reconstructed
unambiguously through algorithm-profile XMLs: 30 SOTA profiles (8 Rank-1 + 22 Rank-2/3) demonstrate that
the axis decomposition carries P01--P33. The variadic API specialization (1/2/$N$ parameters onto
\texttt{std::vector}/\texttt{std::map}/tuple) makes both algorithm- and container-oriented patterns
expressible on one engine. The non-fully-orthogonal axis couplings (node type $\times$ SIMD $\times$ path
compression) are marked as such and explicitly restricted during permutation. The sub-question on
cache-update and cache-line-fill behaviour under \emph{solely} varied ISA/SIMD axis is answered
methodically --- the ISA axis (T12) is isolatable and permutable with an otherwise axis-constant
structure ---; the empirical answer follows with the end-to-end runs (Section~\ref{sec:outlook}).

\paragraph{RQ2 (Measurement methodology).} Measurement is split into a production mode
(\texttt{COMDARE\_\allowbreak MEASUREMENT\_\allowbreak ON} disabled, no overhead) and an experiment
mode: the \texttt{ResultAggregator} is kept host-side in the \texttt{ExperimentDriver} (phases~5--7),
and the experiment mode activates the measurement hooks and the observer coupling in the ABI adapter. Profile-aware workload routing selects the
workload per permutation from the traversal tag, ensuring fair comparison between range-query- and
point-lookup-oriented designs. Each series is collected at three granularities --- micro-, macro-, and
whole-system benchmarking ---, making both individual design components and whole algorithms comparable.
Contradictory findings in the literature on the optimal node/cache-line size (one versus up to sixteen
cache lines) are thus not adopted from incompatible original setups but re-measured axis-isolated, with all
else held equal. A platform counts as successfully verified only once the conformance gate confirms the
\texttt{std::map} contract per module and the platform probe covers the reported axis subset; the
platform-calibrated thresholds themselves are fixed with the real runs.

\paragraph{RQ3 (Probe comparison).} The concrete comparison of PRT-ART against the 30 SOTA profiles is
pending the first measurement runs (Section~\ref{sec:outlook}). The architecture is fully prepared: the
PRT-ART profile, the three mandatory series templates, the \texttt{ExperimentDriver} with auto-pickup of
the SOTA profiles, and the ABI-conformant adapter with the \texttt{std::map} core contract
(conformance subset, Appendix~\ref{app:interfaces}); the full expansion to all seventeen contract
operations including the notify hooks is designated as a follow-up step.
The cheapest local page representation per branch point (point versus compact page) and its
platform-calibrated switching rules are likewise part of the PRT-ART profile.

\paragraph{RQ4 (Compile time vs.\ run time).} The separation is anchored in the build system:
platform-specific axes (ISA, hardware constants) are fixed at compile time and produce
provenance-verifiable binaries per permutation (SHA256-validated original bodies --- whose full implementation Section~\ref{sec:limitations} lists as
open ---,
\texttt{get\_compiler()} identifier), while layout and page selection remain selectable at run time
via a bit-encoded permutation identifier, without violating the common compiler and flag basis. The
forward-looking sub-question of whether the measurement itself can tailor ISA- and
operating-system-optimized binaries for a recombination is taken up by the outlook as heuristics
automation (Section~\ref{sec:outlook}); the operating-system dimension is already anchored in the
measurement plan via the two measurement regimes from Chapter~\ref{ch:eval}.

\section{Limitations}\label{sec:limitations}
\begin{itemize}
  \item \textbf{Module bodies:} the ad-hoc codegen of the active measurement path emits real
  axis-library bodies per permutation (the seventeen fully qualified axis types plus the umbrella
  include); only the legacy profile codegen still produces ABI-conformant skeleton modules, and the
  remaining gaps are the PRT-ART probe stub and the honestly zero-reported storage/prefetch axes of
  the pool families.
  \item \textbf{Hardware validation pending:} configuration is green locally, but the full \texttt{ctest}
  and end-to-end \texttt{messung\_driver} runs across the BlockAO platforms (AVX2/AVX-512, ARM NEON/SVE)
  are outstanding.
  \item \textbf{Results chapter:} currently methodological; concrete measurements follow. A first real
  smoke measurement series (43 permutations) is available in Appendix~\ref{app:measurements}, but does
  not replace the mandatory measurement series.
  \item \textbf{Hardware counters off by default:} PMC capture (L1/L2/L3 and dTLB misses, coherence
  invalidations, energy/RAPL\@; branch misses and IPC are planned as an extension of the counter set and
  are currently not captured) is disabled by default; in the smoke series so far the hardware columns
  are zero --- the cache-aware node/data/index measurement component is thus still pending
  (Figure~\ref{fig:measurement-triangle}).
  \item \textbf{Smoke series with apparatus caveats:} the real measurements so far vary seven of the
  nineteen axes in two sub-blocks (27 SIMD$\times$layout$\times$allocator combinations plus 16
  combinations over node width, path compression, internal search, and telemetry), and partly
  approximate axis effects via wall-clock proxies; single
  effects are apparatus rather than organ artefacts (\enquote{shadowing}). The complete limitation table
  is given in Section~\ref{sec:measurements:limitations}.
\end{itemize}

\section{Outlook}\label{sec:outlook}
In the short term: the end-to-end run of the three mandatory series on real hardware, the generation of
the measurement diagrams, and the closing of the remaining module-body gaps (PRT-ART stub,
honest-0 storage/prefetch axes of the pool families).

\paragraph{Heuristic automation and binary delivery.} The mechanism demanded by the thesis assignment
(sub-task~7), which automatically derives the best configuration from the measurement results, is
deliberately \emph{not yet} implemented and is reserved for a follow-up phase. It presupposes three
components, of which the second is already implemented: (i) a \emph{heuristic/ML classifier} that maps the load-profile and axis
features (operation mix, key characteristics, hardware identifier) to the recommended axis configuration;
(ii) the already implemented \emph{XML parser} (import and export including the extractor) for the heuristic policies stored as an XML load-profile result and reusable per
architecture focus; and (iii) a \emph{binary packager} that delivers the thus-selected overall algorithm
as a standalone binary behind the unified search-algorithm interface. A prerequisite for all three is that
the end-to-end measurement series be fully available beforehand --- only the robust data basis across all
permutations and load profiles makes a measurement-driven, automatically controlled selection meaningful;
until then, the configuration recommendations of the evaluation scheme from
Section~\ref{sec:eval-results} will initially be derived manually once the data are available.

\medskip
In the medium to long term: a GPU adapter (Grace Hopper cluster, ZIH), PIM-allocator integration (A16),
and formal verification of selected permutation classes (StarMalloc-style, A13). The three-repository
architecture lets future students add further probe algorithms analogously to PRT-ART without modifying the
cache-engine building-block library --- a key design principle for the scalability of the research infrastructure.
Beyond the tree-like search structures treated here, the anatomy principle extends through higher-level
abstractions to the further structure classes of Section~\ref{sec:overview} --- hash-based, spatial, as
well as set- and sequence-oriented structures ---, so that the same measurement and permutation apparatus
can be applied to entirely new classes without rebuilding it; herein lies a key lever for future
experimental progress.
