\chapter{Results and Evaluation}\label{ch:results}

This chapter presents the measurement results. Concrete diagrams and tables are generated from the
measurement runs and included from the (language-neutral) \texttt{tikz/} and \texttt{tabellen/}
directories; see the placeholders below. At the time of writing, the methodology and the evaluation
pipeline are in place and verified on sample data; the real-hardware runs are pending
(Section~\ref{sec:limitations}).

\section{Evaluation Pipeline}\label{sec:eval-pipeline}

Per series run, the \texttt{messung\_driver} writes \texttt{Code/\_runs/<date>/<spec\_id>/} with
\texttt{measurements.csv} (one row per permutation: cycles, L1/L2/L3 misses, branches, throughput),
\texttt{measurements.json}, and optional binary records. The evaluation uses
\texttt{binary\_to\_csv} (deserialize and enrich with profile id, paper ref, axes),
\texttt{csv\_to\_latex} (booktabs tables with per-profile fact sheets), and
\texttt{diagram\_generator} (A4-aware TikZ bar charts, scatter plots, heatmaps). The manuscript
consumes these via \texttt{\textbackslash input\{tikz/<spec\_id>/\ldots\}} and
\texttt{\textbackslash input\{tabellen/<spec\_id>\ldots\}}.

\section{Series A: PRT-ART vs.\ State of the Art}\label{sec:result-a}

% After the first messung_driver run, insert e.g.:
%   \input{tikz/A_full/cycles_per_permutation.tikz}
%   \input{tabellen/A_full_summary.tex}
The hypotheses H1--H3 (Section~\ref{sec:hypotheses}) are evaluated here once the measurement data is
available.

\section{Series B: Cache-Engine Permutations}\label{sec:result-b}

% \input{tikz/B_cache_engine_perms/heatmap.tikz}

\section{Series C: Full Join}\label{sec:result-c}

% \input{tikz/C_merge/comparison.tikz}

\section{Axis Sensitivity Analysis}\label{sec:sensitivity}

This analysis directly answers the \emph{separability problem} stated in the thesis assignment:
there it was noted that, in the state of the art, the weightings of the improvements of
\emph{individual} algorithm components cannot so far be separated from one another, because each
publication demonstrates its benefit in its own, incompatible measurement environment. By isolating
each component as an interchangeable axis over an otherwise constant structure, the axis framework
makes the contribution of each individual design decision separable --- for the first time under
otherwise equal conditions --- and thereby attributable in measurable terms.

Using the per-permutation records, this section quantifies, at the micro granularity, which axis
contributes most to the performance variance; macro- and whole-system benchmarking add the operation
and YCSB-load-profile levels. The ranking across the three granularities yields the recommendation
of default configurations per workload class. This closes the arc back to the assignment: the
sensitivity analysis delivers exactly the separated weighting of the individual components whose
absence was the founding problem of the thesis.

\section{Discussion}\label{sec:discussion}

The results are discussed against the leitmotif of the thesis: the transition from heuristic to
informed cache management, i.e.\ whether telemetry-driven, automatically adapted configurations
outperform statically chosen ones.
