% AUTO-GENERATED durch csv_to_latex::write_limitations_longtable (L-e ehrliche Limitierung; lang=en)
% EINE longtable, je Zeile EIN nicht-gefixter Vorbehalt. Zeile 1 = Cache-Misses/PMC (Kernmetrik).
\begin{scriptsize}
\begin{longtable}{@{}r >{\raggedright\arraybackslash}p{5.0cm} >{\raggedright\arraybackslash}p{8.0cm}@{}}
\caption{Honest limitations: unresolved caveats (no finding silently lapses)}\label{tab:le:limitierung}\\
\toprule
\# & Caveat & Status / consequence \\
\midrule
\endfirsthead%
\multicolumn{3}{c}{\tablename\ \thetable{} -- continued}\\
\toprule
\# & Caveat & Status / consequence \\
\midrule
\endhead%
\midrule
\multicolumn{3}{r}{continued on next page}\\
\endfoot%
\bottomrule
\endlastfoot%
1 & \textbf{Cache misses (core metric):} L1/L2/L3 + dTLB + coherence + energy = 0 / not collected & NullPmcSource, available=false; Intel PCM Windows driver + Linux PAPI pending (\#26/P4); cache behaviour only indirect via wall-clock proxy (seg\_memory\_layout\_ns/ns\_per\_op). \\
2 & 15 pinned axes = 0 exchangeability evidence & Only 4 axes vary (search\_algo, node\_type, memory\_layout, prefetch). Pinned (1 value each): cache\_traversal, mapping, path\_compression, allocator, concurrency, serialization, telemetry, value\_handle, isa, index\_organization, io\_dispatch, migration\_policy, filter, queuing\_q1, queuing\_q2. \\
3 & Observer probe counters ($\ast$\_find/$\ast$\_probe/$\ast$\_get) by design $\approx 1$ & Expected behaviour, no phantom: one probe per lookup. Says nothing about internal iterations. \\
4 & Honest-0 of inactive sub-features of pinned strategies & concurrency/migration/io/value\_handle report 0 (feature inactive), not a measurement error. \\
5 & Wall-clock not bit-reproducible & Seed drives keys (deterministic), NOT CPU timing; wall-clock varies run-to-run. \\
6 & RC dimension only nominal & K1: $\times 18 \to \times 3$ degenerated; RC axis does not provide full value breadth. \\
7 & 2/21 scan profiles = no-op (excluded) & ycsb\_e and lp\_range\_scan perform no real scan; excluded for scan diffs (19 valid workloads). \\
8 & Insert profiles measure upserts; stat\_$\ast$ columns include load phase & The insert path is effectively an upsert; the stat\_$\ast$ counters also accumulate the load phase. \\
9 & prefetch measures key values as pseudo-addresses (K9) & The prefetch axis interprets key values as addresses; no real memory prefetch. \\
10 & node\_type + memory\_layout: Q2 step-4 shadowing & Apparatus artefact possible (search\_organ\_ shadowing); search\_algo + prefetch are the least confounded. \\
11 & memory\_layout effect partly sub-noise & Part of the layout differences lies below the measurement noise $\to$ PMC externally gated (\#26). \\
12 & Secondary indirection organs (cache\_traversal/mapping): uint16 key saturation & The uint64 PRIMARY key is narrowed (static\_cast) into the smaller slot\_index/key\_type fields of the PINNED cache\_traversal/mapping organs when auto-coupling (low bits); for records $>$ type width their observer entries collide. Does NOT affect the primary lookup (uint64); observer-only limitation of pinned axes (cf.\ rows 2/4). \\
13 & Capsule-internal (swallowed) OOM in a measured op & An OOM propagated outward now hard-invalidates the measurement (two\_phase\_valid=false). An OOM caught INTERNALLY by tier code via catch(...) (e.g.\ defensive memento skip) is not observable from outside $\to$ documented as a conservative residual caveat, not forced. \\
14 & Timer overhead (\texttt{clock::now()}) not calibrated & The wall-clock segment timings include the non-subtracted intrinsic overhead of the \texttt{std::chrono::steady\_clock::now()} calls (an additive, run-stable ns component per measured segment). NO empty-loop calibration is performed $\to$ documented caveat; affects the absolute ns origin, not the relative per-axis comparisons (G5 audit, MESS-MINOR-7). \\
15 & Design labels = naming convention, no measurement effect & A few pattern labels in code/docs are naming/design caveats with no measurement effect: ``hybrid visitor'' (in reality a memento, no \texttt{accept()}/host visit), plus interpreter/template/decorator labels and \texttt{reserve}/\texttt{if constexpr} hygiene. Pure naming convention $\to$ documented as a caveat, no influence on measured values (G5 audit, PATTERN-MINOR-1..7). \\
\end{longtable}
\end{scriptsize}
