\chapter{Glossary}\label{app:glossary}

% [AP-E2/AP-EN] Mirror of anhang/de/C_glossary.tex. Abbreviations + core terms used in Chs. 1--3.
This glossary summarizes the abbreviations and central terms used in the thesis.

\begin{description}
  \item[ABI] Application Binary Interface --- the binary-stable interface through which the builder
  loads the permutation modules (living beings).
  \item[Adaptive vs.\ dynamic vs.\ measurement-driven] Distinction of terms for adapting the
  algorithm behaviour.\ \emph{Adaptive}: adaptation based on run-time-observed data characteristics
  (e.g.\ adaptive node sizes in ART)\@. \emph{Dynamic}: the run-time selection of the fully compiled
  permutation \emph{module} via the permutation identifier (module loading) --- all nineteen axes are
  baked in per binary at compile time; in addition there are run-time-\emph{variable} parameters of
  dynamic sub-axes (prefetch distance, batch size, inline threshold, pool budget;
  cf.\ Section~\ref{sec:measurement-system}), no run-time
  choice of individual axis strategies.\ \emph{Measurement-driven}: selection of the
  configuration based on empirically collected measurement series --- the leitmotif of this thesis;
  the corresponding selection mechanism is implemented in a first increment
  (\texttt{best\_binary\_selector}: ranking and delivery of the best permutation per metric from the
  measurement CSVs), while the fully automatic, load-profile-spanning selection remains outlook.
  \item[Anatomy] The \emph{body} of a living being: the fixed composition of its nineteen axis
  organs and, above all, their \emph{wiring} --- the generalized interfaces through which one organ
  draws on the services of the others (e.g.\ the shared allocation interface). Not a second model
  beside the search algorithm but its internal structure (cf.\ Section~\ref{sec:anatomy-metaphor}).
  \item[ART] Adaptive Radix Tree~\cite{leis2013art}; reference design with adaptive node sizes
  (Node4/16/48/256).
  \item[Autonomous profile-driven heuristic evaluation] The intended mechanism that automatically
  derives, per load profile, the best axis configuration from the measurement series and stores it
  as a reusable heuristic policy (target mechanism: ML classification of load-profile/axis features
  onto configurations). First building blocks are implemented --- the automatic best-binary selection
  from measurement series (\texttt{best\_binary\_selector}) and the extraction/storage as an XML
  load-profile result ---; the ML classification itself remains outlook
  (cf.\ Section~\ref{sec:outlook}).
  \item[AVX, AVX-512] x86 SIMD extensions (256- and 512-bit, respectively).
  \item[Axis] An orthogonal partial decision (``organ'') of a search algorithm; the framework
  maintains nineteen composition axes.
  \item[CLU] Cache-line utilization --- the fraction of used bytes per loaded cache line.
  \item[CPI, IPC] Cycles per instruction and instructions per cycle, respectively.
  \item[CRTP] Curiously Recurring Template Pattern --- C++ idiom for static polymorphism.
  \item[dTLB] Data translation lookaside buffer; caches data address translations.
  \item[Genus] An externally visible interface (SearchAlgorithm, Container, Graph) --- the topmost
  level (level~1) of the framework; see \emph{Three-level model}.
  \item[Genus subclass] The second level beneath a genus, with a fixed axis set (code term:
  \texttt{AnatomyGenus}, ``animal subclass''). In the biological reading the ``mammal'' = SearchAlgorithm (the full
  nineteen-axis anatomy); under \texttt{Container} the subclasses Set, Sequence, Adapter, and View
  (``bird'', ``reptile'', ``invertebrate'', ``plant'').
  \item[GoF] ``Gang of Four'' --- the canonical design patterns after Gamma et
  al.~\cite{gamma1994patterns}.
  \item[HBM] High-Bandwidth Memory.
  \item[HOT] Height Optimized Trie~\cite{binna2018hot}.
  \item[Living being] A complete search algorithm as a composition of all axes (metaphorical code
  term; technical identifier: SearchAlgorithm instance).
  \item[LOUDS] Level-Ordered Unary Degree Sequence --- succinct tree encoding (Jacobson 1989).
  \item[MESI, MOESI, MESIF] Cache coherence protocols (base protocol and AMD/Intel variants,
  respectively).
  \item[NINE] Non-Inclusive-Non-Exclusive --- inclusion policy of the L3 in Intel server processors.
  \item[NUMA] Non-Uniform Memory Access.
  \item[OLC] Optimistic Lock Coupling --- optimistic concurrency technique (ARTSync,
  P08)~\cite{leis2016olc}.
  \item[Organ] Metaphorical term for an \emph{axis} (level~3 of the three-level model); a search
  algorithm has nineteen organs.
  \item[PMC] (Hardware) Performance Monitoring Counter.
  \item[Probe] The new search algorithm to be integrated for testing (here PRT-ART); it is compiled
  by the builder against the core constituents of the cache engine.
  \item[Probe dock] The externally visible measurement interface of a genus, through which the
  builder loads a living being, drives it through the interface operations, measures it, and persists
  the results (cf.\ Section~\ref{sec:anatomy-levels}).
  \item[PRT-ART] Probst Redirect Tree / ART hybrid --- the probe of this thesis.
  \item[Rank] Centrality classification of the state-of-the-art publications (Rank-1/2/3); explained
  in Section~\ref{sec:rqs}.
  \item[RCU] Read-Copy-Update --- reclamation scheme (McKenney, P29)~\cite{mckenney2001rcu}.
  \item[RRIP, DRRIP] (Dynamic) Re-Reference Interval Prediction --- cache replacement
  heuristic~\cite{jaleel2010rrip}.
  \item[SIMD] Single Instruction, Multiple Data.
  \item[SOSD] Search On Sorted Data --- index benchmark~\cite{kipf2019sosd}.
  \item[SOTA] State of the Art.
  \item[SuRF] Succinct Range Filter~\cite{zhang2018surf}.
  \item[SVE, NEON, RVV] ARM (SVE, NEON) and RISC-V (RVV) SIMD extensions, respectively.
  \item[Three-level model] The layering of the one architecture: \emph{genus} (level~1, the outer
  algorithm interface --- SearchAlgorithm, Container, Graph), \emph{genus subclass} (level~2, fixed
  axis set), and \emph{organs}/\emph{axes} (level~3). See Section~\ref{sec:anatomy-levels}.
  \item[TLB] Translation Lookaside Buffer.
  \item[XML load-profile result] The persistence format demanded by the thesis assignment
  (sub-task~7) in which the heuristic policy extracted for an overall algorithm --- reusable per
  architecture focus --- is stored under reference to the applied load profiles; it serves as input
  to the autonomous profile-driven heuristic evaluation. Import parser and export writer including
  the extractor are implemented (round-trip-capable schema \texttt{comdare\_load\_profile}); what
  remains as outlook is the autonomous heuristic evaluation built on top (ML classification).
  \item[YCSB] Yahoo!\ Cloud Serving Benchmark~\cite{cooper2010ycsb}.
\end{description}
