\chapter{Building-Block Matrix}\label{app:blocks}

This appendix contains the complete building-block matrix referenced by the state of the art
(Section~\ref{sec:sota-instances}). Each axis is an \emph{organ} of the living being: a concrete search algorithm (technically \texttt{SearchAlgorithm}, metaphorically a living being) is the fixed composition of its nineteen axes, each of which is an organ; the building-block variants listed per axis are the interchangeable forms of that organ, and their permutation amounts to a genetic experiment on the living being. The \emph{anatomy} of this living being is the wiring \emph{between} the organs: axis algorithms use the interfaces of other axes (the allocation axis, for instance, provides a shared allocation interface for the remaining organs). For each axis it lists the purpose, the \emph{static} sub-axes
(compile-time-selectable variant families) and the \emph{dynamic} sub-axes (run-time parameters), as
well as the concrete building-block variants with their source (paper identifier or code/baseline).
The entries are verified against the \texttt{comdare-cache-engine} code and the paper-provenance map
(Doc~18).

\setlength{\tabcolsep}{3pt}
\section{T0 Search Algorithm Axis T0 (Search Structures and Methods)}

This axis is the \emph{search organ} (T0) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Systematic permutation of 21 concrete search algorithms and structures (S01--S21: 17 base building blocks --- array-based, sorted vectors, original paper wrappers, search methods like k-ary/\allowbreak{}interpolation, and ordered structures like skip-list/\allowbreak{}B-tree --- plus 4 opt-in per-K k-ary wrappers K2/\allowbreak{}K4/\allowbreak{}K8/\allowbreak{}K16, disabled by default) to measurably quantify the impact of search paradigm choice on performance characteristics within a uniform \texttt{std::map} interface. Sub-axes classify by access density (SA1 dense, SA2 sparse, SA3/\allowbreak{}SA4 multilevel), enable compliance proofs through paper mixins (Habich requirement: \texttt{is\_\allowbreak{}original} marking per function/\allowbreak{}wrapper), and identify SIMD-capable and store-traversable variants.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
SA1 & Direct-Addressed High Density (Dense Access) & Array family: direct-addressed (O(1) lookup), high fill density >= 50\%. Includes \texttt{Array256\allowbreak{}Search\allowbreak{}Algo} (256-slot compact array, ART Node256 re-impl) and \texttt{Array65535\allowbreak{}Search\allowbreak{}Algo} (uint16-based, mid-density 25-50\% direct-addressed). \\
SA2 & Sparse/\allowbreak{}Ordered (Sparse Access) & Sorted vectors, ordered search methods, and tree/\allowbreak{}hash structures: low fill density, range-scan support, variable O(log n)--O(1) lookup. Includes \texttt{Vector\allowbreak{}U8\allowbreak{}U8\allowbreak{}Search\allowbreak{}Algo}, original paper wrappers (ART/\allowbreak{}HOT/\allowbreak{}START/\allowbreak{}Wormhole/\allowbreak{}Su\allowbreak{}RF), search methods (k-ary, interpolation, Eytzinger, hash), and structures (skip-list, BST, B-tree). \\
SA3 & Multi-Byte Discriminators (Multilevel Access) & Multi-byte search structures with indexed sub-discriminators (e.g., START with cost-DP multi-level). \texttt{Vector\allowbreak{}U16\allowbreak{}U16\allowbreak{}Search\allowbreak{}Algo} (u16-keys, sorted lower\_\allowbreak{}bound) and \texttt{Original\allowbreak{}Start\allowbreak{}Search\allowbreak{}Algo} (paper P05) belong here. \\
SA4 & Direct-Addressed Medium Density (Direct Multibyte) & Shares direct-addressing access pattern (like SA1) with multi-byte granularity (uint16 size) and medium fill density 25--50\% (vs. SA1 >= 50\%). Example: prt-art Array65535 (REV 6 \S5.17). Sub-axis added to not break SA1 abstraction but capture density variant. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.5cm} >{\raggedright\arraybackslash}p{6.5cm} >{\raggedright\arraybackslash}p{3.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
Array256\allowbreak{}Search\allowbreak{}Algo (S01) & ART Node256 direct-addressed re-implementation: \texttt{std::array<std::optional<u64>, 256>} with O(1) lookup/\allowbreak{}insert/\allowbreak{}erase. Cache-line-aligned, SIMD-capable, no heap allocation. Baseline for dense accesses. & Code/\allowbreak{}Baseline (Leis ICDE 2013 ART Paper P01) \\
Vector\allowbreak{}U8\allowbreak{}U8\allowbreak{}Search\allowbreak{}Algo (S02) & Sparse sorted key-value vector pair: \texttt{std::vector<u8> keys} + \texttt{std::vector<u64> values} with \texttt{std::lower\_\allowbreak{}bound}. O(log n) lookup, O(n) insert/\allowbreak{}erase. Concept-equivalent to HOT-k-constrained. SIMD-capable (bitmask scan). & Code/\allowbreak{}Baseline (Binna SIGMOD 2018 HOT Paper P02) \\
Vector\allowbreak{}U16\allowbreak{}U16\allowbreak{}Search\allowbreak{}Algo (S03) & Sparse sorted u16 vector pair: \texttt{std::vector<u16> keys} + \texttt{std::vector<u64> values}. Simplified START concept (multi-byte discriminators without cost-DP). O(log n) lookup, O(n) insert/\allowbreak{}erase. & Code/\allowbreak{}Baseline (Fent ICDEW 2020 START Paper P05) \\
Array65535\allowbreak{}Search\allowbreak{}Algo (S09) & Mid-density direct-addressed uint16-based (25--50\% fill density). Migrated from prt-art legacy code (REV 6 \S5.17). O(1) lookup but with memory overhead vs. SA1 denser arrays. & Code/\allowbreak{}Baseline (prt-art F.6-Migration) \\
Original\allowbreak{}Art\allowbreak{}Search\allowbreak{}Algo (S04) & ART paper binding (\texttt{unodb::db}): \texttt{is\_\allowbreak{}original\_\allowbreak{}module()=true} (4/\allowbreak{}4 functions original). Paper mixin via SHA256-validated tool code generation. Habich compliance: \texttt{get\_\allowbreak{}compiler()='gcc-9.5'}, all insert/\allowbreak{}lookup/\allowbreak{}erase/\allowbreak{}clear marked original. & P01 (Leis/\allowbreak{}Kemper/\allowbreak{}Neumann ICDE 2013, Apache-2.0 unodb) \\
Original\allowbreak{}Hot\allowbreak{}Search\allowbreak{}Algo (S05) & HOT paper binding (HOTRowex): \texttt{is\_\allowbreak{}original\_\allowbreak{}module()=false} (2/\allowbreak{}4 original)\@. insert/\allowbreak{}lookup original, but erase/\allowbreak{}clear gaps (Hot\allowbreak{}Rowex is append-only; re-impl fallback). Paper P02. & P02 (Binna et al. SIGMOD 2018, ISC ISC) \\
Original\allowbreak{}Start\allowbreak{}Search\allowbreak{}Algo (S06) & START paper binding (sosd-adapter): \texttt{is\_\allowbreak{}original\_\allowbreak{}module()=false} (2/\allowbreak{}4 original)\@. insert/\allowbreak{}lookup original, erase/\allowbreak{}clear gaps (no remove in sosd-adapter). Paper P05, multi-byte discriminators. & P05 (Fent/\allowbreak{}Jungmair/\allowbreak{}Kipf/\allowbreak{}Neumann ICDEW 2020, MIT) \\
Original\allowbreak{}Wormhole\allowbreak{}Search\allowbreak{}Algo (S07) & Wormhole paper binding: \texttt{is\_\allowbreak{}original\_\allowbreak{}module()=false} (3/\allowbreak{}4 original)\@. put/\allowbreak{}get/\allowbreak{}del original, clear gap. Hybrid hash+trie+B+ index. GPL-3.0 licensed (blocks original code linking for thesis). & P07 (Wu/\allowbreak{}Ni/\allowbreak{}Jiang Euro\allowbreak{}Sys 2019, GPL-3.0) \\
Original\allowbreak{}Surf\allowbreak{}Search\allowbreak{}Algo (S08) & Su\allowbreak{}RF paper binding (FST with LOUDS bitmap): \texttt{is\_\allowbreak{}original\_\allowbreak{}module()=false} (1/\allowbreak{}4 original, lookup only). Succinct range filter, multi-header library. Paper P10. & P10 (Zhang/\allowbreak{}Lim/\allowbreak{}Andersen SIGMOD 2018, Apache-2.0) \\
KAry\allowbreak{}Search\allowbreak{}Algo (S10) & k-ary search method (Schlegel/\allowbreak{}Gemulla/\allowbreak{}Lehner Da\allowbreak{}Mo\allowbreak{}N 2009): k-way partition vs.\ binary halving. K is iterable\_\allowbreak{}aspect (K $\in$ \{2,4,8,16\}). O($\lceil$log\_\allowbreak{}\{K+1\} n$\rceil$) iterations, parallel ILP comparisons. Array family but no dedicated traversal organ (SA2 classification). C++23 re-impl, \texttt{is\_\allowbreak{}original=false}. & Schlegel et al.\ Da\allowbreak{}Mo\allowbreak{}N 2009~\cite{schlegel2009kary} \\
Interpolation\allowbreak{}Search\allowbreak{}Algo (S11) & Interpolation search method (Perl/\allowbreak{}Itai/\allowbreak{}Avni CACM 1978): distribution-aware position estimation vs.\ blind halving. O(log log N) avg with uniform distribution, O(N) worst-case. Not SIMD-vectorizable (data-dependent index). C++23 re-impl. & Perl/\allowbreak{}Itai/\allowbreak{}Avni CACM 1978~\cite{perl1978interpolation} \\
Eytzinger\allowbreak{}Search\allowbreak{}Algo (S12) & Eytzinger/\allowbreak{}BFS-layout branch-free search (Khuong/\allowbreak{}Morin JEA 2017): cache-layout search method with branch-free kernel. Keys in BFS order of implicit search tree. For large arrays per paper fastest general-purpose layout. C++23 re-impl (experiment harness without standard license). & Khuong/\allowbreak{}Morin JEA 2017~\cite{khuong2017arraylayouts} \\
Skip\allowbreak{}List\allowbreak{}Search\allowbreak{}Algo (S13) & Skip-list probabilistic ordered STRUCTURE (Pugh CACM 1990): O(log n) expected search/\allowbreak{}insert/\allowbreak{}erase without explicit rebalancing. Each node has 1..k\allowbreak{}Max\allowbreak{}Level forward pointers (coin-flip probability 0.5). Index-based (\texttt{std::vector<Node>} with uint32 indices, value-semantic). Textbook algorithm, C++23 re-impl. & Pugh CACM 1990~\cite{pugh1990skiplist} \\
Hash\allowbreak{}Search\allowbreak{}Algo (S14) & Open-addressing hash table UNORDERED (Knuth TAOCP 3 \S6.4): Fibonacci hash, linear probing, tombstone-erase on resize. O(1) expected lookup/\allowbreak{}insert at load\_\allowbreak{}factor < 0.7. No range-scan support (distinctive classification). Correct tombstone marking for probe-chain integrity. C++23 re-impl. & Knuth TAOCP 3 \S6.4~\cite{knuth1998taocp3} \\
Linear\allowbreak{}Scan\allowbreak{}Search\allowbreak{}Algo (S15) & Unsorted linear scan baseline (Leis ICDE 2013 ART Node4): O(n) lookup on unsorted KV-array. Simplest comparison null-point structure: insert O(1)-amortized, erase O(1) via swap-and-pop. Store-traversable (array family). No range-scan support. & Code/\allowbreak{}Baseline (Leis ICDE 2013 ART Paper P01 Node4) \\
Binary\allowbreak{}Search\allowbreak{}Tree\allowbreak{}Search\allowbreak{}Algo (S16) & Unbalanced binary search tree (Knuth TAOCP 3 \S6.2.2 Hibbard-deletion): deterministic comparison-tree baseline (vs.\ skip-list probabilistic). Index-based (\texttt{std::vector<Node>} with uint32 indices, free-list). Hibbard-erase (3 cases: leaf/\allowbreak{}1 child/\allowbreak{}2 children via in-order successor). O(n) worst-case with sorted input. & Knuth TAOCP 3 \S6.2.2~\cite{knuth1998taocp3} \\
BTree\allowbreak{}Search\allowbreak{}Algo (S17) & Balanced block-oriented multi-way B-tree (Bayer/\allowbreak{}Mc\allowbreak{}Creight Acta Inf. 1972 /\allowbreak{} CLRS Ch. 18): t=4, so [3..7] keys per node, up to 8 children. ALWAYS balanced (O(log n) guaranteed). Index-based, \texttt{alignas(64)} nodes (cache-line alignment). Top-down splitting on insert, borrow/\allowbreak{}merge on erase. DB index workhorse. & Bayer/\allowbreak{}Mc\allowbreak{}Creight Acta Inf.\ 1972 /\allowbreak{} CLRS~\cite{bayer1972btree} \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} P01, P02, P05, P07, P10, k-ary (Schlegel et al.\ 2009), interpolation (Perl et al.\ 1978), Eytzinger (Khuong/\allowbreak{}Morin 2017), skip list (Pugh 1990), hashing/\allowbreak{}BST (Knuth TAo\allowbreak{}CP 3), B-tree (Bayer/\allowbreak{}Mc\allowbreak{}Creight 1972)~\cite{leis2013art,binna2018hot,fent2020start,wu2019wormhole,zhang2018surf,schlegel2009kary,perl1978interpolation,khuong2017arraylayouts,pugh1990skiplist,knuth1998taocp3,bayer1972btree}.

\section{T1 Cache Traversal (Fanout Lookup)}

This axis is the \emph{cache-traversal organ} (T1) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Axis \texttt{axis\_\allowbreak{}03b} defines three distinct strategies for key resolution in small fanout arrays (typically B+ tree or trie inner nodes). It spans a spectrum from linear scan (O(N), cache-friendly, small), via multiplicative hashing with open addressing (O(1) amortized, Knuth-Fibonacci), to sorted binary search (O(log N), ordered traversal, B+ tree standard). All three are textbook baselines without original code linking, implemented as pure C++23 re-implementations.

\subsection*{Static sub-axes}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
CT1 & Linear Access & Linear-scan-based lookup methods on small, unsorted fanout arrays. Characterized by O(N) resolution with simple comparisons, typical for nodes with fanout $<$ 32 in B+ trees. Sub-axis for \texttt{Linear\allowbreak{}Fanout} and \texttt{Binary\allowbreak{}Search\allowbreak{}Fanout} (both: \texttt{axis\_\allowbreak{}tag} = \texttt{linear\_\allowbreak{}access\_\allowbreak{}tag}). \\
CT2 & Hash-Based Access & Hash lookup with open addressing and linear probing. O(1) amortized resolution, characterized by Fibonacci multiplier (11400714819323198485 $\approx$ $2^{64}/\allowbreak{}\varphi$), power-of-2 bucket count, dynamic rehashing when load factor $>$ 0.7. Only \texttt{Hash\allowbreak{}Lookup} wrapper. Sub-axis: \texttt{axis\_\allowbreak{}tag} = \texttt{hash\_\allowbreak{}access\_\allowbreak{}tag}. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{2.5cm} >{\raggedright\arraybackslash}p{6.5cm} >{\raggedright\arraybackslash}p{4.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Linear\allowbreak{}Fanout} & Unsorted fanout traversal via \texttt{std::find\_\allowbreak{}if} and linear scan. O(N) lookup, no hash overhead, cache-friendly for small N ($<$ 32). Classic B-tree baseline (Bayer/\allowbreak{}Mc\allowbreak{}Creight 1972). Belongs to sub-axis CT1. & Code/\allowbreak{}Baseline --- Bayer/\allowbreak{}Mc\allowbreak{}Creight 1972 (Acta Informatica) \\
\texttt{Hash\allowbreak{}Lookup} & Fibonacci-hash open addressing with linear probing (Knuth TAOCP Vol. 3, §6.4). O(1) amortized, dynamic rehashing (doubles capacity when load $\geq$ 0.7), iterable runtime capacity (\texttt{k\allowbreak{}Iterable\allowbreak{}Initial\allowbreak{}Capacities}: \{8, 16, 64, 256, 1024\}). Standalone C++23 re-implementation, no original code binding. Belongs to sub-axis CT2. & Code/\allowbreak{}Baseline --- Knuth 1998 (The Art of Computer Programming Vol.3) \\
\texttt{Binary\allowbreak{}Search\allowbreak{}Fanout} & Sorted fanout traversal via \texttt{std::lower\_\allowbreak{}bound}. O(log N) lookup with sorted ordering, preserves key sequencing for range-capable routing layers. Classic B+ tree inner-node search (Bayer/\allowbreak{}Mc\allowbreak{}Creight 1972), third distinct strategy of the axis. C++23 re-implementation (R7.2, 2026-05-29), \texttt{is\_\allowbreak{}original}=false. Belongs to sub-axis CT1. & Code/\allowbreak{}Baseline --- Bayer/\allowbreak{}Mc\allowbreak{}Creight 1972 (Acta Informatica) \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} Bayer/\allowbreak{}Mc\allowbreak{}Creight 1972, Knuth TAo\allowbreak{}CP Vol.\,3~\cite{bayer1972btree,knuth1998taocp3}.

\emph{Related literature:} P17, P21, P32~\cite{bender2005coboblivious,chen2001prefetch,schmidt2025hbm}.

\section{T2 Slot-to-Offset Mapping (Traversal Sublayer)}

This axis is the \emph{mapping organ} (T2) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Mediation between logical slot indices and physical memory offsets in trie inner nodes and memory pools. The axis provides two fundamentally different mapping strategies: direct absolute offsets (linear packing) or pool-relative offsets with automatic rebase mechanism supporting memory reallocation and persistent data structures.

\subsection*{Static sub-axes}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
MP1 & Direct-Access Mapping (MP1) & Direct slot-to-absolute-offset mapping without indirection. Wrapper family for classic array-of-pointers layout as in B-tree inner nodes (Bayer/\allowbreak{}Mc\allowbreak{}Creight 1972). Simple cache locality, O(N) lookup via \texttt{std::find\_\allowbreak{}if}. \\
MP2 & Pool-Relative Mapping (MP2) & Slot-to-pool-relative-offset mapping using based-pointer arena pattern. All offsets stored relative to dynamic \texttt{pool\_\allowbreak{}base\_\allowbreak{}address}. Enables O(1) rebase on memory reallocation --- core property for persistent data structures (Driscoll et al. 1989) and Custom\allowbreak{}Aligned\allowbreak{}Structure pattern in prt-art. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{2.6cm} >{\raggedright\arraybackslash}p{8.5cm} >{\raggedright\arraybackslash}p{2.4cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
Direct\allowbreak{}Placement & Wrapper for direct slot-to-absolute-offset mapping (linear packing). \texttt{std::vector<pair<slot\_\allowbreak{}index\_\allowbreak{}type, offset\_\allowbreak{}type>>}; \texttt{resolve\_\allowbreak{}offset}/\allowbreak{}\texttt{register\_\allowbreak{}slot} via linear \texttt{std::find\_\allowbreak{}if}. Usage: prt-art \texttt{INode::placement\_\allowbreak{}page\_\allowbreak{}} + mapping table. Storage: 16-bit slot index + \texttt{size\_\allowbreak{}t} offset per entry. No pool base dependency. & CE-Baseline \\
Pool\allowbreak{}Relative & Wrapper for pool-relative-offset mapping with O(1) rebase capability. \texttt{std::vector<pair<slot\_\allowbreak{}index\_\allowbreak{}type, relative\_\allowbreak{}offset>>} + mutable \texttt{pool\_\allowbreak{}base\_\allowbreak{}address}. On pool reallocation: \texttt{rebase(new\_\allowbreak{}pool\_\allowbreak{}base)} updates only \texttt{pool\_\allowbreak{}base\_\allowbreak{}} --- all relative offsets remain valid. Used with prt-art Custom\allowbreak{}Aligned\allowbreak{}Structure + versioning scenarios (persistent data structures per Driscoll/\allowbreak{}Sarnak/\allowbreak{}Sleator/\allowbreak{}Tarjan 1989). & CE-Baseline \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} --- (CE baseline, own wrappers).

\emph{Related literature:} Bayer/\allowbreak{}Mc\allowbreak{}Creight 1972, Driscoll et al.\ 1989, P03, P07~\cite{bayer1972btree,driscoll1989persistent,mao2012masstree,wu2019wormhole}.

\section{T3 Path Compression}

This axis is the \emph{path-compression organ} (T3) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Defines the compression granularity and strategy for key paths in trie structures. Controls how inner nodes store and navigate redundant byte sequences or significant bits of common prefixes to optimize memory and cache efficiency.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
pc1 & Compression Granularity (PC1) & Defines the minimal unit of path compression: bit-wise (Patricia), byte-wise (ART), or no compression (raw path). Determines trie height and optimal code blocks for traversal. \\
pc2 & Skip Strategy (PC2) & Defines the navigation pattern at inner nodes: linear over 8 bytes (Byte\allowbreak{}Wise), single-bit comparison (Patricia), or direct jump without intermediate steps. Influences branch density and prefetch behavior. \\
pc3 & Decode Complexity (PC3) & Classifies the runtime complexity when extracting prefix information: O(1) for direct extraction, O(log n) for structure-based search, O(n) for sequential check. Important for operations like \texttt{common\_\allowbreak{}prefix\_\allowbreak{}len()} and key comparisons. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{8.5cm} >{\raggedright\arraybackslash}p{2.0cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Path\allowbreak{}Compression\allowbreak{}None} & No path compression; raw key paths are stored completely and uncompressed in nodes. Serves as measurement baseline with \texttt{compression\_\allowbreak{}ratio=1.0}. Used as control case to quantify memory and height savings of compressing strategies. & Code/\allowbreak{}Baseline \\
\texttt{Byte\allowbreak{}Wise\allowbreak{}Path\allowbreak{}Compression} & Byte-by-byte compression (ART-style); stores up to 7 prefix bytes per node in a compact \texttt{u64} word with length tag in high byte. Typical 2x reduction in trie height (\texttt{compression\_\allowbreak{}ratio=0.5}). Origin: Leis et al., ART ICDE 2013 (P01). Employs goldstandard operations \texttt{common\_\allowbreak{}prefix\_\allowbreak{}len()} and \texttt{cut()} for navigation optimization. & P01 \\
\texttt{Patricia\allowbreak{}Path\allowbreak{}Compression} & Single-bit-split path compression (Patricia trie); navigates bit-wise instead of byte-wise, reduces trie height by typically factor 3 (\texttt{compression\_\allowbreak{}ratio=0.3}). Origin: Morrison PATRICIA JACM 1968; modernly implemented in HOT (Binna SIGMOD 2018, P02) and Wormhole (Wu Euro\allowbreak{}Sys 2019, P07). F15-operative primitives: \texttt{key\_\allowbreak{}split\_\allowbreak{}bit()} for MSB-first bit extraction, \texttt{path\_\allowbreak{}descend\_\allowbreak{}scan()} for goldstandard-compliant \texttt{\_\allowbreak{}scan} signature with key-dependent variable depth. & P02 \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} P01, P02, P07, Morrison 1968~\cite{leis2013art,binna2018hot,wu2019wormhole,morrison1968patricia}.

\emph{Related literature:} P04, P05, P09~\cite{boffa2024coco,fent2020start,jacobson1989louds}.

\section{T4 Node Types (Node-Type Axis)}

This axis is the \emph{node-type organ} (T4) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

The axis \texttt{axis\_\allowbreak{}04\_\allowbreak{}node\_\allowbreak{}type} defines the four node formats of the Adaptive Radix Tree (ART) as exchangeable variants: Node4, Node16, Node48, and Node256. Each node type offers a different balance between memory, cache efficiency, and lookup speed. This axis forms the foundation for adaptive tree reorganization and is central to the design of search algorithms.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
NT1 & Capacity Class (NT1) & Classification by fanout size and slot count: small (Node4 with 4 slots, linear search), medium (Node16 with 16 slots, SIMD binary search) to maximal (Node256 with 256 slots, direct-addressed). \\
NT2 & Access Pattern (NT2) & Method of key lookup within a node: linear scan (Node4), SIMD binary search (Node16), indirect indexing (Node48), or direct addressing (Node256). \\
NT3 & Compactness (NT3) & Memory layout and density: sparse and cache-aligned (Node4), dense with indirection (Node48), or maximally direct (Node256), with impacts on memory usage and cache-line misses. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{8.5cm} >{\raggedright\arraybackslash}p{2.0cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Node4\allowbreak{}Node\allowbreak{}Type} & Small ART node with 4 slots and linear search. Ideal for sparsely populated upper tree levels, fits completely in a cache line (64 bytes). Implemented as CE re-implementation of the ART specification with \texttt{node\_\allowbreak{}find\_\allowbreak{}scan()} as linear scan over sorted keys. & P01 \\
\texttt{Node16\allowbreak{}Node\allowbreak{}Type} & Medium ART node with 16 slots, optimized for SIMD-based binary search (SSE2 PCMPESTRM /\allowbreak{} AVX2 VPCMPESTRM). Balance between memory and search speed; largest capacity without indirection. CE re-implementation of the Leis ICDE 2013 specification. & P01 \\
\texttt{Node48\allowbreak{}Node\allowbreak{}Type} & Large ART node with 48 slots, using indirection: 256-byte Child\allowbreak{}Index (byte→slot mapping, 0=empty) plus 48-slot Child\allowbreak{}Array. Trade-off between memory (vs Node256 direct) and cache misses (vs Node4 linear). Double-indirection access pattern (index + child). & P01 \\
\texttt{Node256\allowbreak{}Node\allowbreak{}Type} & Maximal ART node with 256 slots, directly addressed without index. O(1) lookup (no scan, no index touch). High memory overhead (256 pointers), but optimal search speed for dense nodes. Default for lower tree levels. & P01 \\
Internal/\allowbreak{}Border & Node-type pair of Masstree (trie of B\textsuperscript{+} trees): internal nodes navigate the 8-byte key slices of the B\textsuperscript{+} level, border nodes form the leaf level with the values or the transition to the next trie layer. Paper node format (reconstruction target of this thesis, no CE baseline). & P03 (Mao et al.\ 2012)~\cite{mao2012masstree} \\
Decision/\allowbreak{}Span & Node-type pair of the B\textsuperscript{2}-tree (B\textsuperscript{+} with embedded binary inner nodes, cache-aware): decision nodes carry the binary branching decision within the page, span nodes group contiguous key ranges. Paper node format (reconstruction target of this thesis, no CE baseline). & P06 (Schmeisser et al.\ 2022)~\cite{schmeisser2022b2tree} \\
LOUDS-Dense/\allowbreak{}Sparse & Node-type pair of the succinct range filter Su\allowbreak{}RF\@: LOUDS-Dense encodes the upper, densely populated trie levels bitmap-based with direct addressing, LOUDS-Sparse the lower, sparsely populated levels byte-sequentially succinct. Paper node format (reconstruction target of this thesis, no CE baseline). & P10 (Zhang et al.\ 2018)~\cite{zhang2018surf} \\
\texttt{Node\allowbreak{}Type\allowbreak{}Slot\allowbreak{}Store} & Storage organ: node-type-driven bounded slot storage. Spans \texttt{std::array} with capacity \texttt{N::max\_\allowbreak{}capacity()} over a shared uint64 key. Fulfills the Storage\allowbreak{}Organ concept and is drop-in substrate for Composed\allowbreak{}Search. Baseline variant without layout/\allowbreak{}allocator integration. & Code/\allowbreak{}Baseline \\
\texttt{Composed\allowbreak{}Store} & 3-axis storage organ: node\_\allowbreak{}type(N) $\oplus$ layout(L) $\oplus$ allocator(A). Unbounded vector over allocator A\@; layout flows in as static cache\_\allowbreak{}line\_\allowbreak{}size and via optional \texttt{organ\_\allowbreak{}scan\_\allowbreak{}field\_\allowbreak{}sum()}. Really measurable (F15 bridge). & Code/\allowbreak{}Baseline \\
\texttt{Node\allowbreak{}Chunked\allowbreak{}Store} & 3-axis storage organ with runtime impact of node\_\allowbreak{}type axis: Slots lie in chunks of capacity \texttt{N::max\_\allowbreak{}capacity()}. Chunk count = number of node-sized allocations; thus the node\_\allowbreak{}type axis is REALLY measurable (Node4 many chunks, Node256 few). Proof of delegation. & Code/\allowbreak{}Baseline \\
\texttt{Layout\allowbreak{}Aware\allowbreak{}Chunked\allowbreak{}Store} & Layout-honoring variant of Node\allowbreak{}Chunked\allowbreak{}Store: stores records at layout-driven stride (\texttt{eff\_\allowbreak{}stride}). Each record occupies \texttt{eff\_\allowbreak{}stride} bytes (key 0..8, value 8..16, rest padding). Makes memory\_\allowbreak{}layout axis really measurable, OOB-safe for scan methods. & Code/\allowbreak{}Baseline \\
\texttt{Observable\allowbreak{}Node\allowbreak{}Type} & Observable\allowbreak{}Axis wrapper around a node\_\allowbreak{}type strategy. Passes through static API (max\_\allowbreak{}capacity/\allowbreak{}name/\allowbreak{}topic\_\allowbreak{}tag) + offers \texttt{observe\_\allowbreak{}node\_\allowbreak{}find()} as instance driver. Tracks find\_\allowbreak{}count, keys\_\allowbreak{}stored, queries\_\allowbreak{}run, last\_\allowbreak{}checksum. Format-divergent lookup latency measurable (path B). & Code/\allowbreak{}Baseline \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} P01~\cite{leis2013art}, P03~\cite{mao2012masstree}, P06~\cite{schmeisser2022b2tree}, P10~\cite{zhang2018surf}.

\section{T5 Memory Layout}

This axis is the \emph{memory-layout organ} (T5) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

This axis classifies the spatial organization of records in memory, directly impacting cache efficiency, SIMD vectorization potential, and false-sharing characteristics. It captures four orthogonal design dimensions: (1) alignment strategies (cache-line padding vs.\ strict packing), (2) data organization (row-oriented Ao\allowbreak{}S vs.\ column-oriented So\allowbreak{}A vs.\ hybrid Ao\allowbreak{}So\allowbreak{}A), (3) packing density (dense vs.\ succinct bit-packed), (4) stride patterns (sequential vs.\ interleaved). Layout choice is a primary driver of cache locality and SIMD capability during index traversal.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
HM1 & Alignment Strategy & Choice between cache-line-padded layout (false-sharing avoidance) and strict packing (memory density). Determines stride overhead during sequential traversal. \\
HM2 & Data Organization & Primary memory ordering: Array-of-Structs (Ao\allowbreak{}S, row-oriented), Struct-of-Arrays (So\allowbreak{}A, column-oriented), or hybrid Ao\allowbreak{}So\allowbreak{}A (block-mixed). Critical for cache-line utilization during field-selective scans. \\
HM3 & Packing Density & Storage density level: from standard records (8--64 bytes) to compact bit-packing (e.g., LOUDS bitmaps with \textasciitilde{}1 bit per slot). Affects decode latency and L1/\allowbreak{}L2 utilization. \\
HM4 & Stride Pattern & Addressing pattern during traversal: sequential strides (consecutive records), interleaved (e.g., to separate hot/\allowbreak{}cold data), or cache-aware layouts (e.g., Eytzinger BFS). Optimizes prefetch efficiency. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Dynamic sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Parameter & Description \\
\midrule
\endhead%
Cache\allowbreak{}Line\allowbreak{}Size & Cache-Line Size (Subaxis KF-5) & Runtime parameter for target platform cache-line size: 64 bytes (standard x86-64/\allowbreak{}ARM64), 128 bytes (Azure, some Power ISAs), or 32 bytes (embedded). Selected at compile-time via \texttt{Cache\allowbreak{}Line\allowbreak{}Config} NTTP and affects alignment/\allowbreak{}padding overhead for all aligned layouts. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{7.0cm} >{\raggedright\arraybackslash}p{3.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Ao\allowbreak{}SStrict\allowbreak{}Memory\allowbreak{}Layout} & Array-of-Structs, tightly packed without cache-line alignment. Advantage: minimal memory footprint (record\_\allowbreak{}size strides, dense). Disadvantage: false-sharing under concurrent writes, poor cache-hit ratio for narrow field scans. Typical for Wormhole index. & Code/\allowbreak{}Baseline \\
\texttt{Cache\allowbreak{}Line\allowbreak{}Aligned\allowbreak{}Memory\allowbreak{}Layout} & Ao\allowbreak{}S with 64-byte per-record alignment. Avoids false-sharing via padding (stride = round\_\allowbreak{}up(record\_\allowbreak{}size, 64)). Standard for ART/\allowbreak{}HOT/\allowbreak{}Masstree/\allowbreak{}START\@. Operationally: at record\_\allowbreak{}size=48 (bench build), creates 16-byte overhead per record $\rightarrow$ measurably higher latency than aos\_\allowbreak{}strict, genuine cache effect. & Code/\allowbreak{}Baseline \\
\texttt{So\allowbreak{}AMemory\allowbreak{}Layout} & Struct-of-Arrays: all fields stored column-wise sequentially. Advantage: contiguous field scans (e.g., keys-only without values) maximize cache-line utilization (\textasciitilde{}16x fewer cache-lines than Ao\allowbreak{}S for single-field scan). Ideal for SIMD vectorization and OLAP indexes. Benefit described in Abadi et al. SIGMOD 2008. & Abadi et al.\ 2008~\cite{abadi2008columnrow} \\
\texttt{Packed\allowbreak{}Bitmap\allowbreak{}Memory\allowbreak{}Layout} & Bit-packed succinct layout (LOUDS /\allowbreak{} Su\allowbreak{}RF FST). Stores n entries in \textasciitilde{}n*lg(sigma) bits instead of bytes. Extreme compactness (32x fewer cache-lines than Ao\allowbreak{}S), but higher decode cost per lookup. Foundation: Jacobson LOUDS FOCS 1989 (P09). Modern impl: Su\allowbreak{}RF Apache-2.0. & P09 (Jacobson LOUDS FOCS 1989) \\
\texttt{Ao\allowbreak{}So\allowbreak{}AMemory\allowbreak{}Layout} & Array-of-Structures-of-Arrays: hybrid layout with block width W=8 (AVX2 lanes). Within block: So\allowbreak{}A (fields contiguous). Across blocks: Ao\allowbreak{}S (blocks strided). Unites SIMD vectorization (So\allowbreak{}A within) with spatial locality (Ao\allowbreak{}S across). Standard in HPC/\allowbreak{}SIMD engines. Convention; no single originating paper, but ISPC idiom. & Code/\allowbreak{}Baseline (HPC/\allowbreak{}SIMD convention) \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} P09, Abadi et al.\ 2008~\cite{jacobson1989louds,abadi2008columnrow}.

\emph{Related literature:} P06, P11, P12, P32~\cite{schmeisser2022b2tree,rao1999css,rao2000csb,schmidt2025hbm}.

\section{T6 Allocator Axis (T6)}

This axis is the \emph{allocation organ} (T6) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

This axis provides memory management strategy for the cache engine, defining 25 allocator wrappers covering a spectrum from classical baselines (Buddy, Slab) through modern research allocators (mimalloc, jemalloc, snmalloc) to specialized variants for NUMA, PIM, and formally verified systems. All wrappers satisfy standardized concepts and are orthogonally structured by 7 compile-time variant families (freelist topology, size-class schema, thread locality, synchronization, allocation policy, reclamation, fragmentation strategy).

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
AA1 & Free-List Topology & How are free memory blocks organized? Examples: per-page sharding (mimalloc A04), per-thread caching (snmalloc A07), buddy trees (A19), single-global free-list (dlmalloc A20), spans (scalloc A08), hybrid strategies (hmalloc A15). \\
AA2 & Size-Class Schema & How are block sizes classified? Examples: slab per object size (A02), dlmalloc bins (A20), jemalloc size-classes (A05), scalloc spans (A08), buddy power-of-2 (A19), formally verified bins (Star\allowbreak{}Malloc A13), std::pmr pools (A22). \\
AA3 & Thread Locality & How thread/\allowbreak{}core-local is allocation? Examples: tcmalloc thread-local caches (A06), rpmalloc thread-heaps (A10), snmalloc message-passing (A07), tcmalloc-warehouse hyperscale (A14), Hoard per-heap locks (A01). \\
AA4 & Synchronization & What concurrency mechanism? Examples: lock-free CAS (Michael A03, LRMalloc A11), per-heap locks (Hoard A01), single-threaded (Exgen A18), wait-free (Crystalline A17). \\
AA5 & Allocation Policy & How are allocation requests routed? Examples: NUMA-origin awareness (NUMAlloc A09), cache-set awareness (CAMA A12), NUCA awareness (TCMalloc-Warehouse A14), PIM awareness (PIM-malloc A16), polymorphic memory resource (std::pmr A22). \\
AA6 & Reclamation & How are deleted blocks reclaimed? Examples: magazine cache (Slab A02), page-heap release (TCMalloc A06), lock-free reclamation (Crystalline A17), deferred-free (mimalloc A04), Vmem magazine extension (A23). \\
AA7 & Fragmentation Strategy & How is fragmentation prevented? Examples: dlmalloc coalescing (A20), object coloring (Slab A02), buddy splitting (A19), page-local sharding (mimalloc A04), span pooling (scalloc A08). \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{4.0cm} >{\raggedright\arraybackslash}p{6.5cm} >{\raggedright\arraybackslash}p{3.0cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Hoard\allowbreak{}Allocator} (A01) & Per-heap free-lists with SMP scalability. Berger et al. (PPo\allowbreak{}PP 2000). Synchronization via per-heap locks and thread-local heap caches. & P— (OSS, Apache-2.0) \\
\texttt{Slab\allowbreak{}Allocator} (A02) & Object-caching kernel allocator with magazine-cache strategy. Bonwick (USENIX 1994, 2001). Classic textbook paradigm without standalone OSS implementation. & P— (Baseline/\allowbreak{}Textbook) \\
\texttt{Michael\allowbreak{}Lock\allowbreak{}Free\allowbreak{}Allocator} (A03) & Lock-free CAS allocator with hazard pointers. Maged Michael (PLDI 2004). Re-implementation with IBM patent background. & P— (MIT Re-Impl) \\
\texttt{Mimalloc\allowbreak{}Allocator} (A04) & Free-list sharding with deferred-free and page-local management. Leijen/\allowbreak{}Zorn/\allowbreak{}de Moura (MSR-TR-2019-18, APLAS 2019). Original code binding (is\_\allowbreak{}original=2/\allowbreak{}2). & P— (OSS, MIT, is\_\allowbreak{}original) \\
\texttt{Jemalloc\allowbreak{}Allocator} (A05) & Arena-based size-classes with per-CPU arenas and tcache layer. Evans (BSDCan 2006). Standard allocator of Free\allowbreak{}BSD and Facebook. Original code binding (is\_\allowbreak{}original=2/\allowbreak{}2). & P— (OSS, BSD-2-Clause, is\_\allowbreak{}original) \\
\texttt{TCMalloc\allowbreak{}Allocator} (A06) & Thread-caching malloc with central page-heap. Ghemawat/\allowbreak{}Menage (Google ca. 2005) + TEMERAIRE (OSDI 2021). Google production standard. & P— (OSS, Apache-2.0) \\
\texttt{Snmalloc\allowbreak{}Allocator} (A07) & Message-passing allocator with asymmetric heap operations. Paul Liétar et al. (ISMM 2019). Original code binding (is\_\allowbreak{}original=2/\allowbreak{}2). & P— (OSS, MIT, is\_\allowbreak{}original) \\
\texttt{Scalloc\allowbreak{}Allocator} (A08) & Span-based free-list with virtual spans for multi-core scalability. Aigner/\allowbreak{}Kirsch/\allowbreak{}Lippautz/\allowbreak{}Sokolova (OOPSLA 2015). Low fragmentation via span pooling. & P— (OSS, BSD-3-Clause) \\
\texttt{NUMAlloc\allowbreak{}Allocator} (A09) & NUMA-origin-aware allocator for faster allocation in multi-socket systems. UTSASRG (ISMM 2023). Optimized for multi-socket architectures. & P— (OSS, vermutlich Apache-2.0) \\
\texttt{RPMalloc\allowbreak{}Allocator} (A10) & Per-thread span caching without dependencies on other allocator libraries. Mattias Jansson (2017). Original code binding (is\_\allowbreak{}original=2/\allowbreak{}2). & P— (OSS, Public-Domain/\allowbreak{}MIT, is\_\allowbreak{}original) \\
\texttt{LRMalloc\allowbreak{}Allocator} (A11) & Lock-free allocator with hazard pointers for multi-core scalability. Leite/\allowbreak{}Rocha (VECPAR 2018, LNCS 11333). Original code binding (is\_\allowbreak{}original=2/\allowbreak{}2). & P— (OSS, MIT, is\_\allowbreak{}original) \\
\texttt{CAMAAllocator} (A12) & Cache-Aware Memory Allocator with predictable behavior. Herter/\allowbreak{}Backes/\allowbreak{}Haupenthal/\allowbreak{}Reineke (ECRTS 2011, Saarland). No standalone OSS implementation available. & P— (Baseline/\allowbreak{}Research) \\
\texttt{Star\allowbreak{}Malloc\allowbreak{}Allocator} (A13) & Formally verified hardened memory allocator with F*/\allowbreak{}Steel. Inria-Prosecco: Reitz/\allowbreak{}Fromherz/\allowbreak{}Protzenko et al. (OOPSLA 2024). Mathematically verified allocator. & P— (OSS, Apache-2.0) \\
\texttt{TCMalloc\allowbreak{}Warehouse\allowbreak{}Allocator} (A14) & TCMalloc hyperscale extension with hugepage support for warehouse-scale computing. TEMERAIRE (Hunter OSDI 2021) + warehouse characterization (Zhou ASPLOS 2024). NUCA-aware TCMalloc variant. & P— (OSS, Apache-2.0) \\
\texttt{HMalloc\allowbreak{}Allocator} (A15) & Hybrid lock-free allocator with scalable memory management. Li/\allowbreak{}Yao/\allowbreak{}Tang/\allowbreak{}Lin (IEEE ICPADS 2019). No standalone OSS implementation available. & P— (Research Baseline) \\
\texttt{PIMMalloc\allowbreak{}Allocator} (A16) & Processing-in-memory allocator for heterogeneous hardware with fast local memory. VIA-Research (HPCA 2026, ar\allowbreak{}Xiv:2505.13002). Specialized for PIM hardware. & P— (OSS, MIT) \\
\texttt{Crystalline\allowbreak{}Reclamation} (A17) & Wait-free safe-memory-reclamation family (Nikolaev/\allowbreak{}Ravindran DISC 2021 / PLDI 2024). Reclamation policy, not a standalone allocator. & P— (Research) \\
\texttt{Exgen\allowbreak{}Allocator} (A18) & Single-threaded specialized allocator with optimizations for single-threaded workloads. UT Austin (IEEE CAL 2025, ar\allowbreak{}Xiv:2510.10219)\@. \enquote{Old is Gold} principle. & P— (Research, Konfidenz medium) \\
\texttt{Buddy\allowbreak{}Allocator} (A19) & Buddy system with power-of-2 splitting for simple fragmentation control. Knuth (TAo\allowbreak{}CP Vol.1, 1968). Classic textbook paradigm. & P— (Baseline/\allowbreak{}Textbook) \\
\texttt{Dlmalloc\allowbreak{}Allocator} (A20) & Doug Lea malloc with bins-based classification. Lea (1987--2012). Original code binding (is\_\allowbreak{}original=2/\allowbreak{}2). & P— (OSS, Public-Domain/\allowbreak{}CC0, is\_\allowbreak{}original) \\
\texttt{Pt\allowbreak{}Malloc2\allowbreak{}Allocator} (A21) & ptmalloc2 (glibc malloc) with explicit wrapper control. Wolfram Gloger /\allowbreak{} Ulrich Drepper (1990s-2000s). License-blocked (LGPL-2.1+). & P— (OSS, LGPL-2.1+, blockiert Linking) \\
\texttt{Std\allowbreak{}Malloc} (A22a) & Standard libc malloc (proof-of-concept) — portable fallback allocator. Baseline for all standard library implementations. & P— (Standard-Library) \\
\texttt{Pmr\allowbreak{}Resource\allowbreak{}Allocator} (A22b) & std::pmr::memory\_\allowbreak{}resource (polymorphic memory resource). Halpern (N3916, C++17). Abstract interface without behaviorally distinct default implementation. & P— (Standard-Library, C++17) \\
\texttt{Pool\allowbreak{}Resource\allowbreak{}Allocator} (A22c) & std::pmr::unsynchronized\_\allowbreak{}pool\_\allowbreak{}resource with own size-class pool. Halpern (N3916, C++17). Behaviorally distinct without vendor linking (F15-operational). & P— (Standard-Library, C++17, F15-operativ) \\
\texttt{Vmem\allowbreak{}Magazines\allowbreak{}Allocator} (A23) & Vmem plus magazines as extension of slab allocator for multi-core scaling. Bonwick (USENIX ATC 2001). No standalone OSS implementation available. & P— (Baseline/\allowbreak{}Textbook) \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing reclamation sources:} P29, P30~\cite{mckenney2001rcu,michael2004hazard}; full allocator corpus A01--A23 see section \enquote{A-Corpus}.

\section{T7 Prefetch Strategy}

This axis is the \emph{prefetch organ} (T7) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

The Prefetch axis determines the strategy for anticipatory loading of nodes and addresses along the expected search path in a trie data structure. It supports four concrete strategies with different trigger mechanisms and heuristics: no prefetch (measurement baseline), distance-estimated software prefetch based on node density and cache latency (ART standard), explicit hardware prefetch instructions (Wormhole standard), and path-oriented prefetch with bundle granularity (PRT-ART thesis).

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
PF1 & Trigger Mechanism & Type of control over the prefetch timing: whether no mechanism (baseline), through compiler hints via distance estimator, explicit hardware prefetch instructions, or path-oriented prediction \\
PF2 & Distance Heuristic & Calculation method for prefetch distance (how far ahead to load): linear, adaptive based on node density and tier latency, or no heuristic \\
PF3 & Granularity & Prefetch size per trigger: cache-line (64 bytes, individual), page (4\allowbreak{}K, OS window), or bundle (multiple cache-lines, trie-path-oriented) \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{7.0cm} >{\raggedright\arraybackslash}p{3.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{None\allowbreak{}Prefetch} & No prefetch strategy (baseline). Serves as comparison reference for measurement series. Trigger: none; heuristic: none; granularity: N/\allowbreak{}A. \texttt{is\_\allowbreak{}active()} returns false. & Code/\allowbreak{}Baseline \\
\texttt{Distance\allowbreak{}Estimator\allowbreak{}Prefetch} & Distance-estimated software prefetch (ART standard per Leis ICDE 2013). Estimates cache distance to next node based on node density [\%] and tier latency [cycles] with heuristic: estimate = clamp(1 + sparseness*8 + latency\_\allowbreak{}factor, [1,16]). Trigger: compiler hints via \texttt{\_\allowbreak{}\_\allowbreak{}builtin\_\allowbreak{}prefetch}; heuristic: adaptive (density + latency); granularity: cache-line. Applies ART pattern (unodb/\allowbreak{}libart). & P01 (ART/\allowbreak{}unodb, Leis ICDE 2013) \\
\texttt{Hardware\allowbreak{}Prefetch} & Explicit hardware prefetch instructions (Wormhole standard per Wu Euro\allowbreak{}Sys 2019). Uses PREFETCHT0/\allowbreak{}T1/\allowbreak{}T2/\allowbreak{}NTA on hash-anchor-chain without software heuristic; CPU estimates distance autonomously. Trigger: explicit hardware instruction; heuristic: hardware-managed; granularity: hardware-determined. License GPL-3.0 blocks \texttt{is\_\allowbreak{}original}-linking. & P07 (Wormhole, Wu Euro\allowbreak{}Sys 2019) \\
\texttt{Path\allowbreak{}Oriented\allowbreak{}Prefetch} & Path-oriented prefetch heuristic (PRT-ART thesis 2026). Pre-loads all nodes along predicted trie path before search starts. Tracker follows search path and recommends next address via linear step extrapolation. V11.1 Hot-Path-Hints: interprets first 8 bytes of key as uint64 address. Trigger: path-oriented prediction; heuristic: linear (step extrapolation); granularity: bundle (multiple cache-lines). Related literature: Chen/\allowbreak{}Gibbons/\allowbreak{}Mowry SIGMOD 2001 (p\allowbreak{}B+-Trees Prefetching) + Mowry 1994. & P21 (related: Chen/\allowbreak{}Gibbons/\allowbreak{}Mowry SIGMOD 2001); thesis 2026 \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} P01, P07, P21~\cite{leis2013art,wu2019wormhole,chen2001prefetch}.

\emph{Related literature:} P23, P25, P26, P27~\cite{khan2010adaptive,mahling2025hotpath,zhang2024pathprefix,zhang2025hierarchical}.

\section{T8 Concurrency Synchronization Axis (Concurrency Axis T8)}

This axis is the \emph{concurrency organ} (T8) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

This axis controls synchronization patterns and safe memory reclamation for concurrent access to tree/\allowbreak{}index structures. The axis provides compile-time selectable synchronization patterns (None/\allowbreak{}Blocking/\allowbreak{}Reader-Writer/\allowbreak{}Optimistic/\allowbreak{}Lock-Free/\allowbreak{}Wait-Free) as well as orthogonal reclamation schemes (RCU/\allowbreak{}Hazard-Pointer) for safe memory deallocation in lock-free structures.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
CC1 & Synchronization Pattern (CC1) & Compile-time variant family of synchronization patterns: None (single-threaded baseline), Blocking (pessimistic mutex), Reader\allowbreak{}Writer (shared\_\allowbreak{}mutex), Optimistic (lock-free version validation), Lock\allowbreak{}Free (CAS-loops), Wait\allowbreak{}Free (bounded steps). Each family has distinct runtime characteristics (0-overhead to mutex-overhead). \\
CC2 & Reclamation Scheme (CC2) & Orthogonal compile-time variant family for safe memory deallocation in lock-free structures: RCU (deferred grace-period via userspace-rcu), Hazard-Pointer (per-thread in-use marks). Composable with lock-free patterns, optional for Blocking/\allowbreak{}Optimistic. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{8.0cm} >{\raggedright\arraybackslash}p{2.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{None\allowbreak{}Concurrency} & No synchronization, single-threaded baseline. Both acquire() and release() are no-ops (volatile dummy). Reference null-point for concurrency-overhead measurement (F15). & Code/\allowbreak{}Baseline \\
\texttt{Blocking\allowbreak{}Concurrency} & Pessimistic mutex locking with std::mutex. Global coarse-grained lock. Serializes all accesses. Upper-bound for locking overhead. Paper: Dijkstra CACM 1965. & Code/\allowbreak{}Baseline \\
\texttt{Reader\allowbreak{}Writer\allowbreak{}Concurrency} & Single-writer/\allowbreak{}multi-reader via std::shared\_\allowbreak{}mutex. Lock-sharing for read-heavy workloads. Middleweight between pessimistic and optimistic patterns. Paper: Courtois et al. CACM 1971. & Code/\allowbreak{}Baseline \\
\texttt{Olc\allowbreak{}Optimistic\allowbreak{}Concurrency} & Optimistic Lock-Coupling (ART-Sync): version-counter per node, lock-free readers with retry-on-conflict. Cheap under low contention. Paper: Leis Da\allowbreak{}Mo\allowbreak{}N 2016 /\allowbreak{} DEBULL 2019 (P08, is\_\allowbreak{}original=true, Apache-2.0, unodb). & P08 \\
\texttt{Lock\allowbreak{}Free\allowbreak{}Concurrency} & Lock-free via Compare-And-Swap (std::atomic CAS-loops). System-wide progress guarantee. Generic CAS-pattern without blocking. Paper: Michael/\allowbreak{}Scott PODC 1996 (anchor paper). & Code/\allowbreak{}Baseline \\
\texttt{Wait\allowbreak{}Free\allowbreak{}Concurrency} & Wait-free: each thread completes in bounded steps (strongest guarantee). Without retry-loop like Lock\allowbreak{}Free. Realized via sequence-lock /\allowbreak{} fetch-add. Paper: Herlihy TOPLAS 1991 (theory anchor). & Code/\allowbreak{}Baseline \\
\texttt{Rcu\allowbreak{}Concurrency} & Read-Copy-Update (Mc\allowbreak{}Kenney OLS 2001): lock-free readers via per-thread nesting counter (relaxed-order, no memory-barrier on read-side). Reclamation-scheme orthogonal to pattern. Local re-impl (P29 userspace-rcu LGPL-2.1, is\_\allowbreak{}original=false). & P29 \\
\texttt{Hazard\allowbreak{}Pointer\allowbreak{}Concurrency} & Hazard Pointers (Michael TPDS 2004, P30): per-thread marked in-use pointers prevent premature deallocation (ABA-safe). Reclamation-scheme. Local re-impl (P30 NO LICENSE, is\_\allowbreak{}original=false). SEQ\_\allowbreak{}CST-store on acquire(). & P30 \\
\texttt{Olc\allowbreak{}Reserved\allowbreak{}Blocks\allowbreak{}Concurrency} & OLC + cache-line-reserved value blocks (multi-writer cache-coherence-storm avoidance). Specialization of Optimistic pattern with additional atomic block-ID reservation (fetch\_\allowbreak{}add). F.6 Doku 19 (prt-art optional, not is\_\allowbreak{}original). & Code/\allowbreak{}Baseline \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} Dijkstra 1965, Courtois et al.\ 1971, P08, Michael/\allowbreak{}Scott 1996, Herlihy 1991, P29, P30~\cite{dijkstra1965concurrent,courtois1971readers,leis2016olc,michael1996queue,herlihy1991waitfree,mckenney2001rcu,michael2004hazard}.

\section{T9 Serialization (Encoding and Compression)}

This axis is the \emph{serialization organ} (T9) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Axis 10 determines how data is encoded and compressed within the Cache-Engine system: from raw-binary storage layouts to variable-length encoding and succinct bit-packing. The serialization strategy choice directly impacts CPU cost (delta-encoding, bit-manipulation, LEB128 varint), memory overhead, and IO-bandwidth utilization --- critical for LSM-trees and read-heavy workloads with approximate filters (Su\allowbreak{}RF/\allowbreak{}LOUDS).

\subsection*{Static sub-axes}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
SR1 & Byte-Order and Word Format & How are bytes arranged within a record? Raw (memcpy raw-layout), Varlen (LEB128-varint with signaling bits), or Packed (bit-by-bit packing)? \\
SR2 & Density (Memory Utilization) & Memory overhead per record: Sparse (complete records), Dense (with padding), or Succinct (information-theoretically dense, n*H + o(n) bits)? \\
SR3 & Compression and Preprocessing & Block compression (LZ4/\allowbreak{}Snappy, LZ77-based) or delta-encoding (zigzag-transform for negative deltas)? Or no compression? \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{8.5cm} >{\raggedright\arraybackslash}p{2.0cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Raw\allowbreak{}Binary\allowbreak{}Serialization} & Raw byte storage layout (memcpy). Baseline without transformation. Minimal CPU cost. Typical for in-memory workloads without compression. Variant: \texttt{serialize\_\allowbreak{}scan()} reads 32-bit fields + order-sensitive checksum. & Code/\allowbreak{}Baseline \\
\texttt{Var\allowbreak{}Len\allowbreak{}Serialization} & Variable-length encoding with signaling bits (LEB128-Var\allowbreak{}Int). Standard for ART (Leis ICDE 2013 P01). Small values = few bytes. Signaling bits mark byte-length. Data-dependent loop (1-5 bytes). FNV-1a-mix of varint bytes. & P01 \\
\texttt{Succinct\allowbreak{}Serialization} & Bit-by-bit packing with minimal bit-width (LOUDS/\allowbreak{}Su\allowbreak{}RF). Achieves information-theoretic minimum n*H + o(n) bits. Read-only after build. Optimal for approx-filters + read-heavy. Highest per-element CPU cost (real bit-manipulation). Variant: \texttt{serialize\_\allowbreak{}scan()} with 64-bit accumulator. & P10,P09 \\
\texttt{Compressed\allowbreak{}Serialization} & Block compression (LZ4/\allowbreak{}Snappy) layered on raw-binary. Trade-off: CPU cost vs IO/\allowbreak{}memory bandwidth. Variant: delta-encoding against predecessor + zigzag-transform (negative delta $\rightarrow$ small unsigned). Typical for LSM-trees. Medium CPU cost. & Code/\allowbreak{}LZ77 (Ziv/\allowbreak{}Lempel 1977)~\cite{ziv1977lz77} \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} P01, P09, P10, Ziv/\allowbreak{}Lempel 1977~\cite{leis2013art,jacobson1989louds,zhang2018surf,ziv1977lz77}.

\section{T10 Telemetry and Measurement Instrumentation (axis\_\allowbreak{}11)}

This axis is the \emph{telemetry organ} (T10) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Defines strategies for measuring index operations (insert, lookup, erase) with different overhead profiles. Supports density tracking for adaptive node-type selection, global insert counters for throughput measurements, HDR histograms for latency quantiles, and leaf-only counters to avoid cache-line false sharing in measurement cycles.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
TM1 & Scope & Validity scope of counter management: leaf-only (leaf nodes only), per-node (all nodes), or global (single accumulator for entire structure). Compile-time choice, impacts overhead through contention on shared cache lines. \\
TM2 & Metric Type & Type of measured metric: counter (event counter), histogram (distribution with quantiles), or density (per-node slot occupancy). Impacts memory footprint and measurement granularity. \\
TM3 & Overhead Level & Measurement overhead magnitude: low (atomic operations), medium (per-node data structures), high (histogram logging). Determines suitability for different workload scenarios (benchmarking vs.\ production). \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{2.5cm} >{\raggedright\arraybackslash}p{7.5cm} >{\raggedright\arraybackslash}p{3.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Leaf\allowbreak{}Only\allowbreak{}Counter} & Counter in leaf nodes only (avoids cache-line ping-pong by skipping inner-node updates). Low overhead with leaf-only scope. Implements coherence-friendly counter concept from Kuehn Da\allowbreak{}Mo\allowbreak{}N 2023 motivation, although original paper describes cache-friendly B+-layout, not leaf-counter algorithm --- CE baseline for false-sharing avoidance. & P28 (Kuehn Da\allowbreak{}Mo\allowbreak{}N 2023, context; Code/\allowbreak{}Baseline) \\
\texttt{Density\allowbreak{}Tracker} & Per-node density/\allowbreak{}slot-occupancy tracking for adaptive node-type selection (4→16→48→256). Contextual telemetry tied to ART structure (Leis ICDE 2013), measures slot occupancy across all nodes. Higher overhead than Leaf\allowbreak{}Only, but provides adaptation signal for dynamic rebalancing decisions. & P01 (ART/\allowbreak{}Leis ICDE 2013, context; Code/\allowbreak{}Baseline) \\
\texttt{Insert\allowbreak{}Counter} & Global atomic insert counter with minimal overhead (1 atomic increment per insert). Contextual telemetry tied to HOT structure (Binna SIGMOD 2018), global scope. Sufficient for throughput measurement series, no latency granularity. & P02 (HOT/\allowbreak{}Binna SIGMOD 2018, context; Code/\allowbreak{}Baseline) \\
\texttt{Latency\allowbreak{}Histogram} & HDR histogram for lookup latency distribution (p50/\allowbreak{}p95/\allowbreak{}p99 quantiles). Based on Hdr\allowbreak{}Histogram (Gil Tene), CC0-1.0 license. Highest overhead (per lookup), but provides complete latency statistics for SLA-oriented and research-grade measurement series. Only is\_\allowbreak{}original candidate of the axis. & Hdr\allowbreak{}Histogram (github.com/\allowbreak{}Hdr\allowbreak{}Histogram/\allowbreak{}Hdr\allowbreak{}Histogram\_\allowbreak{}c, CC0-1.0, is\_\allowbreak{}original) \\
Probability-Hints & Probability hints of the cache-oblivious Bender line (P16/\allowbreak{}P17): distribution- and probability-related occupancy hints that serve as input to the layout decisions. Paper building block of the axis dissection (reconstruction target of this thesis, no CE baseline). & P16/\allowbreak{}P17 (Bender 2000/\allowbreak{}2005)~\cite{bender2000cobtree,bender2005coboblivious} \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} Hdr\allowbreak{}Histogram (Gil Tene)~\cite{tene_hdrhistogram}, P16/\allowbreak{}P17 (Bender)~\cite{bender2000cobtree,bender2005coboblivious}.

\emph{Related literature:} P01, P02, P28~\cite{leis2013art,binna2018hot,kuehn2023bplustree}.

\section{T11 Value Handle Storage Strategy}

This axis is the \emph{value-handle organ} (T11) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

The Value-Handle axis defines how data values (Values) are stored and accessed in the cache index: inline within node slots, externally in a pool, via immutable-shared references (RCU-style), versioned pointers (MVCC), or chained external references (multi-value). This axis characterizes the storage-layout strategy for values and determines access latency (direct vs.\ indirect) and concurrency semantics (snapshot vs.\ versioning).

\subsection*{Static sub-axes}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
VH1 & Storage Location & Compile-time variants for physical value placement: inline in node slot (combines key/\allowbreak{}value without indirection) or external in separate pool (node holds offset/\allowbreak{}pointer only). \\
VH2 & Ownership Semantics & Compile-time variants for ownership and sharing strategy: owned (exclusive per key), shared (immutable reference-counted via RCU-style), weak (borrowing without increment). \\
VH3 & Versioning Strategy & Compile-time variants for concurrency control: no versioning (simple ownership), version-tags in pointer (MVCC for reader snapshot isolation). \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.0cm} >{\raggedright\arraybackslash}p{2.3cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Inline\allowbreak{}Value\allowbreak{}Handle} & Value embedded directly in node slot (no indirection). Standard for small values like uint64\_\allowbreak{}t. Minimal access latency (1 direct read), but node memory consumed by larger values. Source: ART (Leis ICDE 2013, P01). & P01 \\
\texttt{External\allowbreak{}Pool\allowbreak{}Value\allowbreak{}Handle} & Value stored externally in pool; node holds offset only. Enables variable value sizes and compact nodes, costs 1 additional cache miss per lookup (pointer chasing). Source: Wormhole (Wu Euro\allowbreak{}Sys 2019, P07). & P07 \\
\texttt{Immutable\allowbreak{}Shared\allowbreak{}Ref\allowbreak{}Value\allowbreak{}Handle} & Value via reference-counted pointer (RCU-style snapshot sharing). Optimal for reader concurrency: readers see stable snapshots, writers create new immutable snapshots. Requires Ref\allowbreak{}Count acquire/\allowbreak{}release per access. Source: RCU (Mc\allowbreak{}Kenney OLS 2001, P29). & P29 \\
\texttt{Versioned\allowbreak{}Pointer\allowbreak{}Value\allowbreak{}Handle} & Pointer with version tag in upper byte (MVCC). Enables multi-version concurrency control: readers get snapshot version for visibility control, writers create new version. Tombstone detection via version bit. Source: Masstree (Mao Euro\allowbreak{}Sys 2012, P03) and SMART (OSDI 2023). & P03 \\
\texttt{Chain\allowbreak{}Ref\allowbreak{}Value\allowbreak{}Handle} & Multi-value chained external reference (linked-list). Node holds offset to chain head in pool; each chain node holds (value\_\allowbreak{}offset, next\_\allowbreak{}offset). Most expensive variant with 2 dependent dereferences per value access. Source: prt-art (local reimplementation, F.6 migration). & Code/\allowbreak{}Baseline \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} P01, P03, P07, P29, SMART (OSDI 2023)~\cite{leis2013art,mao2012masstree,wu2019wormhole,mckenney2001rcu,luo2023smart}.

\section{T12 Instruction Set Architecture (ISA)}

This axis is the \emph{ISA organ} (T12) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Axis T12 specifies the target CPU instruction set architecture (ISA) as a compile-time constant, enabling cross-platform optimizations across four dominant server architectures: x86-64 (Intel/\allowbreak{}AMD), AArch64 (ARM), Power-ISA (IBM), and RISC-V (open-source). Each variant defines a hardware-specific SIMD baseline (SSE2, NEON, VSX, scalar) and thus characterizes the vector acceleration potential for cache-engine operations such as \texttt{simd\_\allowbreak{}field\_\allowbreak{}sum} (accumulation of 32-bit word fields).

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
IS1 & Word Size & CPU data-path width: 32-bit, 64-bit, or 128-bit. In the cache engine standardized to 64-bit for all four ISA profiles. \\
IS2 & CPU Vendor/\allowbreak{}Family & Instruction set family: \texttt{intel\_\allowbreak{}amd} (x86-64), \texttt{arm} (AArch64), \texttt{ibm} (Power-ISA), \texttt{open} (RISC-V). Determines SIMD extension compatibility: SSE2/\allowbreak{}AVX2/\allowbreak{}AVX-512 (x86-64), NEON/\allowbreak{}SVE/\allowbreak{}SVE2 (AArch64), VSX (Power), RVV (RISC-V). \\
IS3 & ZIH Cluster Affinity & Production machines at TU Dresden: Barnard (Intel Xeon Platinum 8470, Sapphire Rapids, x86-64), Capella (AMD EPYC 9334, x86-64), Grace\allowbreak{}Hopper (NVIDIA Grace Hopper ARMv9 Neoverse V2, AArch64), N/\allowbreak{}A (research boards, RISC-V/\allowbreak{}Power). Enables hardware-specific experiments with real production cluster profiles. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{2.4cm} >{\raggedright\arraybackslash}p{7.6cm} >{\raggedright\arraybackslash}p{3.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Amd64\allowbreak{}Isa} & x86-64 /\allowbreak{} Intel 64 ISA (AMD/\allowbreak{}Intel Server+Desktop). Dominant server platform with SSE2 SIMD baseline (4 uint32 lanes per 128-bit register). Implements genuine SSE2 vectorization via \texttt{\_\allowbreak{}mm\_\allowbreak{}loadu\_\allowbreak{}si128} + \texttt{\_\allowbreak{}mm\_\allowbreak{}add\_\allowbreak{}epi32} on native x86-64 hosts; scalar fallback on non-x86 build hosts. ZIH presence: Barnard (Intel Xeon Platinum 8470, Sapphire Rapids), Capella (AMD EPYC 9334). & Vendor spec (AMD 2000 /\allowbreak{} Intel ongoing) \\
\texttt{Aarch64\allowbreak{}Isa} & ARMv8-A 64-bit ISA (ARM Holdings). Growing server relevance (Apple M-Series, AWS Graviton, Ampere Altra). SIMD baseline: NEON (128-bit, 4 uint32 lanes). Cache engine implementation: 4-way unrolled scalar fallback (models NEON width without ARM intrinsics on x86 build hosts); genuine NEON vectorization on native ARMv8 hosts. ZIH presence: Grace\allowbreak{}Hopper (NVIDIA Grace Hopper, Neoverse V2 + Hopper GPU). & Vendor spec (ARM Ltd., since 2011/\allowbreak{}2013) \\
\texttt{Power\allowbreak{}Pc\allowbreak{}Isa} & Power ISA (POWER9/\allowbreak{}10, ppc64le little-endian). IBM Power Systems (AC922, IC922, S1022). SIMD baseline: VSX (Vector-Scalar Extension, 128-bit, 4 uint32 lanes). Cache engine implementation: 4-way unrolled scalar fallback (models VSX width without Power intrinsics on x86 build hosts). Relevant for HPC workloads with NVLink GPU connectivity (AC922). ZIH presence: not primary, but available for research comparisons. & Vendor/\allowbreak{}standards spec (IBM/\allowbreak{}Open\allowbreak{}POWER, v3.0 2017, v3.1 2020) \\
\texttt{Risc\allowbreak{}VIsa} & RISC-V RV64\allowbreak{}GC open-source ISA (RV64 General + Compressed). Growing relevance (Si\allowbreak{}Five, Alibaba Xuantie, Andes, ETH Zurich cores). SIMD baseline: NONE (RVV is optional, not part of RV64\allowbreak{}GC baseline). Cache engine implementation: deliberately plain scalar path (unrolled) -- models lack of vector unit in base ISA and thus creates real, slower runtime versus x86-64 and ARM\@. ZIH presence: embedded + research boards. & Standards spec (RISC-V Intl.; UCB TR since 2011) \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} ---

\section{T13 Index Organization}

This axis is the \emph{index-organization organ} (T13) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Classification of data storage organization and index structuring across three orthogonal dimensions: storage-order (heap/\allowbreak{}clustered/\allowbreak{}unordered), index-count (none/\allowbreak{}single/\allowbreak{}multiple), and data-embedding (separate vs.\ leaf-embedded). This axis models fundamental DBMS architectural decisions that decisively influence cache behavior, access patterns, and I/\allowbreak{}O characteristics based on alignment of index-order with storage-order.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
\texttt{io1\_\allowbreak{}storage\_\allowbreak{}order} & Storage Order & Arrangement of tuples in storage: unordered (insertion-order in heap), sorted by primary key (clustered), or via pointer indirection (non-clustered). Determines sequential vs.\ random access patterns during range scans. \\
\texttt{io2\_\allowbreak{}index\_\allowbreak{}count} & Index Multiplicity & Number of supported indexes per table: none (heap baseline), single primary index (clustered), or multiple secondary indexes (non-clustered). Influences OLTP workload support. \\
\texttt{io3\_\allowbreak{}data\_\allowbreak{}embedding} & Data Embedding & Storage placement of data relative to index: separate row storage (Clustered/\allowbreak{}Non\allowbreak{}Clustered with pointer-hop), or directly embedded in index leaf-pages (IOT/\allowbreak{}trie-style like ART/\allowbreak{}HOT). \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.5cm} >{\raggedright\arraybackslash}p{6.0cm} >{\raggedright\arraybackslash}p{4.0cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Clustered\allowbreak{}Index\allowbreak{}Organization} & Storage-order aligns with index-order (1 primary key). Optimized for range-scans along primary key with sequential disk reads (cache-friendly). Standard: My\allowbreak{}SQL Inno\allowbreak{}DB, Postgre\allowbreak{}SQL CLUSTER, SQL Server CLUSTERED INDEX\@. Access pattern: sequential forward iteration without predicate evaluation, O(n) aggregation scan. & Bayer \& Mc\allowbreak{}Creight, Acta Informatica 1972 (10.1007/\allowbreak{}BF00288683) \\
\texttt{Heap\allowbreak{}Index\allowbreak{}Organization} & Storage-order = insertion-order, NO index (baseline). Lookup via O(n) full-scan with per-record predicate evaluation (data-dependent branch, no early exit). Standard: Postgre\allowbreak{}SQL heap tables, HBase. Access pattern: unordered sequential scan with comparison overhead and branch misprediction. & Garcia-Molina, Ullman \& Widom, Database Systems (Textbook Baseline) \\
\texttt{Iot\allowbreak{}Index\allowbreak{}Organization} & Data directly embedded in index leaf-pages (Oracle IOT, SQL Server CLUSTERED INDEX style). Saves pointer indirection vs.\ non-clustered. Equivalent to B+tree leaves with embedded data or trie nodes (ART/\allowbreak{}HOT/\allowbreak{}Wormhole). Access pattern: sequential like clustered, but with overhead from fat leaf layout (two fields per record). & Srinivasan et al., VLDB 2000 (Oracle8i IOT)~\cite{srinivasan2000oracle8iiot} \\
\texttt{Non\allowbreak{}Clustered\allowbreak{}Index\allowbreak{}Organization} & Index-order $\neq$ storage-order, multiple secondary indexes per table. Lookup via index + pointer-hop to row data (random-stride, cache-misses). Standard: SQL Server NONCLUSTERED INDEX, Postgre\allowbreak{}SQL CREATE INDEX\@. Access pattern: index navigation followed by random storage hops (LCG-deterministic, REAL cache-miss latency from TLB/\allowbreak{}L2 invalidation). & Comer, ACM Computing Surveys 1979 (The Ubiquitous B-Tree, 10.1145/\allowbreak{}356770.356776) \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} Bayer \& Mc\allowbreak{}Creight 1972, Comer 1979, Srinivasan et al.\ 2000, Garcia-Molina, Ullman \& Widom (Textbook)~\cite{bayer1972btree,comer1979btree,srinivasan2000oracle8iiot,garciamolina2009dbsystems}.


\section{T14 I/\allowbreak{}O Dispatch Axis}

This axis is the \emph{I/O-dispatch organ} (T14) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Modeling and selection of I/\allowbreak{}O strategies for persistence and memory access in index operations: in-memory storage only (baseline), buffered via OS page cache (standard), direct I/\allowbreak{}O with \texttt{O\_\allowbreak{}DIRECT} flag (NVMe-optimized), or memory-mapped file I/\allowbreak{}O (persistent memory). This axis decouples search and tree structures from persistence implementation and enables A/\allowbreak{}B measurements of dispatch overhead profiles under identical workloads.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
IO1 & Persistence Model & Classifies storage type: RAM-only (no persistence, pure measurement baseline), or disk/\allowbreak{}persistent-memory (file-backed indices with various caching strategies). \\
IO2 & Caching Strategy & Characterizes OS buffer usage: Buffered (OS page cache, read-ahead + write-back, standard), Direct I/\allowbreak{}O (\texttt{O\_\allowbreak{}DIRECT}, page cache bypass, 512\allowbreak{}B/\allowbreak{}4\allowbreak{}KB alignment), or Memory-Mapped (file-backed \texttt{mmap} with automatic OS paging and page-fault-driven access). \\
IO3 & Write Durability & Defines consistency guarantees: no \texttt{fsync} (highest performance, no durability), \texttt{fsync} after batch (compromise, batch commits), or atomic writes (highest safety, lowest performance). In this implementation primarily present as a conceptual sub-axis; operational realization occurs through the caching strategy. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{2.5cm} >{\raggedright\arraybackslash}p{8.5cm} >{\raggedright\arraybackslash}p{2.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{In\allowbreak{}Memory\allowbreak{}Only} & Pure RAM-based index without persistence. Direct sequential record access without dispatch overhead. Fastest path and measurement baseline of the axis. Used for performance reference and measurement under zero-I/\allowbreak{}O-latency conditions. & Code/\allowbreak{}Baseline \\
\texttt{Buffered\allowbreak{}Io} & Standard \texttt{read()}/\allowbreak{}\texttt{write()} via OS page cache with automatic read-ahead and write-back. Default for general-purpose DBMS\@. Trade-off: convenient but double-buffering (app cache + OS cache results in higher memory footprint). Paper anchor: Stonebraker CACM 1981 (position paper Operating System Support for Database Management, DOI 10.1145/\allowbreak{}358699.358703). Confidence: medium (engineering pattern, no algorithm code). & P-Stonebraker-1981 \\
\texttt{Direct\allowbreak{}Io} & Bypasses OS page cache via \texttt{O\_\allowbreak{}DIRECT} flag (\texttt{open(2)}), NVMe-SSD-optimal. Requires 512\allowbreak{}B/\allowbreak{}4\allowbreak{}KB alignment. Enables DBMS-owned cache strategies (Rocks\allowbreak{}DB, My\allowbreak{}SQL). Higher CPU cost but no cache pollution from background activity. No dedicated paper source (Linux API since 2.4.10); treated as intentional engineering baseline. & Code/\allowbreak{}Baseline \\
\texttt{Mmap\allowbreak{}Io} & Memory-mapped file I/\allowbreak{}O (\texttt{mmap}) with persistent memory or read-heavy workloads. OS paging automatic, no copy between user and kernel space. Trade-off: page-fault latency spikes vs.\ reuse-friendliness. Paper anchor: Crotty et al. CIDR 2022 (anti-pattern baseline Are You Sure You Want to Use MMAP in Your DBMS?, db.cs.cmu.edu/\allowbreak{}\ldots/\allowbreak{}cidr2022-p13-crotty.pdf). Deliberately cited as anti-pattern reference to document limits and overhead. OSS reference MIT-licensed. & P-Crotty-2022 \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} Stonebraker-1981, Crotty-2022~\cite{stonebraker1981os,crotty2022mmap}

\section{T15 Migration Strategy Axis}

This axis is the \emph{migration organ} (T15) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

The Migration Strategy axis (T15, \texttt{axis\_\allowbreak{}migration}) encapsulates decision logic for data placement across multi-tier storage hierarchies (RAM/\allowbreak{}SSD/\allowbreak{}HDD, Hot/\allowbreak{}Cold tiers). It models when and how frequently accessed data should be promoted to faster storage or demoted when cold, measuring decision logic only without actual tier migration. Supports four concrete strategies: static (baseline), Hot/\allowbreak{}Cold separation, multi-tier hierarchy, ML-driven adaptive.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
MG1 & Trigger Mechanism (MG1) & Determines when migration decision is triggered: none (static), threshold-based, or observation-driven. Compile-time variant per strategy. \\
MG2 & Migration Direction (MG2) & Determines direction of data movement: none (static), up-migration (RAM→SSD), down-migration (SSD→RAM), or bidirectional. Compile-time variant. \\
MG3 & Migration Granularity (MG3) & Specifies the unit granularity for migration decisions: none (static), page (4\allowbreak{}KB), block (64\allowbreak{}KB), object (single item). Compile-time category. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Dynamic sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Parameter & Description \\
\midrule
\endhead%
D1 & Decision Threshold & Runtime parameter: Hot/\allowbreak{}Cold classification boundary (e.g., recency counter limit for LRU-based promotions/\allowbreak{}demotions). User-configurable. \\
D2 & Feature Scoring Weights & Runtime parameters for Adaptive\allowbreak{}Migration: weights (w\_\allowbreak{}freq, w\_\allowbreak{}rec) for frequency and recency features in ML linear combination. Self-tuning possible. \\
D3 & Tier Configuration & Runtime: number of storage tiers (e.g., k\allowbreak{}Tiers=3 for RAM/\allowbreak{}SSD/\allowbreak{}HDD in Tier\allowbreak{}Based\allowbreak{}Migration), capacities per tier, promotion/\allowbreak{}demotion thresholds. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{7.0cm} >{\raggedright\arraybackslash}p{3.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{No\allowbreak{}Migration} & Static placement without migration (baseline). Data positioned once, no movement between tiers. Benchmark zero-point for comparison with active strategies. & Code/\allowbreak{}Baseline \\
\texttt{Hot\allowbreak{}Cold\allowbreak{}Migration} & Hot/\allowbreak{}Cold separation via LRU + probabilistic recency counting. Frequently accessed data (hot) promoted to faster cache, rarely accessed (cold) demoted to slow storage. Standard for LRU-K and LFU cache systems. & Anti-Caching (PVLDB 2013)~\cite{debrabant2013anticaching} \\
\texttt{Tier\allowbreak{}Based\allowbreak{}Migration} & Multi-tier migration RAM→NVM→SSD→HDD with modulo-based tier mapping. Each data item mapped to target tier via modulo field value. Standard for Rocks\allowbreak{}DB, LSM-trees, and tiered cache systems with configurable thresholds. & Managing NVM in DBMSs (SIGMOD 2018)~\cite{vanrenen2018nvm} \\
\texttt{Adaptive\allowbreak{}Migration} & ML-driven adaptive migration with online learning (Cachelib/\allowbreak{}Le\allowbreak{}Ca\allowbreak{}R-style). Weighted linear combination of frequency and recency features computes adaptive score; running learning accumulator models online learning. Higher overhead versus optimal hit-rate. & ML-based Le\allowbreak{}Ca\allowbreak{}R (USENIX Hot\allowbreak{}Storage 2018)~\cite{vietri2018lecar} \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} Anti-Caching (PVLDB 2013), LeCaR (Hot\allowbreak{}Storage 2018), Managing NVM (SIGMOD 2018)~\cite{debrabant2013anticaching,vietri2018lecar,vanrenen2018nvm}.

\section{T16 Filter Axis}

This axis is the \emph{filter organ} (T16) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

The Filter Axis T16 provides selectable probabilistic and deterministic membership-query filters for point and range queries. It covers classic Bloom filters with k hash functions, the deletable Cuckoo Filter with bucket hashing and fingerprints, the succinct Range Filter Su\allowbreak{}RF using LOUDS-Trie navigation, and the space-efficient Xor Filter with constant 3-way-XOR latency. Purpose: Make space, query latency, mutability, and false-positive-rate trade-offs directly measurable in cache-engine experiments.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
FT1 & Query Type & Distinction between point queries (membership: key present?) and range queries (range: keys in [a,b]?). Point filters are Bloom/\allowbreak{}Cuckoo/\allowbreak{}Xor; only Range\allowbreak{}Surf\allowbreak{}Filter supports range semantics. \\
FT2 & Mutability & Division into immutable filters (Bloom, Range\allowbreak{}Surf, Xor immutable after build) vs.\ mutable (Cuckoo with delete operation). Determines wrapper design and caching possibilities during runtime. \\
FT3 & Error Profile & Characterization of error-tolerant semantics: false-positive-only (Bloom, Cuckoo, Xor --- filter answers \enquote{yes} incorrectly) vs.\ lossless (deterministic answer). Bloom: approx.\ k*n hash tests; Xor: exactly 3 XOR ops (branch-free). \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{2.6cm} >{\raggedright\arraybackslash}p{7.4cm} >{\raggedright\arraybackslash}p{3.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Bloom\allowbreak{}Filter} & Classical Bloom Filter per Bloom 1970: m-bit bitmap, k independent hash functions (double-hashing per Kirsch/\allowbreak{}Mitzenmacher 2006). Point queries only; false positives possible, no deletes. Implementation with a real persistent m-bit bitmap (8~KiB, m=65536, k=4 double-hashing), probed in measurement path~B over the actually inserted keys; the honestly declared pseudo-bitmap probe (k=4 hash tests per query, characteristic early-exit latency on mismatch) remains only in the path-A segment-timing scan. & Bloom 1970 (CACM 13:422-426) \\
\texttt{Cuckoo\allowbreak{}Filter} & Cuckoo Filter per Fan et al. Co\allowbreak{}NEXT 2014: Bucketed cuckoo hashing with 1-byte fingerprints. Supports deletes and lookup counting (advantage over Bloom). Probe characteristic: 2 bucket lookups + fingerprint comparison (partial-key cuckoo per Co\allowbreak{}NEXT 2014 \S3: i2 = i1 XOR hash(fp)). Higher insert costs, but flexible mutability. & Fan et al. Co\allowbreak{}NEXT 2014 (DOI:10.1145/\allowbreak{}2674005.2674994) \\
\texttt{Range\allowbreak{}Surf\allowbreak{}Filter} & Su\allowbreak{}RF (Succinct Range Filter) per Zhang et al. SIGMOD 2018: LOUDS-bitmap trie for range queries. Only filter with range semantics (lower\_\allowbreak{}bound/\allowbreak{}upper\_\allowbreak{}bound at byte-prefix level). Probe characteristic: multi-level trie descent with succinct-LOUDS navigation (up to k\allowbreak{}Max\allowbreak{}Depth=6), not exactly 1 hash test, hence latency variable per prefix length. Practical in Rocks\allowbreak{}DB SST optimizations. & P10 (Zhang et al. SIGMOD 2018) \\
\texttt{Xor\allowbreak{}Filter} & Xor Filter per Graf+Lemire 2020 (JEA 25): XOR-based filter with approx. 9 bits/\allowbreak{}key (vs. 10 bits/\allowbreak{}key Bloom at same FP rate). Immutable after build. Probe characteristic: exactly 3 independent table accesses (3 disjoint table thirds h0, h1, h2) + 3-way XOR, branch-free, constant latency without early-exit. Safe against path optimization, smallest memory footprint of all filters. & Graf+Lemire 2020 (JEA 25:Article 1.5, DOI:10.1145/\allowbreak{}3376122) \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} Bloom 1970, Fan et al. Co\allowbreak{}NEXT 2014, Zhang et al. SIGMOD 2018, Graf+Lemire JEA 2020~\cite{bloom1970,fan2014cuckoo,zhang2018surf,graf2020xor}.

\section{T17 Buffering Q1 (Operations Buffer)}

This axis is the \emph{buffering organ (Q1)} (T17) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

The \texttt{axis\_\allowbreak{}q1\_\allowbreak{}queuing} realizes various buffering strategies for operations in the cache-engine context. It encompasses textbook ADTs (FIFO, LIFO, Append-Only, Priority-Heap), specialized versioning techniques (Delta-Chain, Copy-on-Write, Epoch, Tombstone), and modern lock-free queues (SPSC/\allowbreak{}MPMC following Lamport and Vyukov). Strategies are selected at compile time via the axis registry (the \texttt{mp\_list} of research wave 2, filtered by enable flags) per permutation binary (no runtime switch) and are classifiable by access pattern (sequential, ordered, cyclic, versioned, batched, lock-free).

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
QS1 & Sequential Access & FIFO/\allowbreak{}LIFO/\allowbreak{}Append-Only: Access at head or tail, typical for write-coalescing and memtable buffering \\
QS2 & Ordered Access & Skip-List/\allowbreak{}Priority-Heap: Sorted drain, typical for LSM compaction and hot-key eviction \\
QS3 & Cyclic Access & Bounded Ring-Buffer: Fixed capacity, Disruptor pattern, wrap-around with overflow handling \\
QS4 & Versioned Access & Delta-Chain/\allowbreak{}Tombstone/\allowbreak{}Co\allowbreak{}W/\allowbreak{}Epoch: MVCC, snapshot isolation, reclamation windows \\
QS5 & Batched Operations & Batched-Insert-Buffer: Accumulation into sub-batches, SIMD vectorization, cache-locality optimization \\
QS6 & Lock-Free Concurrent & SPSC/\allowbreak{}MPMC\@: CAS-loop or wait-free Lamport model, true lock-freedom without mutex \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{8.0cm} >{\raggedright\arraybackslash}p{2.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{No\allowbreak{}Buffer} & Passthrough/\allowbreak{}Identity (Baseline): \texttt{put()} is no-op, \texttt{get()} always returns \texttt{std::nullopt}. Useful as control point and in classic B+/\allowbreak{}ART scenarios with direct leaf access. Family Q01, QS1 \texttt{sequential\_\allowbreak{}access}. & Code/\allowbreak{}Baseline \\
\texttt{Append\allowbreak{}Only\allowbreak{}Buffer} & Append-Only Linear Buffer (LSM memtable, Bw-Tree delta): monotonically growing \texttt{std::vector} without middle deletion. Optimized for pure append + bulk drain. Family Q02, QS1 \texttt{sequential\_\allowbreak{}access}. References: Levandoski/\allowbreak{}Lomet/\allowbreak{}Sengupta Bw-Tree ICDE 2013, O'Neil LSM 1996. & Code/\allowbreak{}Baseline \\
\texttt{FIFOQueue\allowbreak{}Buffer} & FIFO-Queue (\texttt{std::deque}): First-In-First-Out via \texttt{deque::push\_\allowbreak{}back} /\allowbreak{} \texttt{deque::pop\_\allowbreak{}front}. Unbounded, efficient middle growth. Textbook ADT, typical for write-coalescing. Family Q03, QS1 \texttt{sequential\_\allowbreak{}access}. & Code/\allowbreak{}Baseline \\
\texttt{LIFOStack\allowbreak{}Buffer} & LIFO-Stack (\texttt{std::vector}): Last-In-First-Out via \texttt{vector::push\_\allowbreak{}back} /\allowbreak{} \texttt{vector::pop\_\allowbreak{}back}. Unbounded, optimized for hot-path reuse and version tombstones. Family Q04, QS1 \texttt{sequential\_\allowbreak{}access}. & Code/\allowbreak{}Baseline \\
\texttt{Bounded\allowbreak{}Ring\allowbreak{}Buffer} & Bounded Ring-Buffer (Disruptor Pattern): Fixed capacity, cyclic access with wrap-around. Overflow: drop\_\allowbreak{}oldest (default). \texttt{iterable\_\allowbreak{}aspect\_\allowbreak{}t}: Capacity as runtime loop. Family Q05, QS3 \texttt{cyclic\_\allowbreak{}access}. Reference: LMAX Disruptor 2011. & LMAX 2011 \\
\texttt{Priority\allowbreak{}Heap\allowbreak{}Buffer} & Priority-Heap (\texttt{std::priority\_\allowbreak{}queue}): Max-heap, \texttt{get()} returns highest priority first. First strategy with \texttt{supports\_\allowbreak{}priority\_\allowbreak{}ordering()}=true. Unbounded, textbook ADT (Williams 1964). Application: LRU approximation, hot-key promotion, eviction queues. Family Q06, QS2 \texttt{ordered\_\allowbreak{}access}. & Code/\allowbreak{}Baseline \\
\texttt{Delta\allowbreak{}Chain\allowbreak{}Buffer} & Delta-Chain Buffer (Bw-Tree, Levandoski 2013): First versioned strategy (\texttt{is\_\allowbreak{}versioned()}=true). Each \texttt{put()} appends delta-record with ascending version-ID to top of chain. \texttt{get()} LIFO-drain with version-ID\@. Append-versioning paradigm. Family Q07, QS4 \texttt{versioned\_\allowbreak{}access}. & Bw-Tree (Levandoski et al.\ ICDE 2013)~\cite{levandoski2013bwtree} \\
\texttt{Skiplist\allowbreak{}Buffer} & Skip-List Buffer (sorted buffer, Rocks\allowbreak{}DB-style): \texttt{put()} inserts in sorted order via \texttt{std::set}. \texttt{get()} ASCENDING-drain (minimum first), typical for LSM-compact output. Distinction from \texttt{Priority\allowbreak{}Heap\allowbreak{}Buffer}: skiplist=ascending, heap=descending. Reference: Pugh CACM 1990. Family Q08, QS2 \texttt{ordered\_\allowbreak{}access}. & Pugh 1990, Rocks\allowbreak{}DB \\
\texttt{Tombstone\allowbreak{}Buffer} & Tombstone-Marker Buffer (LSM + MVCC): \texttt{put()} creates live entry, deleted entries remain as tombstones (\texttt{std::nullopt}) until compaction. \texttt{get()} skips tombstones FIFO-like, \texttt{is\_\allowbreak{}versioned()}=true (erase-versioning, unlike delta-chain append-versioning). \texttt{size()} counts live-slots only. Family Q09, QS4 \texttt{versioned\_\allowbreak{}access}. Reference: O'Neil LSM-tree 1996. & O'Neil 1996 \\
\texttt{Copy\allowbreak{}On\allowbreak{}Write\allowbreak{}Buffer} & Copy-on-Write Snapshot-Buffer (Persistent Data Structures): \texttt{put()}/\allowbreak{}\texttt{get()} increments snapshot version, earlier snapshots remain logically via \texttt{shared\_\allowbreak{}ptr}. Snapshot-versioning paradigm (unlike delta-chain append, tombstone erase). \texttt{is\_\allowbreak{}versioned()}=true. Family Q10, QS4 \texttt{versioned\_\allowbreak{}access}. Reference: Driscoll/\allowbreak{}Sarnak/\allowbreak{}Sleator/\allowbreak{}Tarjan JCSS 1989. & Driscoll et al.\ 1989 \\
\texttt{Epoch\allowbreak{}Buffer} & Epoch-based Buffer /\allowbreak{} QSBR (Quiescent-State-Based-Reclamation): Entries accumulated in current epoch. After \texttt{epoch\_\allowbreak{}threshold} transition => prior epoch can be reclaimed. \texttt{iterable\_\allowbreak{}aspect\_\allowbreak{}t}: \texttt{epoch\_\allowbreak{}threshold} as runtime parameter. \texttt{is\_\allowbreak{}versioned()}=true (reclamation-window-versioning). Family Q11, QS4 \texttt{versioned\_\allowbreak{}access}. References: Mc\allowbreak{}Kenney RCU OLS 2001, Masstree Euro\allowbreak{}Sys 2012, SMART OSDI 2023. & P29 (RCU OLS 2001), Masstree 2012 \\
\texttt{Batched\allowbreak{}Insert\allowbreak{}Buffer} & Batched-Insert Buffer (OLAP indexing, ART bulk-insert): \texttt{put()} buffers into sub-batches of \texttt{batch\_\allowbreak{}size\_\allowbreak{}}, complete batches passed on (bulk-drain). Goal: amortized insert cost via bulk operations, SIMD vectorization, cache-locality. First QS5 strategy. Family Q12, QS5 \texttt{batched\_\allowbreak{}access}. Reference: Leis ART ICDE 2013 §6. & P01 (Leis ART ICDE 2013) \\
\texttt{Lock\allowbreak{}Free\allowbreak{}SPSCBuffer} & Lock-Free SPSC Ring-Queue (Lamport 1977): Single-Producer/\allowbreak{}Single-Consumer. \texttt{put()} wait-free (1 atomic.store + 1 atomic.load), \texttt{get()} wait-free. No CAS needed due to disjoint modified-sets. \texttt{Progress\allowbreak{}Guarantee::Lock\allowbreak{}Free} (even wait-free). \texttt{iterable\_\allowbreak{}aspect\_\allowbreak{}t}: Capacity. Overflow: dropping. Contract: exactly 1 producer + 1 consumer. Family Q13a, QS6 \texttt{lock\_\allowbreak{}free\_\allowbreak{}access}. & Lamport 1977 \\
\texttt{Lock\allowbreak{}Free\allowbreak{}MPMCBuffer} & Lock-Free Bounded MPMC Queue (Vyukov 2010--11): Multi-Producer/\allowbreak{}Multi-Consumer via per-cell atomic sequence-number and CAS-loop. \texttt{Cell.sequence} == pos (producer writes), then pos+1 (consumer reads). Retries on contention (wait-free per op NOT guaranteed). Capacity MUST be power-of-2. \texttt{iterable\_\allowbreak{}aspect\_\allowbreak{}t}. Family Q13b, QS6 \texttt{lock\_\allowbreak{}free\_\allowbreak{}access}. \texttt{is\_\allowbreak{}original\_\allowbreak{}eligible}=true (Vyukov BSD-2). & Vyukov 1024cores 2010--11, Michael/\allowbreak{}Scott PODC 1996 \\
\texttt{Original\allowbreak{}Lock\allowbreak{}Free\allowbreak{}Mpmc\allowbreak{}Concurrent\allowbreak{}Queue} & \texttt{moodycamel::Concurrent\allowbreak{}Queue} (Cameron Desrochers, block-based MPMC): Habich-compliance wrapper with original-code linking (2/\allowbreak{}6 methods original: put, get; 4/\allowbreak{}6 gaps: emplace, peek\_\allowbreak{}front, peek\_\allowbreak{}back, clear are re-impl). \texttt{is\_\allowbreak{}original\_\allowbreak{}module()}=false due to gap-filling. Header-only, BSD-2/\allowbreak{}BSL-1.0 dual. Analog \texttt{Lock\allowbreak{}Free\allowbreak{}MPMCBuffer} but with true submodule binding. Family Q15, QS6 \texttt{lock\_\allowbreak{}free\_\allowbreak{}access}. \texttt{is\_\allowbreak{}original\_\allowbreak{}eligible}=true (moodycamel BSD-2). & moodycamel 2014, BSD-2/\allowbreak{}BSL-1.0 \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} LMAX 2011, P01 (Leis ART ICDE 2013), Bw-Tree (Levandoski et al.\ ICDE 2013), P29 (RCU OLS 2001), Lamport 1977, Vyukov 1024cores 2010--11, Michael/\allowbreak{}Scott PODC 1996, Pugh CACM 1990, O'Neil LSM-tree 1996, Driscoll et al.\ 1989, Masstree Euro\allowbreak{}Sys 2012, SMART OSDI 2023, Williams 1964, moodycamel 2014~\cite{thompson2011disruptor,leis2013art,levandoski2013bwtree,mckenney2001rcu,lamport_spsc,vyukov_mpmc,michael1996queue,pugh1990skiplist,oneil1996lsm,driscoll1989persistent,mao2012masstree,luo2023smart,williams1964heapsort,desrochers_concurrentqueue}.

\section{T18 Buffering Q2 (Flush Policy)}

This axis is the \emph{buffering organ (Q2)} (T18) of the living being \texttt{SearchAlgorithm} (cf.\ Section~\ref{sec:sota-instances}).

Axis Q2 defines automatic flush strategies for write buffers. It determines WHEN a filled buffer propagates its output to the next stage --- after every operation (Eager), never automatically (Lazy), when threshold is exceeded (Watermark), after time window (Timed), or adaptively by workload frequency via EWMA-controlled LSM-style logic. This is central to memory/\allowbreak{}throughput tradeoff and batching efficiency in cache engines.

\subsection*{Static sub-axes}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
FS1 & Event-Triggered & Flush is triggered by explicit events (put/\allowbreak{}get operation or eviction). Encompasses Eager\allowbreak{}Flush (after every operation) and Lazy\allowbreak{}Flush (only on eviction). \\
FS2 & Threshold-Triggered & Flush is triggered when buffer fill exceeds a threshold (default 75\%). Reduces unnecessary flushes in sparse traffic. \\
FS3 & Time-Triggered & Flush is triggered periodically after a time window (micro-batching), independent of buffer fill. Classic Hadoop/\allowbreak{}Spark/\allowbreak{}Kafka pattern for bounded-latency guarantees. \\
FS4 & Adaptive-Triggered & Flush threshold is adaptively learned from workload frequency via EWMA\@. High write-rate $\rightarrow$ flush early (60\% threshold); low rate $\rightarrow$ flush late (95\% threshold). First Q2 strategy with self-tuning. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}
\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{2.6cm} >{\raggedright\arraybackslash}p{6.5cm} >{\raggedright\arraybackslash}p{4.4cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Eager\allowbreak{}Flush} & Pessimistic Write-Through policy: flushes completely (Full\allowbreak{}Flush) after EVERY \texttt{put()} operation. Lowest latency, but maximum flush frequency and highest throughput overhead --- classic synchronous-write pattern. & Textbook Write-Through Policy /\allowbreak{} Code Baseline \\
\texttt{Lazy\allowbreak{}Flush} & Optimistic Write-Back policy: never flushes automatically, only on explicit eviction from buffer. Maximum batching, minimum flush frequency --- classic deferred-write pattern. \texttt{should\_\allowbreak{}flush()} always returns No\allowbreak{}Flush. & Textbook Write-Back Policy /\allowbreak{} Code Baseline \\
\texttt{Watermark\allowbreak{}Flush} & Threshold-driven flush policy (default 75\% capacity). Iterable aspect with 5 variants (50\%, 65\%, 75\%, 85\%, 95\%) for runtime permutation. Related to LSM-Tree Memtable-Trigger (O'Neil 1996). & LSM-Tree O'Neil 1996 (related) /\allowbreak{} Code Implementation \\
\texttt{Timed\allowbreak{}Flush} & Time-window-based micro-batching policy (iterable aspect: 10/\allowbreak{}100/\allowbreak{}1000/\allowbreak{}10000 ms). Inspired by Kafka (Net\allowbreak{}DB 2011) and Spark D-Streams (SOSP 2013) for bounded-latency guarantees. Independent of buffer fill. & Kafka Net\allowbreak{}DB 2011 /\allowbreak{} Spark SOSP 2013 /\allowbreak{} Code Implementation \\
\texttt{Adaptive\allowbreak{}Lsm\allowbreak{}Flush} & Self-tuning adaptive watermark policy with EWMA over burst-rate. Threshold interpolates between 60\% (high write-rate) and 95\% (low rate). Conceptually Rocks\allowbreak{}DB Dynamic Leveled Compaction + O'Neil LSM 1996; EWMA logic is CE-specific. First Q2 strategy with \texttt{is\_\allowbreak{}adaptive=true}. & Rocks\allowbreak{}DB Dynamic Leveled Compaction /\allowbreak{} O'Neil LSM-Tree 1996 /\allowbreak{} Code Implementation \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} O'Neil et al.\ 1996 (LSM-Tree), Kafka (Net\allowbreak{}DB 2011), Spark D-Streams (SOSP 2013)~\cite{oneil1996lsm,kreps2011kafka,zaharia2013dstreams}.

\section{Build-PG Page Type Axis (PAGE-TYPE)}

The page type axis defines the structural and indexing strategies for branch pages in the trie index: from dense direct-addressed byte indexing (Dense\allowbreak{}Byte), through multi-level density variants (Extended\allowbreak{}Dense, Sparse\allowbreak{}Patricia), to path-collapse optimizations (Redirect, Custom\allowbreak{}Cache) and B+-style sorted pages (BPlus). All page types are cache-line aware and implement the \texttt{Page\allowbreak{}Type\allowbreak{}Strategy} concept interface with guaranteed interchangeability (F15 permutation).

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
PG1 & Structure Role (Branching) & Classification of branch page architecture: full-branching branch page, collapsed single-path node (Redirect), or specialized cache-optimized variant. \\
PG2 & Density Class (Density) & Degree of slot utilization and resulting indexing strategy: densely populated (0-256 slots), extended (>256 slots), or sparse with Patricia compression. \\
PG3 & Path Collapse Strategy & Handling of unique remainder paths: no compression (standard), or collapse via Redirect/\allowbreak{}Co\allowbreak{}Co technique for space minimization. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{8.0cm} >{\raggedright\arraybackslash}p{2.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Dense\allowbreak{}Byte\allowbreak{}Page\allowbreak{}Type} & Dense byte-indexed branch page with direct 256-slot addressing (direct-addressed). Page type 1: direct mapping from key byte to slot index without indirection. Source: Adaptive Radix Tree (ART), Leis et al. 2013. & P01 \\
\texttt{Extended\allowbreak{}Dense\allowbreak{}Page\allowbreak{}Type} & Extended dense page for >256 entries with multi-level indexing scheme (e.g., 2-level lookup or variable k-byte span). Page type 2: density class for complex slot arrangements. Source: Adaptive Radix Tree extensions, PRT-ART\@. & P01 (extended) \\
\texttt{Sparse\allowbreak{}Patricia\allowbreak{}Page\allowbreak{}Type} & Sparsely populated page with Patricia path compression (single-bit split). Page type 3: density class with bit-vector traversal for low branching factor. Source: HOT (Height Optimized Trie, Binna et al. 2018, based on Patricia Trie, Morrison 1968). & P02 \\
\texttt{Redirect\allowbreak{}Page\allowbreak{}Type} & Collapsed unique remainder path (single-path node). Page type 4: path-collapse strategy for space optimization under monopoly unique successor sequences. Neither classical branch nor leaf, stores only suffix until next real branching or value node. Source: Co\allowbreak{}Co-Trie (Boffa et al. 2024). & P04 \\
\texttt{Custom\allowbreak{}Cache\allowbreak{}Page\allowbreak{}Type} & Cache-line-oriented custom page with explicit layout design for 64-byte alignment. Page type 5: structure role branch with hardware-aware geometry. Source: PRT-ART proprietary implementation with TLB-offset + cache-line-aligned memory layout (Axis 5). & PRT-ART \\
\texttt{BPlus\allowbreak{}Page\allowbreak{}Type} & B+-style sorted branch page with range-walk traversal. Page type 6: structure role branch, but with sorted key-array organization instead of byte indexing. Source: Masstree (Mao et al. 2012), B+-tree family. & P03 \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} P01, P02, P03, P04, Morrison 1968, PRT-ART~\cite{leis2013art,binna2018hot,mao2012masstree,boffa2024coco,morrison1968patricia}.

\emph{Related literature:} P05, P06, P10, P11, P12, P13, P20~\cite{fent2020start,schmeisser2022b2tree,zhang2018surf,rao1999css,rao2000csb,hankins2003nodesize,mueller2025btreesback}.

\section[Build-SIMD: Vector/Accelerator Extension Axis]{Build-SIMD: Vector/\allowbreak{}Accelerator Extension Axis (Hardware Accelerators)}

The \texttt{axis\_\allowbreak{}09b\_\allowbreak{}simd\_\allowbreak{}extension} models available SIMD/\allowbreak{}accelerator extensions of the CPU instruction set architecture as a sub-axis to \texttt{axis\_\allowbreak{}09} (main ISA). It enables compile-time selection among various vector extensions (SSE2/\allowbreak{}AVX2/\allowbreak{}AVX-512 for x86; NEON/\allowbreak{}SVE2 for ARM\@; RVV for RISC-V) plus baseline (scalar) and GPU acceleration (CUDA), with backward-cumulative layer hierarchy and hardware-specific topology (SIMD units per socket, P/\allowbreak{}E core access).

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
SE1 & Vector Width & Compile-time-known maximum vector width of the SIMD extension in bits: 0 (none), 128, 256, 512, or -1 (scalable). Determines lane width and parallelization degree. \\
SE2 & Accelerator Type & Categorizes accelerator class: none (baseline), simd (vector SIMD), matrix (matrix engine), gpu (GPU), npu (neural processing unit). Influences composition with hardware topology (\texttt{axis\_\allowbreak{}12}). \\
SE3 & Compatibility Family & Constraint to main ISA (\texttt{axis\_\allowbreak{}09}): none (universal), x86, arm, riscv, cuda. Ensures compile-time validation: e.g., \texttt{Sse2\allowbreak{}Simd\allowbreak{}Extension} compatible only with \texttt{Amd64\allowbreak{}Isa}. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{7.5cm} >{\raggedright\arraybackslash}p{3.0cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{No\allowbreak{}Simd\allowbreak{}Extension} & Baseline variant without SIMD acceleration. Scalar fallback strategy for all permutations without accelerator. Universally compatible with all main ISAs (x86/\allowbreak{}ARM/\allowbreak{}RISC-V/\allowbreak{}Power\allowbreak{}PC). Mandatory comparison baseline. & Code/\allowbreak{}Baseline \\
\texttt{Sse2\allowbreak{}Simd\allowbreak{}Extension} & Intel SSE2 128-bit vector extension (x86\_\allowbreak{}64 ABI baseline since 2003). Compatible only with \texttt{Amd64\allowbreak{}Isa}. Baseline on x86\_\allowbreak{}64; provides 4× uint32 lanes/\allowbreak{}iteration. Standard on cluster nodes Barnard/\allowbreak{}Capella. & Intel ISA Ext Ref 319433-060 \\
\texttt{Avx2\allowbreak{}Simd\allowbreak{}Extension} & Intel AVX2 256-bit vector extension (Haswell+ 2013). X86\_\allowbreak{}64 only. Backward-cumulative: includes all SSE+AVX\@. Default optimization level for server DBMS\@. 2 AVX2 units/\allowbreak{}socket. ZIH standard on Barnard (Intel Sapphire Rapids) and Capella (AMD EPYC 9334). & Intel ISA Ext Ref 319433 \\
\texttt{Avx512\allowbreak{}Simd\allowbreak{}Extension} & Intel AVX-512 512-bit vector extension (Skylake-X+, Ice Lake Server, Zen 4+). Backward-cumulative: all SSE/\allowbreak{}AVX/\allowbreak{}AVX2 + 15 separate flags (VNNI for vector indexes, VPOPCNTDQ for bitmap indexes). Not on all x86\_\allowbreak{}64 CPUs (e.g., Haswell/\allowbreak{}Raptor Lake disabled). 1 AVX-512 unit/\allowbreak{}socket; E-cores no access (Intel Hybrid). & Intel ISA Ext Ref 319433-060 \\
\texttt{Neon\allowbreak{}Simd\allowbreak{}Extension} & ARM NEON /\allowbreak{} Advanced SIMD 128-bit (AArch64 ABI baseline ARMv8+). Compatible only with \texttt{Aarch64\allowbreak{}Isa}. Baseline on all AArch64 CPUs (Apple M-Series, AWS Graviton 2/\allowbreak{}3/\allowbreak{}4, NVIDIA Grace). 1 NEON unit/\allowbreak{}socket; all cores accessible (big.LITTLE architecture). & ARM ARM DDI 0487 /\allowbreak{} Cortex-A8 TRM \\
\texttt{Sve2\allowbreak{}Simd\allowbreak{}Extension} & ARM SVE2 (Scalable Vector Extension 2) ARMv9 mandatory (scalable 128--2048 bit). Compatible only with \texttt{Aarch64\allowbreak{}Isa}. ZIH cluster Grace Hopper (NVIDIA Grace = ARM Neoverse V2 with 128-bit SVE2). Scalable lane width; 1 unit/\allowbreak{}socket; all cores. & SVE~\cite{stephens2017sve} (IEEE Micro 2017); SVE2/\allowbreak{}Armv9 per ARM ARM~\cite{arm_arm} \\
\texttt{Rvv\allowbreak{}Simd\allowbreak{}Extension} & RISC-V Vector Extension (RVV v1.0, scalable VLEN 128--65536 bit). Compatible only with \texttt{Risc\allowbreak{}VIsa}. Optional (not baseline like SSE2/\allowbreak{}NEON). Growing relevance: Si\allowbreak{}Five Performance P870 (128-bit), T-Head Xuantie C908 (128-bit). Research boards at TU Dresden/\allowbreak{}ZIH\@. & RVV-v1.0-Spec~\cite{riscv_vector_spec_1_0}; Vitruvius+ (TACO 2023)~\cite{minervini2023vitruvius} \\
\texttt{Cuda\allowbreak{}Gh200\allowbreak{}Simd\allowbreak{}Extension} & NVIDIA CUDA on Hopper GH200 (Grace Hopper Superchip GPU accelerator). ZIH cluster Alpha Centauri with NVIDIA HGX\@. Compatible with \texttt{Aarch64\allowbreak{}Isa} (Grace CPU) AND \texttt{Amd64\allowbreak{}Isa} (Hopper host). NVLink-C2\allowbreak{}C 900 GB/\allowbreak{}s between Grace CPU and Hopper GPU\@. Massive parallel (96 GB HBM3 GPU\@; \textasciitilde16,896 CUDA cores). \texttt{units\_\allowbreak{}per\_\allowbreak{}socket()} = -1 (GPU is separate device). & NVIDIA Tesla~\cite{lindholm2008tesla} (IEEE Micro 2008) + NVIDIA GH200 Whitepaper \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Sources:} ISA specifications and whitepapers (Intel, ARM, IEEE Micro, ACM TACO, NVIDIA); see the source column of the building blocks.

\section{Build-HW — Platform Hardware Strategy}

Defines the platform hardware family (CPU architecture, SIMD capabilities, cache topology, memory paging) as compile-time constants for all cache-engine algorithms. The axis captures which platforms are configurable at build time and provides static properties (cache-line size, page size, SIMD bit-width, NUMA capability, hugepage capability) for algorithm-appropriate data structures.

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
\texttt{hw1\_\allowbreak{}simd\_\allowbreak{}family} & 12.1 SIMD-Family & Which vector instruction set is actively used? (AVX2, AVX-512, NEON, SVE2, scalar/\allowbreak{}none). Determines the maximum vector width for parallel operations within an algorithm. \\
\texttt{hw2\_\allowbreak{}cache\_\allowbreak{}level\_\allowbreak{}targeting} & 12.2 Cache-Level-Targeting & Which cache level is optimized (L1-aware, L2-aware, L3-aware, HBM-aware)? Critical for prefetch strategies and memory layout decisions. \\
\texttt{hw3\_\allowbreak{}numa\_\allowbreak{}strategy} & 12.3 NUMA-Strategy & How is NUMA topology considered (single-node, multi-node, NUMA-aware-pinning, NUMA-balanced)? Affects allocator locality and thread-pinning strategy. \\
\texttt{hw4\_\allowbreak{}prefetch\_\allowbreak{}hardware} & 12.4 Prefetch-Hardware-Instructions & Which explicit hardware prefetch instructions are available (PREFETCH, PREFETCHNTA, PREFETCHW, none)? Core part of distance-estimator and path-oriented prefetch. \\
\texttt{hw5\_\allowbreak{}atomic\_\allowbreak{}instruc\allowbreak{}tion\_\allowbreak{}family} & 12.5 Atomic-Instruction-Family & Which atomic read-modify-write instructions are available (CAS, LL-SC, RMW-extended via HTM/\allowbreak{}XBEGIN/\allowbreak{}XEND, none)? Fundamental basis for lock-free concurrency and optimistic lock coupling. \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.0cm} >{\raggedright\arraybackslash}p{2.3cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{X86\_\allowbreak{}64\allowbreak{}Hardware\allowbreak{}Profile} & Modern x86-64 desktop/\allowbreak{}server platform (Intel/\allowbreak{}AMD). Build-time constants: \texttt{cache\_\allowbreak{}line\_\allowbreak{}size}=64, \texttt{memory\_\allowbreak{}page\_\allowbreak{}size}=4096, \texttt{simd\_\allowbreak{}width\_\allowbreak{}bits}=256 (AVX2 conservative), \texttt{numa\_\allowbreak{}capable}=true, \texttt{huge\_\allowbreak{}page\_\allowbreak{}capable}=true. Baseline for desktop and server-side experiments. & Code/\allowbreak{}Baseline \\
\texttt{Aarch64\allowbreak{}Hardware\allowbreak{}Profile} & ARM64/\allowbreak{}AArch64 server and mobile platforms (Ampere Altra, Apple Silicon M-series, Linux ARM). Build-time constants: \texttt{cache\_\allowbreak{}line\_\allowbreak{}size}=64, \texttt{memory\_\allowbreak{}page\_\allowbreak{}size}=4096, \texttt{simd\_\allowbreak{}width\_\allowbreak{}bits}=128 (NEON standard), \texttt{numa\_\allowbreak{}capable}=true, \texttt{huge\_\allowbreak{}page\_\allowbreak{}capable}=true. Enables cross-platform comparisons. & Code/\allowbreak{}Baseline \\
\texttt{Generic\allowbreak{}Hardware\allowbreak{}Profile} & Conservative fallback profile for platform-agnostic configurations. Build-time constants: \texttt{cache\_\allowbreak{}line\_\allowbreak{}size}=64, \texttt{memory\_\allowbreak{}page\_\allowbreak{}size}=4096, \texttt{simd\_\allowbreak{}width\_\allowbreak{}bits}=0 (scalar-only), \texttt{numa\_\allowbreak{}capable}=false, \texttt{huge\_\allowbreak{}page\_\allowbreak{}capable}=false. Lowest common denominator when no specialized adapter matches. & Code/\allowbreak{}Baseline \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Sources:} platform profiles as code/\allowbreak{}baseline (no external papers).

\section{A-Korpus Allocator Corpus (A01-A23)}

The allocator axis documents 25 memory management strategies (identifiers A01--A21, the three standard-library baselines A22a--A22c, and A23) affecting dynamic allocation in search workloads. It covers textbook allocators (Buddy, Slab), industrial standards (dlmalloc, jemalloc, TCMalloc, mimalloc, snmalloc), specialized variants (NUMA-aware, PIM-Malloc, cache-aware), and modern hardened implementations (Star\allowbreak{}Malloc). All 25 wrappers satisfy the \texttt{Allocator\allowbreak{}Strategy} concept and enable systematic measurement of memory management's impact on search tree performance (F15 operativity).

\subsection*{Static sub-axes}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.8cm} >{\raggedright\arraybackslash}p{3.2cm} >{\raggedright\arraybackslash}p{8.5cm}@{}}
\toprule
ID & Name & Description \\
\midrule
\endhead%
AA1 & Free-List Topology & Organization of free memory blocks: per-page sharding (mimalloc), per-thread caching (snmalloc, rpmalloc, TCMalloc), buddy tree (Buddy), single global free-list (dlmalloc), per-heap locks (Hoard). \\
AA2 & Size-Class Schema & Classification of block sizes: slab-per-object-size (Slab), dlmalloc bins, jemalloc size-classes, scalloc spans, Buddy power-of-2. \\
AA3 & Thread Locality & Degree of thread isolation: tcmalloc thread-local caches, rpmalloc per-thread spans, snmalloc message-passing, Hoard per-heap locks. \\
AA4 & Synchronization & Concurrency mechanism: lock-free CAS (Michael, LRMalloc), per-heap locks (Hoard), wait-free (Crystalline), mutex-based. \\
AA5 & Allocation Policy & Routing of allocation requests: NUMA-origin-aware (NUMAlloc), cache-set-aware (CAMA), NUCA-aware (TCMalloc-Warehouse), PIM-aware (PIM-Malloc). \\
AA6 & Reclamation & Reclamation of deleted blocks: magazine cache (Slab), page-heap release (TCMalloc), lock-free reclamation (LRMalloc), deferred free (mimalloc). \\
AA7 & Fragmentation Strategy & Fragmentation avoidance: dlmalloc coalescing, object coloring (Slab), buddy splitting, page-local sharding (mimalloc), span pooling (scalloc). \\
\bottomrule
\end{longtable}
\end{scriptsize}

\subsection*{Building blocks}

\begin{scriptsize}
\begin{longtable}{@{}>{\raggedright\arraybackslash}p{3.0cm} >{\raggedright\arraybackslash}p{6.0cm} >{\raggedright\arraybackslash}p{4.5cm}@{}}
\toprule
Variant & Description & Source \\
\midrule
\endhead%
\texttt{Hoard\allowbreak{}Allocator} (A01) & Per-heap free-lists with SMP scalability. Three-tier structure: thread-local heaps, global heap for cross-thread migration, false-sharing reduction via segregated allocation. & A01~\cite{berger2000hoard} (Berger/\allowbreak{}Mc\allowbreak{}Kinley/\allowbreak{}Blumofe/\allowbreak{}Wilson, ASPLOS 2000, 10.1145/\allowbreak{}378993.379232) + Code/\allowbreak{}Baseline \\
\texttt{Slab\allowbreak{}Allocator} (A02) & Object-caching kernel allocator. Slab-per-object-size with per-CPU magazine cache (Bonwick 1994/\allowbreak{}2001). Classic Solaris kernel implementation for pre-allocated object pools. & A02~\cite{bonwick1994slab} (Bonwick, USENIX 1994) + Code/\allowbreak{}Baseline \\
\texttt{Michael\allowbreak{}Lock\allowbreak{}Free\allowbreak{}Allocator} (A03) & Lock-free CAS allocator per Michael 2004. Hazard-pointer-protected lock-free data structures, MIT re-implementation with IBM patent background. & A03~\cite{michael2004lockfree} (Michael, PLDI 2004, 10.1145/\allowbreak{}996841.996848) + Code/\allowbreak{}Baseline \\
\texttt{Mimalloc\allowbreak{}Allocator} (A04) & Free-list sharding (3 free-lists per page). Microsoft Research 2019: deferred\_\allowbreak{}free hooks, temporal cadence, page-local sharding.\ is\_\allowbreak{}original: MIT direct binding. & A04~\cite{leijen2019mimalloc} (Leijen/\allowbreak{}Zorn/\allowbreak{}de Moura, APLAS 2019, 10.1007/\allowbreak{}978-3-030-34175-6\_\allowbreak{}13) + is\_\allowbreak{}original (github.com/\allowbreak{}microsoft/\allowbreak{}mimalloc, MIT) \\
\texttt{Jemalloc\allowbreak{}Allocator} (A05) & Arena-based size-class allocator. Evans 2006 /\allowbreak{} Free\allowbreak{}BSD\@. Per-CPU arenas with tcache layer for thread-local caches. Facebook standard.\ is\_\allowbreak{}original: BSD-2-Clause direct binding. & A05~\cite{evans2006jemalloc} (Evans, BSDCan 2006, papers.freebsd.org/\allowbreak{}\ldots/\allowbreak{}jemalloc) + is\_\allowbreak{}original (github.com/\allowbreak{}jemalloc/\allowbreak{}jemalloc, BSD-2-Clause) \\
\texttt{TCMalloc\allowbreak{}Allocator} (A06) & Thread-cache + central-cache + page-heap. Google gperftools 2005. Later TEMERAIRE (OSDI 2021) with warehouse-scale optimizations. & A06~\cite{hunter2021temeraire} (TCMalloc 2005; TEMERAIRE, OSDI 2021) + Code/\allowbreak{}Baseline \\
\texttt{Snmalloc\allowbreak{}Allocator} (A07) & Message-passing allocator. Paul Liétar et al. ISMM 2019: free operations via lock-free queue to allocating thread, cross-thread lock contention avoidance\@. is\_\allowbreak{}original: MIT direct binding. & A07~\cite{lietar2019snmalloc} (Liétar, ISMM 2019, 10.1145/\allowbreak{}3315573.3329980) + is\_\allowbreak{}original (github.com/\allowbreak{}microsoft/\allowbreak{}snmalloc, MIT) \\
\texttt{Scalloc\allowbreak{}Allocator} (A08) & Span-based, virtual-spans technique. Aigner/\allowbreak{}Kirsch/\allowbreak{}Lippautz/\allowbreak{}Sokolova OOPSLA 2015. Multicore-scalable with low fragmentation. & A08~\cite{aigner2015scalloc} (Aigner et al., OOPSLA 2015) + Code/\allowbreak{}Baseline \\
\texttt{NUMAlloc\allowbreak{}Allocator} (A09) & NUMA-origin-aware allocator. UTSASRG ISMM 2023. Node hint for local NUMA allocation, multi-node systems optimized. & A09~\cite{yang2023numalloc} (Yang et al., ISMM 2023) + Code/\allowbreak{}Baseline \\
\texttt{RPMalloc\allowbreak{}Allocator} (A10) & Per-thread span caching. Mattias Jansson 2017, OSS without paper. Special case: mandatory init (rpmalloc\_\allowbreak{}initialize, rpmalloc\_\allowbreak{}thread\_\allowbreak{}initialize)\@. is\_\allowbreak{}original: Public-Domain/\allowbreak{}MIT direct binding. & Code/\allowbreak{}Baseline (github.com/\allowbreak{}mjansson/\allowbreak{}rpmalloc, Public-Domain/\allowbreak{}MIT) + is\_\allowbreak{}original \\
\texttt{LRMalloc\allowbreak{}Allocator} (A11) & Lock-free + hazard pointers. Leite/\allowbreak{}Rocha VECPAR 2018 (LNCS 11333). Modern lock-free allocator with ABA protection.\ is\_\allowbreak{}original: MIT direct binding. & A11~\cite{leite2018lrmalloc} (Leite/\allowbreak{}Rocha, VECPAR 2018, 10.1007/\allowbreak{}978-3-030-15996-2\_\allowbreak{}17) + is\_\allowbreak{}original (github.com/\allowbreak{}ricleite/\allowbreak{}lrmalloc, MIT) \\
\texttt{CAMAAllocator} (A12) & Cache-aware memory allocator. Herter/\allowbreak{}Backes/\allowbreak{}Haupenthal/\allowbreak{}Reineke ECRTS 2011 (Saarland). Predictable cache behavior, no public code reference. & A12~\cite{herter2011cama} (Herter et al., ECRTS 2011) + Code/\allowbreak{}Baseline \\
\texttt{Star\allowbreak{}Malloc\allowbreak{}Allocator} (A13) & Formally verified hardened allocator. Inria-Prosecco (Reitz/\allowbreak{}Fromherz/\allowbreak{}Protzenko et al.) OOPSLA 2024. F*/\allowbreak{}Steel formal verification, Apache-2.0 license. & A13~\cite{reitz2024starmalloc} (Reitz et al., OOPSLA 2024) + Code/\allowbreak{}Baseline \\
\texttt{TCMalloc\allowbreak{}Warehouse\allowbreak{}Allocator} (A14) & TCMalloc-Warehouse /\allowbreak{} Hyperscale. Google Cloud OSDI 2021 /\allowbreak{} ASPLOS 2024. Hugepage-aware variant, marketing variant without separate artifact. & A14~\cite{hunter2021temeraire,zhou2024tcmallocwarehouse} (Hunter OSDI 2021 / Zhou ASPLOS 2024) + Code/\allowbreak{}Baseline \\
\texttt{HMalloc\allowbreak{}Allocator} (A15) & Hybrid, scalable, lock-free memory allocator. Li/\allowbreak{}Yao/\allowbreak{}Tang/\allowbreak{}Lin ICPADS 2019. Hybrid approach for scalability, no public code reference. & A15~\cite{li2019hmalloc} (Li/\allowbreak{}Yao/\allowbreak{}Tang/\allowbreak{}Lin, ICPADS 2019, 10.1109/\allowbreak{}ICPADS47876.2019.00064) + Code/\allowbreak{}Baseline \\
\texttt{PIMMalloc\allowbreak{}Allocator} (A16) & Processing-in-memory allocator. VIA-Research HPCA 2026 (ar\allowbreak{}Xiv:2505.13002). Memory distribution across compute nodes, PIM hardware optimized. & A16~\cite{pimmalloc2026} (VIA-Research, HPCA 2026, arxiv.org/\allowbreak{}abs/\allowbreak{}2505.13002) + Code/\allowbreak{}Baseline \\
\texttt{Crystalline\allowbreak{}Reclamation} (A17) & Wait-free / lock-free safe-memory-reclamation family for non-blocking data structures. Not a standalone malloc/\allowbreak{}free allocator but a reclamation policy. & A17~\cite{nikolaev2021crystalline,nikolaev2024crystalline} (Nikolaev/\allowbreak{}Ravindran, DISC 2021 / PLDI 2024) + Code/\allowbreak{}Baseline \\
\texttt{Exgen\allowbreak{}Allocator} (A18) & Single-threaded specialized allocator. UT Austin IEEE CAL 2025 (ar\allowbreak{}Xiv:2510.10219)\@. \enquote{Old is Gold} optimizations for single-threaded workloads, no public code reference. & A18~\cite{li2025exgen} (UT Austin, IEEE CAL 2025, arxiv.org/\allowbreak{}abs/\allowbreak{}2510.10219) + Code/\allowbreak{}Baseline \\
\texttt{Buddy\allowbreak{}Allocator} (A19) & Buddy system (power-of-2 splitting). Knuth TAo\allowbreak{}CP Vol.1 (1968), CACM 1965. Classic textbook allocator, binary tree splitting without OSS code reference. & A19~\cite{knowlton1965buddy} (Knuth/\allowbreak{}Carlson, CACM 1965, 10.1145/\allowbreak{}365628.365655) + Code/\allowbreak{}Baseline \\
\texttt{Dlmalloc\allowbreak{}Allocator} (A20) & Doug Lea malloc (bins-based). Lea 1987--2012, public-domain (CC0). Classic since 1987, bins size-class schema, oldest production implementation.\ is\_\allowbreak{}original: public-domain direct binding. & A20~\cite{lea1996dlmalloc} (Lea, gee.cs.oswego.edu/\allowbreak{}\ldots/\allowbreak{}malloc.html, 1996) + is\_\allowbreak{}original (github.com/\allowbreak{}ennorehling/\allowbreak{}dlmalloc, Public-Domain/\allowbreak{}CC0) \\
\texttt{Pt\allowbreak{}Malloc2\allowbreak{}Allocator} (A21) & ptmalloc2 (glibc malloc). Wolfram Gloger 1990s-2000s. LGPL-2.1+ copyleft, linking blocked (no is\_\allowbreak{}original). Standard basis for glibc-based systems. & Code/\allowbreak{}Baseline (malloc.de/\allowbreak{}en/\allowbreak{}) + License restriction \\
\texttt{Std\allowbreak{}Malloc} (A22a) & Standard libc malloc (proof of concept), portable fallback allocator. Baseline for all standard-library implementations without vendor linking. & Code/\allowbreak{}Baseline (C++ Standard Library) + Baseline \\
\texttt{Pmr\allowbreak{}Resource\allowbreak{}Allocator} (A22b) & std::pmr::memory\_\allowbreak{}resource. WG21 N3916 (2014 $\rightarrow$ C++17). Polymorphic Memory Resources standard library, fallback to new/\allowbreak{}delete allocator. & Code/\allowbreak{}Baseline (C++17 Standard Library, open-std.org/\allowbreak{}\ldots/\allowbreak{}n3916.pdf) + Baseline \\
\texttt{Pool\allowbreak{}Resource\allowbreak{}Allocator} (A22c) & std::pmr::unsynchronized\_\allowbreak{}pool\_\allowbreak{}resource. WG21 N3916 (C++17). Own size-class pool, F15-operative (first wrapper with own memory retention behavior without vendor linking). & Code/\allowbreak{}Baseline (C++17 Standard Library, open-std.org/\allowbreak{}\ldots/\allowbreak{}n3916.pdf) + is\_\allowbreak{}operative \\
\texttt{Vmem\allowbreak{}Magazines\allowbreak{}Allocator} (A23) & Vmem + magazines (slab extension). Bonwick 2001 USENIX ATC\@. Per-CPU magazines for scalability, kernel classic without public code reference. & A23~\cite{bonwick2001vmem} (Bonwick, USENIX ATC 2001, usenix.org/\allowbreak{}\ldots/\allowbreak{}bonwick.html) + Code/\allowbreak{}Baseline \\
\bottomrule
\end{longtable}
\end{scriptsize}

\emph{Contributing sources:} cited per allocator A01--A23 in the source column of the building-block table.

\setlength{\tabcolsep}{6pt}

