\chapter{Architecture Decisions}\label{app:adr}

% Architecture Decision Records ADR-1..ADR-12, extracted verbatim-faithfully from the
% consolidated, user-verified decisions in the project ledger
% (DIPLOMARBEIT-ZIELE-OFFENE-PUNKTE-LEDGER.md, sections 18/19/20 + goal V6), 2026-07-16/17.
% Only documented decisions are recorded; no invented architecture. The earlier F1--F15
% round (docs/bausteine/04_...) is historical and partly superseded (newer facts win).

This appendix records the load-bearing architecture decisions as Architecture Decision Records
(ADRs). Each ADR states the context, the decision and its consequences, and carries the date on
which it was fixed. The decisions are the consolidated, verified choices of the architect (the
author, in the supervising role); they were established across the design sessions of
2026-07-16/17 and supersede an earlier decision round where they conflict. The code structure they
govern is documented in Appendix~\ref{app:code}.

\section{ADR-1: Recursive dock architecture}\label{asec:adr-dock}
\emph{Date: 2026-07-16/17.}
\paragraph{Context.} The framework must both compose measurement systems and drive the individual
tier binaries, at two different granularities, while keeping every boundary binary-stable.
\paragraph{Decision.} The same dock/ABI-stable-shared-object/contract pattern is applied
recursively on two levels. The \emph{experiment dock} (planner~$\leftrightarrow$~cache-engine
builder, CEB) is \emph{one-directional}: the planner steers the settings of the CEB it compiles and
receives no return channel. The \emph{check dock} (CEB~$\leftrightarrow$~tier binary) is
\emph{bidirectional}: the CEB drives the tier binary and the tier binary reports back ABI-stably.
The CEB is itself a system-axis-versioned shared object; it writes the requested tasks and measured
values itself to the target given in the experiment XML, not through the one-directional experiment
dock.
\paragraph{Consequences.} The Anatomy major version rides on the experiment-planner dock. The two
levels share their concepts, so the pattern is genuinely recursive. Result write-back is decoupled
from the drive path.

\section{ADR-2: Compile-time-only composition (CRTP + concept guard)}\label{asec:adr-crtp}
\emph{Date: 2026-07-16 (standing doctrine).}
\paragraph{Context.} The design space is reassembled by permutation, not by runtime adaptation; the
hot path must not pay for dynamic dispatch.
\paragraph{Decision.} On the hot path, composition is compile-time only: axes are CRTP bases guarded
by concepts, with no vtable and no runtime switch. Any compile-time-to-runtime degradation is
corrected back. Runtime polymorphism is permitted only internally, never at the product interfaces
or the ABI. Only named textbook design patterns are used; ad-hoc constructions are disallowed.
\paragraph{Consequences.} Each composed configuration is monomorphised per point; the axis roof
\texttt{topics::Axis} enforces the guard (empty, non-polymorphic) so that an accidentally introduced
vtable fails to compile. The orphaned runtime-vtable scheduling strategy is deprecated and replaced
(see ADR-7).

\section{ADR-3: Allocators are algorithms of the allocator axis}\label{asec:adr-alloc}
\emph{Date: 2026-07-16.}
\paragraph{Context.} Allocation is one of the cache-aware design decisions and must be measurable on
equal footing with the other axes.
\paragraph{Decision.} An allocator wrapper \emph{is} the algorithm choice, modelled as a compile-time
strategy on the allocator axis (\texttt{axis\_06}). All vendor wrappers live in one central type
list; a compile-time filter selects the enabled subset, and code generation bakes exactly one choice
into each per-permutation binary.
\paragraph{Consequences.} Allocator selection is a static axis value, never a runtime switch; this
keeps per-permutation measurement fair and directly motivates the DLL-load decision (ADR-5), because
each shared object must carry its own allocator archive.

\section{ADR-4: Four level-one genera, each with a head framework}\label{asec:adr-genera}
\emph{Date: 2026-07-16/17.}
\paragraph{Context.} Structurally different data structures (ordered map, sequential container, set,
graph) must be expressed as points of the same design space behind one interface, without collapsing
their distinct outer hulls.
\paragraph{Decision.} There are four level-one genera --- map (the search-algorithm hull),
(sequential) container, set and graph --- each with its own head framework of shared genus concepts.
Each genus is its own genus with its own ABI and only partially reused observers; genera do not
intersect but may use one another. A \emph{set} is its own genus with its own ABI (not a subtype of
sequence); a \emph{view} is a read-only projection of a sequential container.
\paragraph{Consequences.} Promoting set from a sub-genus of container to a level-one genus is part of
the coordinated ABI bump (ADR-7); the genera and their enforcement are non-negotiable.

\section{ADR-5: Dynamic-library loading via Option B (self-linked vendor archive)}\label{asec:adr-dll}
\emph{Date: 2026-07-17.}
\paragraph{Context.} The runtime build path for a per-permutation shared object did not link the
vendor allocator archive, so \texttt{dlopen} failed on undefined allocator symbols and no
measurements were produced.
\paragraph{Decision.} Each permutation shared object self-links its own vendor allocator archive
(Option~B). The rejected Option~A (host-side symbol export via \texttt{-rdynamic}) was found
doctrine-adverse on five counts: it breaks the allocator-as-algorithm axis through one process-wide
allocator instance with arena contamination between permutations, breaks the destroy contract,
violates the self-contained ABI, breaks the truthfulness of the build-variant metadata, and couples
static dispatch to runtime symbol resolution.
\paragraph{Consequences.} Each shared object is loaded under \texttt{RTLD\_LOCAL} with private
allocator symbols, giving fair per-library isolation and a clean teardown. The fix is purely
additive and ABI-compatible; it restored non-zero measurement output and unblocked the real
measurement run (see Appendix~\ref{app:measurements}).

\section{ADR-6: System axes over tier-binary axes (the target layering)}\label{asec:adr-layering}
\emph{Date: 2026-07-16/17.}
\paragraph{Context.} The CEB-level configuration axes were scattered and flat: all axes were
reflected into one flat list, with no shared layer separating the system level from the organ level.
\paragraph{Decision.} System axes (at CEB level, compile-time static, versioned) sit \emph{over} the
tier-binary axes (organs plus governor/prefetcher). The CEB is built under its system axes, which in
turn determine how the subordinate tier binaries are built statically; the compile order is
system configuration first, then the measurement/heuristic configuration. The CEB configuration
system axes are: scheduling, hardware/ISA, telemetry regime, load and load frameworks, compiler
(\texttt{gcc}\,$\vert$\,\texttt{clang}) and extension hardware (see ADR-12) --- six in total. System
axes get their own abstract root (``blood''); organ and system axes are never mixed.
\paragraph{Consequences.} A dedicated system-axis level is split off above the nineteen organ axes;
this layering was introduced ABI-compatibly, without a major bump, closing the layering gap between
the model and the implementation.

\section{ADR-7: The single coordinated 4\,$\to$\,5 ABI bump}\label{asec:adr-abi}
\emph{Date: 2026-07-17.}
\paragraph{Context.} Four changes each break the same ABI boundary (module symbols, POD layouts, the
genus discriminator and the \texttt{binary\_id} convention). Doing them separately would touch the
frozen ABI boundary four times and re-materialise the golden set four times.
\paragraph{Decision.} They are bundled into one coordinated Anatomy major bump (4\,$\to$\,5) at the
experiment-planner dock: telemetry moved out of the \texttt{binary\_id} into a system axis; the set
level-one promotion; the native set ABI; and scheduling as a compile-time CRTP system axis (replacing
the orphaned runtime vtable). The magic marker advances so the loader cleanly rejects all
old-major libraries. This is the one reserved major bump; it proceeds only under the coordinated
approval, with one backup and review.
\paragraph{Consequences.} The check dock becomes the single bidirectional contract carrier (the
former iterator bypass is unified). Documentation and measurement data are never deleted; the bump
carries a per-increment verification checkpoint showing the concrete POD/ABI diff.

\section{ADR-8: Measurement-preserving rebuild of the golden set}\label{asec:adr-golden}
\emph{Date: 2026-07-17.}
\paragraph{Context.} The golden pilot set is a fixed 320-point configuration that also serves as a
round-trip gate. The coordinated ABI bump (ADR-7) breaks it three ways (system axes leave the path,
the genus discriminator changes, the major advances).
\paragraph{Decision.} The golden set is deliberately decomposed and rebuilt rather than kept
byte-identical, but strictly measurement-preserving: the old fixtures and their measurement
CSV/PDF are frozen additively and untouched; the serialisation is \emph{extended, not replaced}; the
set is re-materialised once per genus; and the gate is re-pointed, keeping the old set as a
historical reference.
\paragraph{Consequences.} No measurement datum is ever lost; the previous round-trip invariant is
replaced by the measurement-preserving rebuild only inside the one coordinated bump.

\section{ADR-9: Serialisation --- flat system-axis folders, deep tier tree, dynamic-only columns}\label{asec:adr-serial}
\emph{Date: 2026-07-16.}
\paragraph{Context.} Encoding every axis as a CSV column is redundant, because the static and system
axes are constant within a run.
\paragraph{Decision.} System axes are serialised \emph{flat} as a folder structure; the tier-binary
axes together with the CSV form a \emph{deeper} serialised permutation tree. Static compile-time axes
and system axes do not become columns; they move into folder and file labels and into a sidecar
manifest. Only the dynamic sub-axes (workload, resource control, P/E-core, NUMA, multithreading)
remain columns. On disk this is a tree plus index; results are written first into the build path and
then atomically copied to the single output target named in the experiment XML.
\paragraph{Consequences.} The result layout is compact and non-redundant; provenance for the static
and system choices is carried beside the binary rather than inflating every row.

\section{ADR-10: Failure visibility --- \texttt{failed} cell instead of null}\label{asec:adr-failure}
\emph{Date: 2026-07-16.}
\paragraph{Context.} A silently missing value cannot be distinguished from a genuine zero, which would
let a mis-programmed axis or an ISA/OS incompatibility disappear.
\paragraph{Decision.} A failed axis algorithm yields the literal cell value ``failed'' (never null),
together with an aggregated log written next to the CSV. The harness keeps measuring the remaining
configurations.
\paragraph{Consequences.} Mis-programming and platform incompatibilities become visible in the result
tables; measurement is never disabled by writing zeros.

\section{ADR-11: Measurement modes --- debug parallel, measurement single-thread}\label{asec:adr-modes}
\emph{Date: 2026-07-16.}
\paragraph{Context.} Debugging benefits from full parallelism, whereas a fair latency measurement must
avoid harness-induced scheduling noise.
\paragraph{Decision.} In debug mode all threads run in parallel; in measurement mode a single
measurement thread is used, with the genus threading taken from the tier binary rather than
overridden by the harness. Measurement is never switched off by writing zeros. From the same
experiment XML the planner generates either the CI code or the local CMake code, synchronously.
\paragraph{Consequences.} Measurement runners are laid out as ISA\,$\times$\,OS\,$\times$\,system-axes,
disjoint and parallel across ISA, with concurrency one per measuring machine.

\section{ADR-12: Extension-hardware system axis and compiler axis (Q2, Option C)}\label{asec:adr-q2}
\emph{Date: 2026-07-17.}
\paragraph{Context.} Instruction-set extensions (such as SIMD width) change the compiled tier binary
and hence its measurements, and must be controlled explicitly without polluting the \texttt{binary\_id}.
\paragraph{Decision.} Option~C: the CEB receives its frame from the experiment planner but itself
permutes the extension-hardware value (initially the SIMD extension) at its own runtime, and for each
value compiles and measures tier binaries at compile time. The extension hardware (SIMD, later GPU)
is therefore its own system axis; its value is passed as compiler flags at the compile-function seam,
its provenance is recorded in the build-variant sidecar, and it never enters the \texttt{binary\_id}.
The compiler (\texttt{gcc}\,$\vert$\,\texttt{clang}) is a further full system axis, so that the
planner can build and compare with both compilers.
\paragraph{Consequences.} This yields the sixth CEB system axis (extension hardware) and the compiler
axis of ADR-6; the hardware/ISA coupling drives \texttt{-march}/\texttt{-mavx} selection at the
compile seam, and the versioned build-variant metadata (with an AVX10 field) documents each static
compilation beside its binary.
